### 2021-11-26 Manuscript.Rmd
Add manuscript, datamaking and fitted models

---
title: "Do NBA Teams Avoid Trading within Their Own Division?^[We appreciate feedback we received on this paper from Skye Bender-deMoll, Pavel Krivitsky and David Schaefer.]"
author: "jimi adams & Michał Bojanowski"
date: "`r format(Sys.time(), '%d %B, %Y')`"
abstract: "Within US professional sports, trades within one's own division are often perceived to be disadvantageous. We ask how common this practice is. To examine this question, we construct a date-stamped network of all trades in the NBA between June 1976 and May 2019. We then use yearly weighted exponential random graph models to estimate the likelihood of teams avoiding within-division trade partners, and whether that pattern changes through time. In addition to the empirical question, this analysis serves to demonstrate the necessity and difficulty of constructing the proper baseline for statistical comparison. We find limited-to-no support for the popular perception."
output: 
  bookdown::html_document2:
    keep_md: true
link-citations: true
bibliography: nba_trades.bib
---

```{r setup, include=FALSE, cache=FALSE}
requireNamespace("knitr")
library(dplyr)
knitr::opts_chunk$set(
  echo = FALSE
)
```

# Introduction

A common refrain among sports commentators is that professional sports teams avoid trading players with other teams from within their own division [@wong2017; @simmons]. Some offered conjectures on why this should be avoided include: not wanting to improve the competitiveness of a team's direct rivals [@ley2017], especially in leagues where games within the division are more common than other match-ups [@bates2015], or wanting to avoid fans being reminded of players they gave up [@fs2015], especially if they turn out to play better for their new teams [@ley2017]. Despite the frequency of this speculation, attempts to quantify the actual frequency of such a prohibition has been rare [for one limited exception, see @ahr2018].

Beyond the popular perception of the avoidance of intra-division trades, there is also ample scholarly literature that would lead us to the same expectation, particularly in professional sports [@Stewart1999]. For example, within a field where the actors are (or are perceived to be) competitors with one another, they may avoid cooperating with one another because it is perceived to be a competitive disadvantage [@Hoffmann2018], though that assumption has been questioned [@Bengtsson1999; @peng12]. Instead of cooperating directly with one's competitors, actors may therefore seek out means of collaboration with those outside the competitive field [@Soda_2017]. For example, in the case of NBA player development, this could lead to a variety of strategic patterns. Teams may prioritize trading with teams they do not perceive to be direct competitors [@Barman_2002], such as those outside their own division. Alternatively, general managers may develop strong stable relationships as resources [@Elfenbein2014], for example with particular player agents to provide comparative advantages in access to players on the free agent market (thus reducing their reliance on trades). Teams also could carve out recruitment niches that avoid direct competition [@Soltis_2010; @Barman_2002]. For example, recent shifts in player "apprentice" opportunities through the developmental league and foreign partnerships has opened opportunities for teams to gain contractual rights to players outside of the trade system [@Keiper_2020]. Each of these possibilities could bolster teams' ability to avoid cooperating through trading with their direct competitors.

In sum, we therefore investigate how strongly division shapes trade partners among National Basketball Association (NBA) franchises, focused especially on whether teams tend to avoid trading with other teams within their own division.

# Data

Here, we draw on a database of all player transactions in the NBA from the beginning of the 1976 season---when the NBA merged with the ABA uniting major professional basketball in the US under one league---through the completion of the 2018-19 season [@richardson2020]. We compile this list of 1,977 trades into 43 annual trade networks, with the nodes representing teams, and each edge representing a unique trade between those teams. Each trade is assigned to the "trade season" which runs from the end of the previous season through the corresponding season's trade deadline.^[The opening of the trade season begins either (1) with the conclusion of the previous season's playoffs (early in the observed period), or on a specified date each summer (later in the observed window). Using these dates, there were no ambiguous trades that we were unable to attribute to a particular season.] The number of teams within each annual slice changes over time as the league gradually expanded from 22 to 30 teams.^[Unless otherwise specifically noted, all historical accounts in the text were pulled from Wikipedia and confirmed on nba.com.]. The number of trades observed within each slice varies from year to year (range 22-86; see Figure 1), and teams exhibit differing rates of trading (see Figure 2).

![Figure 1. **Number of Trades by Year.** (normalized to a per team rate)](EDA_files/figure-gfm/trades-per-season-seasonwise-1.png)

Figure 1 presents the trend of number of trades per season. Since, the number of teams varies across the window, we standardize these values to represent the number of trades *per team* in each season. This pattern shows a roughly U-shaped trend, with the highest rates of trades (nearing 2 per team) occurring early and late in the period and the lowest rates occurring in the early 1990s. Notably, this low-point came *after* the introduction of unrestricted free-agency (in 1988), allowing players to sign with any team after their contract expires. As a corollary, Figure 2 presents the frequency distribution of trades per team across the full observed window. There are some teams (e.g., the Spurs) who consistently trade less frequently, and others who are much more trade active (e.g., the Mavericks).

![Figure 2. **Number of Trades by Team.** (normalized to a per season rate)](EDA_files/figure-gfm/trades-per-season-teamwise-1.png)

To address our research question, we needed to supplement these trade data by compiling a list of each team's division membership. From the 1976-77 season through 2003-04, there were four divisions (Atlantic, Central, Midwest, and Pacific). From the 2004-05 season onward, there were six divisions (Atlantic, Central, Southeast, Northwest, Pacific, and Southwest).^[The divisions are aggregated into conferences, with the Atlantic, Central, and Southeast being in the Eastern Conference, and the Midwest, Pacific, Northwest, and Southwest being in the Western Conference.] While those division/conference names otherwise remained stable, teams changed divisions at various times, whether for geographic reasons or as new expansion teams were added to the league. Accordingly, each team's division is assigned as current to a particular trade season (for example, the Buffalo Braves were in the Atlantic division in 1977-78, but became the San Diego Clippers and moved to the Pacific division for the 1978-79 season). Combined, these data allow us to investigate the tendency for teams to avoid trading with other teams in their own division.

# Analysis and Findings

Analytically, we proceed in three steps, which allow us to address two simultaneous aims. Primarily, these steps allow us to build an appropriate test of our research question. Secondarily, these steps also allow us to illustrate the proper way to statistically condition a question such as this, and to explain the need for doing so. This combination motivates our usage of a relatively recent development for statistically modeling weighted network data.

```{r, echo=FALSE}
htmltools::tags$iframe(src = 'NBA_Trades.html', width = '80%', height = '600px')
```

Figure 3. **Dynamic Visualization of Annual Trade Networks** Note: Intra-division trades are highlighted in red. Node color and position refers to division.

Our first step is to tabulate the number of trades observed both within and across divisions. This tabulation is presented in the form of what is referred to as a mixing matrix, which shows---at the level of divisions---how many trades occur between teams of the same division and between teams of differing divisions. We present these mixing matrices in two aggregations. The first represents all years prior to division expansion, when there were four divisions, a period which spans from 1976-2004. The second window began in 2005 and runs through the end of our analytic period at the end of the 2018-19 season, during which the teams were separated into six divisions. To provide a visual representation of the changes in these division-specific trade patterns across the examined period, Figure 3 presents a dynamic visualization of each year's trade network.^[The appendices also include full replication code for all analyzes conducted and the generation of this visualization.]

```{r mm-1976-2004}
mm_1976_2004 <- readRDS("mm_1976_2004.rds")
withr::with_options(
  list(knitr.kable.NA = ""),
  mm_1976_2004 %>%
    rename(`**1976 - 2004**` = div_from) %>%
    knitr::kable(
      caption = "**Mixing Matrix for the period 1976 - 2004**"
    )
)
```

```{r mm-2005-2019}
mm_2005_2019 <- readRDS("mm_2005_2019.rds")
withr::with_options(
  list(knitr.kable.NA = ""),
  mm_2005_2019 %>%
    rename(`**2005 - 2019**` = div_from) %>%
    knitr::kable(
      caption = "**Mixing Matrix for the period 2005 - 2019**"
    )
)
```

As can be seen in Tables \@ref(tab:mm-1976-2004) and \@ref(tab:mm-2005-2019), these tabulations would appear to suggest that there is an avoidance of trading with members of one's own division. For example in the pre-expansion era only 22% (231/1062) of trades were within division, a ratio of nearly 3.6 trades across divisions for every one trade within division.

While these mixing matrices may appear to support the prohibition against teams trading within their division, there are two caveats for interpreting these patterns which we address in each of our subsequent analytic steps. In order to test for whether teams avoid trading within their division we must first account for each team's tendency to trade at all, before we can assess with whom they conduct those trades. For example, in the 1976-77 trade season, there were 35 trades, of which 7 were within division and 28 across divisions. If we account for how many trades each team made, yet assumed that they made those trades without respect to division of their trade partner, random expectation would be that on average 7.6 trades (sd 2.2) would be within division and 27.4 (sd=2.2) would be across divisions. That is, this observed mixing ratio of 4.0 trades across division for every one within is actually consistent with the random expectations, and thus does not conform to a within division avoidance pattern.

Our second analytic step therefore addresses these conditional distribution assumptions that are not apparent from the mixing matrices alone. To accomplish these properly conditioned versions of the test we use what is known as the $\alpha$-index for computing segregation in networks [@moody2001]. Essentially, what the $\alpha$-index does is allow the estimates of frequency of ties between teams from the same division to be estimated as an odds ratio, which is properly conditioned to account for the number of trade partners one would expect to be within and across division, given the observed base rates of trading by each team, *if* they chose with whom to trade without respect to division [@Bojanowski:2014aa]. Since this is a homophily statistic, testing for nodes within the *same* division, a significantly *negative* $\alpha$-index would be consistent with teams following the proposed prohibition. Figure 4 presents this $\alpha$-index for each trade season individually. What can be seen here is that in none of the observed trade years does the $\alpha$-index differ significantly from random expectation (OR=1). In other words this improved conditional form of the test suggests that there is no observed prohibition against teams trading within division.

![Figure 4. **Divisional Homophily in NBA Trades by Season.** (as computed by Moody's $\alpha$-index )](segregation_files/figure-gfm/or2-1.png)

Our third analytic step addresses one additional caveat necessary for appropriate interpretation. The $\alpha$-index used above only allows for ties in the network to be dichotomous. That is, within each trade season the test only asks whether each pair of teams trades with one another or not. However in these data it is possible (and observed) that some pairs of teams trade with each other multiple times within the same trade season. So we further need a version of the test that allows for the weighting of the edges in the network, to properly allow for these multiple trades to be incorporated into the estimate. We rely on a recently developed advance in Exponential-family Random Graph Models [ERGM, @strauss-ikeda1990;@lusher-etal2013;@handcock2019], that allow for estimation with weighted networks [@krivit2012;@krivitsky2019]. ERGM is a statistical model for explaining the structure of a network by means of various local tendencies for ties to be present or absent quantified by model terms such as density, degree, homophily, transitivity and so on. The magnitude of effect of each term is measured by an associated coefficient. In the player trade networks we are interested in verifying whether there is any divisional homophily effect, i.e. analogously to the above presented $\alpha$. This would indicated if (multiple) trades are more or less likely in pairs of teams belonging to the same division compared to to pairs in different divisions. A positive value of the coefficient would suggest homophily while negative values would indicate heterophily (the hypothesized avoidance of within-division trading).

Once properly conditioning the test for each of these necessary caveats, we find that in none of the seasons are teams more likely to avoid trading with other teams from within their division. As can be verified in the "funnel plot" presented in Figure 5, each the seasonal homophily coefficients fall within the white cone of statistical *in*significance.

![Figure 5. **Homophily Coefficients and Standard Errors from Weighted Exponential Random Graph Models.** (including base-rate effects)](ergm-results_files/figure-gfm/B-nodematch-funnel-allinone-1.png)

To summarize the results from Figure 5 in a single model, we also fit a pooled ERGM across all seasons, including a base-rate effect for each year's trade volume. These results are presented in Table \@ref(tab:model1-table-short).

```{r model1-table-short}
 withr::with_options(
   list(knitr.kable.NA = ""),
   readRDS("model1-table-short.rds") |>
     knitr::kable(
       caption = "**Valued ERGM fitted to pooled seasonal data**. The table excludes seasonal constants, complete table is presented in the Appendix.",
       align = "lrrrr"
     )
 )
```

We find an aggregate same-division homophily effect of 0.0099 (standard error of 0.059, z-statistic of 0.169), again showing no avoidance pattern. Fitting similar models to the periods 1976-2004 and 2004-2019 separately leads to almost identical results.[^periodModels]

[^periodModels]: Results can be obtained from the authors upon request.

# Discussion

In short, the empirical patterns do not support the popular claim that NBA teams are likely to avoid trading with teams from within their own division. However, the lack of support for this prohibition is only observed once the strategy for testing the question first accounts for the differential rates at which teams trade and therefore have the opportunity to trade with teams from within/outside their division. The lingering weak associations are even further reduced once including the weighting that is possible from multiple trades between partners in a single season.

It is worth mentioning that the models we have estimated in the results for Figure 5 are actually no different from what would have been possible through a generalized linear model predicting the (weighted) number of ties observed between pairs of teams. However, the value of introducing the weighted ERGM approach with these data is that if subsequent researchers want to take the next step to provide an explanatory account that further examines which teams are likely to trade with which others, this modeling framework would allow for the simultaneous estimation of predictive factors that variously operate at the node level (e.g., is a team that finished lower in the previous year's standings more likely to trade), the dyad level (e.g., are teams with higher total salaries more likely to trade with other teams who have lower total salaries), or structural features above the dyad (e.g., the tendency to form trade-triples or larger configurations through multi-team deals). The GLM framework would not be able to estimate each of these types of features in such a model without violating model assumptions (e.g., some of these are explicitly modeling the *dependency* between multiple trades). As such, we hope that the illustration if how the weighted ERGM could be beneficial serves as a useful introduction to this modeling framework, which could be extended by future researchers using the data we provide.

# Appendix

## ERGM model fit

Table: Full results for the ERG model fit to pooled data.

|Effect                |   Estimate|        SE|95% CI         | p-value|
|:---------------------|----------:|---------:|:--------------|-------:|
|*Seasonal base rates* |           |          |               |        |
|1976-1977             | -0.9026534| 0.0792530|(-1.06; -0.75) |   0.000|
|1977-1978             | -0.8427925| 0.0724760|(-0.98; -0.70) |   0.000|
|1978-1979             | -0.9263500| 0.0796106|(-1.08; -0.77) |   0.000|
|1979-1980             | -0.8762352| 0.0771426|(-1.03; -0.73) |   0.000|
|1980-1981             | -0.9910653| 0.0854415|(-1.16; -0.82) |   0.000|
|1981-1982             | -0.9050884| 0.0754043|(-1.05; -0.76) |   0.000|
|1982-1983             | -0.8069531| 0.0682278|(-0.94; -0.67) |   0.000|
|1983-1984             | -0.9092292| 0.0766022|(-1.06; -0.76) |   0.000|
|1984-1985             | -1.0526877| 0.0872369|(-1.22; -0.88) |   0.000|
|1985-1986             | -1.0782139| 0.0930195|(-1.26; -0.90) |   0.000|
|1986-1987             | -0.9198185| 0.0757693|(-1.07; -0.77) |   0.000|
|1987-1988             | -0.9403157| 0.0781448|(-1.09; -0.79) |   0.000|
|1988-1989             | -1.1192392| 0.0854069|(-1.29; -0.95) |   0.000|
|1989-1990             | -1.1674130| 0.0852806|(-1.33; -1.00) |   0.000|
|1990-1991             | -1.1015000| 0.0789942|(-1.26; -0.95) |   0.000|
|1991-1992             | -1.3314738| 0.1014767|(-1.53; -1.13) |   0.000|
|1992-1993             | -1.2131200| 0.0885597|(-1.39; -1.04) |   0.000|
|1993-1994             | -1.2949074| 0.0953019|(-1.48; -1.11) |   0.000|
|1994-1995             | -1.3952511| 0.1080338|(-1.61; -1.18) |   0.000|
|1995-1996             | -1.2254082| 0.0852894|(-1.39; -1.06) |   0.000|
|1996-1997             | -1.2243585| 0.0837553|(-1.39; -1.06) |   0.000|
|1997-1998             | -1.1148373| 0.0727527|(-1.26; -0.97) |   0.000|
|1998-1999             | -1.2492779| 0.0861211|(-1.42; -1.08) |   0.000|
|1999-2000             | -1.2789135| 0.0904522|(-1.46; -1.10) |   0.000|
|2000-2001             | -1.0430409| 0.0700097|(-1.18; -0.91) |   0.000|
|2001-2002             | -1.0608807| 0.0696335|(-1.20; -0.92) |   0.000|
|2002-2003             | -1.2653868| 0.0888219|(-1.44; -1.09) |   0.000|
|2003-2004             | -1.0915922| 0.0733866|(-1.24; -0.95) |   0.000|
|2004-2005             | -1.0507651| 0.0685301|(-1.19; -0.92) |   0.000|
|2005-2006             | -1.0769347| 0.0703190|(-1.21; -0.94) |   0.000|
|2006-2007             | -1.1925306| 0.0774736|(-1.34; -1.04) |   0.000|
|2007-2008             | -1.1603188| 0.0759557|(-1.31; -1.01) |   0.000|
|2008-2009             | -0.9697841| 0.0623303|(-1.09; -0.85) |   0.000|
|2009-2010             | -0.9605862| 0.0613118|(-1.08; -0.84) |   0.000|
|2010-2011             | -0.9335487| 0.0588953|(-1.05; -0.82) |   0.000|
|2011-2012             | -1.1465080| 0.0721213|(-1.29; -1.01) |   0.000|
|2012-2013             | -0.9791587| 0.0622563|(-1.10; -0.86) |   0.000|
|2013-2014             | -0.9789426| 0.0643977|(-1.11; -0.85) |   0.000|
|2014-2015             | -0.8316791| 0.0542273|(-0.94; -0.73) |   0.000|
|2015-2016             | -1.0257114| 0.0647017|(-1.15; -0.90) |   0.000|
|2016-2017             | -1.0875715| 0.0680398|(-1.22; -0.95) |   0.000|
|2017-2018             | -1.0105469| 0.0640282|(-1.14; -0.89) |   0.000|
|2018-2019             | -0.9172200| 0.0586872|(-1.03; -0.80) |   0.000|
|*Homophily*           |  0.0099434| 0.0587577|(-0.11;  0.13) |   0.866|

# References {-}

### 2021-11-26 make_data.Rmd
Add manuscript, datamaking and fitted models

---
title: "Make data files"
output: 
  github_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

---

```{r setup, include=FALSE}
library(tibble)
library(dplyr)
library(tidyr)
library(stringi)
library(lubridate)
library(intervals)
library(igraph)
requireNamespace("here")

knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE
)
```

Processing the files from `data-src` to produce data frames and igraph/network objects for analysis.

The exact way in which the `gl200210.Rdata` was created is not 100% recovered thus we create the data files based on `wgl.Rdata`. Thus some of the code below is not executed until the full pipeline is re-checked.

# Trade data

## From source files

> Code in this section is not run.

Load original data files to separate environments (I'm not sure what object names are and where):

```{r load-data, eval = FALSE}
adata <- new.env()
load(here::here("data-src", "data", "NBA_AnalysisData.RData"), envir=adata)
ls(adata)

tdata <- new.env()
load(here::here("data-src", "data", "NBA_Transactions.RData"), envir=tdata)
ls(tdata)

mdata <- new.env()
load(here::here("data", "NBA_Meta.RData"), envir=mdata)
ls(mdata)

# Add rowid for back checking
main.df <- tdata$main.df %>%
  as_tibble() %>%
  tibble::rowid_to_column(".rowid")
```

Load standings data for team and division names:

```{r load-standings, eval = FALSE}
standings <- readRDS(here::here("data", "standings.rds")) %>%
  mutate(
    team_name = map_chr(strsplit(full_team_name, " "), tail, 1),
    Team = recode(team_name, "SuperSonics"="Sonics") # For joining with `standings`
  )
```

Create edgelist

```{r tradedf, eval=FALSE}
trade.df <- main.df %>%
  as_tibble() %>%
  select(.rowid, Date, Team, Notes) %>%
  mutate( 
    # Trade has "trade with" in `Notes` field...
    is_trade = grepl("trade with ", Notes),
    # ... but does not have the following phrases
    has_cancelled = grepl("cancelled ", Notes),
    has_rescinded = grepl("rescinded ", Notes),
    has_nullified = grepl("nullified ", Notes),
    has_voided = grepl("voided ", Notes)
  ) %>%
  filter(is_trade) %>%
  filter_at(vars(starts_with("has_")), all_vars(!.)) %>%
  select(-starts_with("has_"), -is_trade) %>%
  mutate( 
    # Remove stuff in parentheses from `Notes`
    notes = gsub("\\s*\\([^\\)]+\\)","", Notes),
    # Extract unique words starting with capital letter or digit (prospective team names)
    partner = map(stri_extract_all_regex(
      notes,
      paste0('\\b', paste0(unique(standings$Team), collapse="|"), '\\b'),
    ), unique),
    # Count extracted names
    num_trade_partners = map_int(partner, length)
  ) %>%
  transmute(
    team1 = Team,
    team2 = partner,
    date = Date,
    id = .rowid,
  ) %>%
  unnest("team2")
```

## Load seasonal weighted igraph objects

List of seasonal weighted igraph objects should be created from the items above. At this moment we just load it and create edge and vertex dataframes from it.

```{r load-wgl}
load(here::here("data", "wgl.Rdata"))
```

## Vertex database

A data frame for `season` x `name` with:

- `name` -- time-constant "franchise name"
- `label` -- time-varying team name
- `div` -- division name
- `season`, `season_from`, `season_to` -- season dates

```{r vdb}
vdb <- graphlist %>%
  setNames(nm = vapply(graphlist, graph_attr, character(1), name = "season")) %>%
  lapply(igraph::as_data_frame, what = "vertices") %>%
  bind_rows(.id = "season") %>%
  as_tibble() %>%
  separate(season, into = c("season_from", "season_to"), remove = FALSE) %>%
  mutate(
    across(starts_with("season_"), as.integer)
  )
vdb
```

## Edge database

A data frame for `season` x `from` x `to` where `from < to` with:

- `from`, `to` -- franchise names
- `season`, `season_from`, `season_to` -- season dates
- `weight` -- number of trades between `from` and `to` in `season`

```{r edb}
edb <- graphlist %>%
  setNames(nm = vapply(graphlist, graph_attr, character(1), name = "season")) %>%
  lapply(igraph::as_data_frame, what = "edges") %>%
  bind_rows(.id = "season") %>%
  as_tibble() %>%
  separate(season, into = c("season_from", "season_to"), remove = FALSE) %>%
  mutate(
    across(starts_with("season_"), as.integer),
    .from = pmin(from, to),
    .to = pmax(from, to)
  ) %>%
  select(
    from = .from, 
    to = .to, 
    season, 
    season_from,
    season_to,
    weight
  )
edb
```

## Checks

Team names in `vdb` and `edb` match:

```{r check}
u_vdb_teams <- with(vdb, unique(name))
u_edb_teams <- with(edb, unique(c(from, to)))
stopifnot(all(u_vdb_teams %in% u_edb_teams))
stopifnot(all(u_edb_teams %in% u_vdb_teams))
```

## Save

```{r save}
saveRDS(vdb, file = here::here("data", "df-vdb.rds"))
saveRDS(edb, file = here::here("data", "df-edb.rds"))
```

<!-- Include folded session info -->
<details>
<summary>Session info</summary>
```{r session-info}
sessioninfo::session_info()
```
</details>

### 2021-11-26 ergm-pooled.R

@@ -1,0 +1,90 @@
+requireNamespace("here")
+requireNamespace("igraph")
+library(ggplot2)
+library(ergm)
+library(ergm.count)
+
+load(here::here("data", "wgl.Rdata"))
+graphlist <- graphlist |>
+  setNames(
+    vapply(graphlist, igraph::graph_attr, character(1), name = "season")
+  ) |>
+  lapply(function(g) {
+    igraph::V(g)$name <- paste0(g$season, "-", igraph::V(g)$name)
+    igraph::V(g)$season <- g$season
+    g$season <- NULL
+    g
+  }) |>
+  head(-1) # Drop 2019-2020
+
+season_from <- names(graphlist) |>
+  substr(1, 4) |>
+  as.numeric()
+
+bigg <- do.call(igraph::disjoint_union, graphlist)
+net <- intergraph::asNetwork(bigg)
+
+library(ergm.count)
+
+
+# Fit model 1 -------------------------------------------------------------
+
+# All seasons/periods
+
+fit <- ergm(
+  net ~ nodefactor("season", levels = TRUE) + nodematch("div"),
+  constraints = ~ blockdiag("season"),
+  reference = ~ Poisson,
+  response = "weight",
+  verbose = TRUE,
+  control = control.ergm(
+    parallel = 7
+  )
+)
+saveRDS(fit, file="ergm-pooled1.rds")
+
+
+# Fit model 2 -------------------------------------------------------------
+
+# Seasons 1976 - 2004
+
+bigg2 <- do.call(igraph::disjoint_union, graphlist[season_from <= 2003])
+net2 <- intergraph::asNetwork(bigg2)
+
+fit2 <- ergm(
+  net2 ~ nodefactor("season", levels = TRUE) + nodematch("div"),
+  constraints = ~ blockdiag("season"),
+  reference = ~ Poisson,
+  response = "weight",
+  verbose = TRUE,
+  control = control.ergm(
+    parallel = 7
+  )
+)
+
+summary(fit2)
+
+saveRDS(fit2, file="ergm-pooled2.rds")
+
+
+# Fit model 3 -------------------------------------------------------------
+
+# Seasons 2005 - 2019
+
+bigg3 <- do.call(igraph::disjoint_union, graphlist[season_from >= 2004])
+net3 <- intergraph::asNetwork(bigg3)
+
+fit3 <- ergm(
[…]

### 2021-11-27 make_data.Rmd
Create graphlist

@@
 ```{r setup, include=FALSE}
-library(tibble)
 library(dplyr)
@@
 
-The exact way in which the `gl200210.Rdata` was created is not 100% recovered thus we create the data files based on `wgl.Rdata`.
-Thus some of the code below is not executed until the full pipeline is re-checked.
+# Trades data
 
-# Trade data
+Load trades data from Zachary Richardson. Group trades into seasons using April 1 of each year.
 
-## From source files
+```{r}
+e <- new.env()
+load(here::here("data-src", "data", "NBA_AnalysisData.RData"), envir=e)
 
-> Code in this section is not run.
-
-Load original data files to separate environments (I'm not sure what object names are and where):
-
-```{r load-data, eval = FALSE}
-adata <- new.env()
-load(here::here("data-src", "data", "NBA_AnalysisData.RData"), envir=adata)
-ls(adata)
-
-tdata <- new.env()
-load(here::here("data-src", "data", "NBA_Transactions.RData"), envir=tdata)
-ls(tdata)
-
-mdata <- new.env()
-load(here::here("data", "NBA_Meta.RData"), envir=mdata)
-ls(mdata)
-
-# Add rowid for back checking
-main.df <- tdata$main.df %>%
-  as_tibble() %>%
-  tibble::rowid_to_column(".rowid")
-```
-
-Load standings data for team and division names:
-
-```{r load-standings, eval = FALSE}
-standings <- readRDS(here::here("data", "standings.rds")) %>%
+trades <- e$trade.df_long |>
+  select(-season) |> # This one is wrong
   mutate(
-    team_name = map_chr(strsplit(full_team_name, " "), tail, 1),
-    Team = recode(team_name, "SuperSonics"="Sonics") # For joining with `standings`
+    season_from = case_when(
+      month(Date) >= 4 ~ as.numeric(year(Date)),
+      month(Date) <= 3 ~ as.numeric(year(Date)) - 1,
+      TRUE ~ as.numeric(NA)
+    ),
+    season_to = season_from + 1,
+    season = paste0(season_from, "-", season_to)
   )
@@
 
-Create edgelist
+# Seasonal networks
 
-```{r tradedf, eval=FALSE}
-trade.df <- main.df %>%
-  as_tibble() %>%
-  select(.rowid, Date, Team, Notes) %>%
-  mutate( 
-    # Trade has "trade with" in `Notes` field...
-    is_trade = grepl("trade with ", Notes),
-    # ... but does not have the following phrases
-    has_cancelled = grepl("cancelled ", Notes),
-    has_rescinded = grepl("rescinded ", Notes),
-    has_nullified = grepl("nullified ", Notes),
-    has_voided = grepl("voided ", Notes)
-  ) %>%
-  filter(is_trade) %>%
-  filter_at(vars(starts_with("has_")), all_vars(!.)) %>%
-  select(-starts_with("has_"), -is_trade) %>%
-  mutate( 
-    # Remove stuff in parentheses from `Notes`
-    notes = gsub("\\s*\\([^\\)]+\\)","", Notes),
-    # Extract unique words starting with capital letter or digit (prospective team names)
-    partner = map(stri_extract_all_regex(
-      notes,
-      paste0('\\b', paste0(unique(standings$Team), collapse="|"), '\\b'),
-    ), unique),
-    # Count extracted names
-    num_trade_partners = map_int(partner, length)
-  ) %>%
-  transmute(
-    team1 = Team,
-    team2 = partner,
-    date = Date,
-    id = .rowid,
-  ) %>%
-  unnest("team2")
+Create a list of igraph objects with edge attribute `weight` counting the trades:
+
+```{r graphlist}
+graphlist <- lapply(
+  sort(unique(trades$season)),
+  function(s) {
+    g <- graph_from_data_frame( 
+      edb %>%
+        filter(season == s) %>%
+        select(P1, P2), 
+      directed=FALSE)
+    E(g)$weight <- count_multiple(g)
+    g <- simplify(g)
+    g$season <- s
+    g
+  }
+)
 ```
 
-## Load seasonal weighted igraph objects
+Since that was built from edges, non-trading teams are inadvertently excluded. Here, we're adding them back in.
 
-List of seasonal weighted igraph objects should be created from the items above. At this moment we just load it and create edge and vertex dataframes from it.
+```{r graphlist-add-nodes}
+graphlist[[3]] <- add.vertices(graphlist[[3]], 1, name="Spurs", label="Spurs", div="Central")
 
-```{r load-wgl}
-load(here::here("data", "wgl.Rdata"))
+graphlist[[4]] <- add.vertices(graphlist[[4]], 1, name="Celtics", label="Celtics", div="Atlantic")
+graphlist[[4]] <- add.vertices(graphlist[[4]], 1, name="Warriors", label="Warriors", div="Pacific")
+
+graphlist[[5]] <- add.vertices(graphlist[[5]], 1, name="76ers", label="76ers", div="Atlantic")
+
+graphlist[[9]] <- add.vertices(graphlist[[9]], 1, name="Nets", label="Nets", div="Atlantic")
+graphlist[[9]] <- add.vertices(graphlist[[9]], 1, name="Suns", label="Suns", div="Pacific")
+
+graphlist[[10]] <- add.vertices(graphlist[[10]], 1, name="Nuggets", label="Nuggets", div="Midwest")
+graphlist[[10]] <- add.vertices(graphlist[[10]], 1, name="Knicks", label="Knicks", div="Atlantic")
+
+graphlist[[11]] <- add.vertices(graphlist[[11]], 1, name="Suns", label="Suns", div="Pacific")
+graphlist[[11]] <- add.vertices(graphlist[[11]], 1, name="Warriors", label="Warriors", div="Pacific")
+
+graphlist[[12]] <- add.vertices(graphlist[[12]], 1, name="Mavericks", label="Mavericks", div="Midwest")
+
+graphlist[[14]] <- add.vertices(graphlist[[14]], 1, name="Lakers", label="Lakers", div="Pacific")
+graphlist[[14]] <- add.vertices(graphlist[[14]], 1, name="Wizards", label="Bullets", div="Atlantic")
+graphlist[[14]] <- add.vertices(graphlist[[14]], 1, name="Jazz", label="Jazz", div="Midwest")
+
+graphlist[[15]] <- add.vertices(graphlist[[15]], 1, name="Pacers", label="Pacers", div="Central")
+graphlist[[15]] <- add.vertices(graphlist[[15]], 1, name="Celtics", label="Celtics", div="Atlantic")
+graphlist[[15]] <- add.vertices(graphlist[[15]], 1, name="Knicks", label="Knicks", div="Atlantic")
+graphlist[[15]] <- add.vertices(graphlist[[15]], 1, name="Cavaliers", label="Cavaliers", div="Central")
+
+graphlist[[16]] <- add.vertices(graphlist[[16]], 1, name="76ers", label="76ers", div="Atlantic")
+graphlist[[16]] <- add.vertices(graphlist[[16]], 1, name="Pacers", label="Pacers", div="Central")
+graphlist[[16]] <- add.vertices(graphlist[[16]], 1, name="Nets", label="Nets", div="Atlantic")
+graphlist[[16]] <- add.vertices(graphlist[[16]], 1, name="Magic", label="Magic", div="Atlantic")
+
+graphlist[[18]] <- add.vertices(graphlist[[18]], 1, name="Heat", label="Heat", div="Atlantic")
+graphlist[[18]] <- add.vertices(graphlist[[18]], 1, name="Celtics", label="Celtics", div="Atlantic")
+
+graphlist[[19]] <- add.vertices(graphlist[[19]], 1, name="Kings", label="Kings", div="Pacific")
+graphlist[[19]] <- add.vertices(graphlist[[19]], 1, name="Nets", label="Nets", div="Atlantic")
+graphlist[[19]] <- add.vertices(graphlist[[19]], 1, name="Bulls", label="Bulls", div="Central")
+graphlist[[19]] <- add.vertices(graphlist[[19]], 1, name="Cavaliers", label="Cavaliers", div="Central")
+
+graphlist[[20]] <- add.vertices(graphlist[[20]], 1, name="Pacers", label="Pacers", div="Central")
+graphlist[[20]] <- add.vertices(graphlist[[20]], 1, name="Rockets", label="Rockets", div="Midwest")
+graphlist[[20]] <- add.vertices(graphlist[[20]], 1, name="Jazz", label="Jazz", div="Midwest")
+graphlist[[20]] <- add.vertices(graphlist[[20]], 1, name="Mavericks", label="Mavericks", div="Midwest")
+
+graphlist[[21]] <- add.vertices(graphlist[[21]], 1, name="76ers", label="76ers", div="Atlantic")
+graphlist[[21]] <- add.vertices(graphlist[[21]], 1, name="Bulls", label="Bulls", div="Central")
+graphlist[[21]] <- add.vertices(graphlist[[21]], 1, name="Cavaliers", label="Cavaliers", div="Central")
+graphlist[[21]] <- add.vertices(graphlist[[21]], 1, name="Spurs", label="Spurs", div="Midwest")
+graphlist[[21]] <- add.vertices(graphlist[[21]], 1, name="Clippers", label="Clippers", div="Pacific")
+
+graphlist[[22]] <- add.vertices(graphlist[[22]], 1, name="Lakers", label="Lakers", div="Pacific")
+graphlist[[22]] <- add.vertices(graphlist[[22]], 1, name="Jazz", label="Jazz", div="Midwest")
+graphlist[[22]] <- add.vertices(graphlist[[22]], 1, name="Wizards", label="Wizards", div="Atlantic")
+graphlist[[22]] <- add.vertices(graphlist[[22]], 1, name="Spurs", label="Spurs", div="Midwest")
+
+graphlist[[23]] <- add.vertices(graphlist[[23]], 1, name="Heat", label="Heat", div="Atlantic")
+graphlist[[23]] <- add.vertices(graphlist[[23]], 1, name="Pacers", label="Pacers", div="Central")
+graphlist[[23]] <- add.vertices(graphlist[[23]], 1, name="Clippers", label="Clippers", div="Pacific")
+
+graphlist[[24]] <- add.vertices(graphlist[[24]], 1, name="Heat", label="Heat", div="Atlantic")
+graphlist[[24]] <- add.vertices(graphlist[[24]], 1, name="Jazz", label="Jazz", div="Midwest")
+graphlist[[24]] <- add.vertices(graphlist[[24]], 1, name="Knicks", label="Knicks", div="Atlantic")
+
+graphlist[[25]] <- add.vertices(graphlist[[25]], 1, name="Timberwolves", label="Timberwolves", div="Midwest")
+graphlist[[25]] <- add.vertices(graphlist[[25]], 1, name="Nets", label="Nets", div="Atlantic")
+
+graphlist[[26]] <- add.vertices(graphlist[[26]], 1, name="Jazz", label="Jazz", div="Midwest")
+
+graphlist[[27]] <- add.vertices(graphlist[[27]], 1, name="Heat", label="Heat", div="Atlantic")
+graphlist[[27]] <- add.vertices(graphlist[[27]], 1, name="Pacers", label="Pacers", div="Central")
+graphlist[[27]] <- add.vertices(graphlist[[27]], 1, name="Timberwolves", label="Timberwolves", div="Midwest")
+graphlist[[27]] <- add.vertices(graphlist[[27]], 1, name="Mavericks", label="Mavericks", div="Midwest")
+
+graphlist[[28]] <- add.vertices(graphlist[[28]], 1, name="Heat", label="Heat", div="Atlantic")
+graphlist[[28]] <- add.vertices(graphlist[[28]], 1, name="Lakers", label="Lakers", div="Pacific")
+
+graphlist[[29]] <- add.vertices(graphlist[[29]], 1, name="Timberwolves", label="Timberwolves", div="Northwest")
+
+graphlist[[30]] <- add.vertices(graphlist[[30]], 1, name="Mavericks", label="Mavericks", div="Southwest")
+graphlist[[30]] <- add.vertices(graphlist[[30]], 1, name="Spurs", label="Spurs", div="Southwest")
+graphlist[[30]] <- add.vertices(graphlist[[30]], 1, name="Warriors", label="Warriors", div="Pacific")
+
+graphlist[[31]] <- add.vertices(graphlist[[31]], 1, name="Heat", label="Heat", div="Southeast")
+graphlist[[31]] <- add.vertices(graphlist[[31]], 1, name="Clippers", label="Clippers", div="Pacific")
+graphlist[[31]] <- add.vertices(graphlist[[31]], 1, name="Kings", label="Kings", div="Pacific")
+graphlist[[31]] <- add.vertices(graphlist[[31]], 1, name="Wizards", label="Wizards", div="Southeast")
+graphlist[[31]] <- add.vertices(graphlist[[31]], 1, name="Knicks", label="Knicks", div="Atlantic")
+
+graphlist[[32]] <- add.vertices(graphlist[[32]], 1, name="Bucks", label="Bucks", div="Central")
+
+graphlist[[33]] <- add.vertices(graphlist[[33]], 1, name="Hawks", label="Hawks", div="Southeast")
+
+graphlist[[34]] <- add.vertices(graphlist[[34]], 1, name="Pacers", label="Pacers", div="Central")
+
+graphlist[[35]] <- add.vertices(graphlist[[35]], 1, name="Pistons", label="Pistons", div="Central")
+
+graphlist[[36]] <- add.vertices(graphlist[[36]], 1, name="Pistons", label="Pistons", div="Central")
+graphlist[[36]] <- add.vertices(graphlist[[36]], 1, name="Suns", label="Suns", div="Pacific")
+
+graphlist[[37]] <- add.vertices(graphlist[[37]], 1, name="Spurs", label="Spurs", div="Southwest")
+
+graphlist[[38]] <- add.vertices(graphlist[[38]], 1, name="Magic", label="Magic", div="Southeast")
+
+graphlist[[39]] <- add.vertices(graphlist[[39]], 1, name="Warriors", label="Warriors", div="Pacific")
+
+graphlist[[41]] <- add.vertices(graphlist[[41]], 1, name="Timberwolves", label="Timberwolves", div="Northwest")
+
+graphlist[[42]] <- add.vertices(graphlist[[42]], 1, name="Spurs", label="Spurs", div="Southwest")
+
+graphlist[[43]] <- add.vertices(graphlist[[43]], 1, name="Warriors", label="Warriors", div="Pacific")
+
+graphlist[[44]] <- add.vertices(graphlist[[44]], 1, name="Raptors", label="Raptors", div="Atlantic")
 ```
+
+Save
+
+```{r graphlist-save}
+saveRDS(graphlist, file = "wgl.rds")
+```
+
+# Edge and node data frames
 

### 2021-11-27 make_data.Rmd
Standings and division mapping

@@
 
+# Divisions
+
+From data on standings
+
+```{r}
+divisions <- readRDS(here::here("data", "standings.rds")) %>%
+  distinct(
+    season_from,
+    division = purrr::map_chr(strsplit(division, " "), 1),
+    team_acronym = purrr::map_chr(strsplit(full_team_name, " "), tail, 1)
+  )
+```
+
 # Seasonal networks
@@
 graphlist <- lapply(
-  sort(unique(trades$season)),
+  sort(unique(trades$season_from)),
   function(s) {
-    g <- graph_from_data_frame( 
-      edb %>%
-        filter(season == s) %>%
-        select(P1, P2), 
-      directed=FALSE)
+    # Select trades
+    e <- trades %>%
+      filter(season_from == s) %>%
+      select(P1, P2)
+    # Nodes
+    n <- data.frame(
+      name = sort(unique(c(e$P1, e$P2)))
+    ) %>%
+      left_join(
+        divisions %>%
+          filter(season_from == s) %>%
+          rename(div = division),
+        by = c("name" = "team_acronym")
+      )
+    g <- graph_from_data_frame(e, vertices = n, directed=FALSE)
     E(g)$weight <- count_multiple(g)
     g <- simplify(g)
-    g$season <- s
+    g$season <- paste0(s, "-", s+1)
     g
@@
 ```{r graphlist-save}
-saveRDS(graphlist, file = "wgl.rds")
+saveRDS(graphlist, file = here::here("data", "wgl.rds"))
 ```

### 2021-11-27 data/standings.R

@@ -1,0 +1,168 @@
+# Download standings
+
+library(tidyverse)
+library(glue)
+library(xml2)
+library(rvest)
+library(stringi)
+library(labelled)
+
+base_url <- "https://www.basketball-reference.com/leagues"
+years <- 1976:2019 # Season (year-1)-year
+urls <- glue("{base_url}/NBA_{years}.html")
+cat(urls, file=here::here("data", "fetchlist.txt"), sep="\n")
+
+# NOW execute `data/fetch.sh` to wget all the pages and save them to
+# `data/standings` folder. What follows assumes that they are there.
+
+# List of downloaded files
+fnames <- list.files(
+  here::here("data", "standings"),
+  "^NBA_[0-9]{4}\\.html",
+  full.names = TRUE
+)
+
+# Extract division standings ----------------------------------------------
+
+# Names of numerical columns
+numcols <- c("W", "L", "W/L%", "GB", "PS/G", "PA/G", "SRS")
+
+east <- fnames %>%
+  lapply(read_html, options=c("RECOVER", "NOERROR", "NOBLANKS", "NONET") ) %>%
+  lapply(html_node, css="table#divs_standings_E") %>%
+  lapply(html_table, fill=TRUE) %>%
+  bind_rows(.id="id_season") %>%
+  rename(
+    team = `Eastern Conference`
+  ) %>%
+  mutate(
+    is_division_header = grepl("Division", team),
+    division = mbtools::carryover(ifelse(is_division_header, team, as.character(NA)))
+  ) %>%
+  filter(!is_division_header) %>%
+  mutate(
+    file = basename(fnames)[as.numeric(id_season)],
+    season_to = as.numeric(stri_extract_first_regex(file, "[0-9]{4}")),
+    season_from = season_to - 1,
+    in_playoffs = grepl("\\*", team),
+    standing = as.integer(stri_extract_last_regex(team, "(?<=\\()[0-9]+(?=\\))")),
+    full_team_name = stri_extract_first_regex(team, "\\b[A-Za-z0-9 ]+(?=\\*?)\\b"),
+    conference = "Eastern"
+  ) %>%
+  select(
+    season_from,
+    season_to,
+    division,
+    standing,
+    full_team_name,
+    one_of(numcols)
+  ) %>%
+  mutate_at(
+    numcols,
+    ~ as.numeric(recode(., "—"=as.character(NA)))
[…]

### 2021-11-29 make_data.Rmd
Teams and franchises

--- 2021-11-27 make_data.Rmd
@@
 
-Processing the files from `data-src` to produce data frames and igraph/network objects for analysis.
+Processing the files created by Zachary Ricardson (in `data-src`) to produce data frames and igraph objects for analysis.
 
-# Trades data
+# Team and franchiese names
 
-Load trades data from Zachary Richardson. Group trades into seasons using April 1 of each year.
+Based on data on standings
 
 ```{r}
-e <- new.env()
-load(here::here("data-src", "data", "NBA_AnalysisData.RData"), envir=e)
+team_franchise <- readRDS(here::here("data", "standings.rds")) %>%
+  distinct(
+    season_from,
+    season_to,
+    team_acronym = purrr::map_chr(strsplit(full_team_name, " "), tail, 1)
+  ) %>%
+  mutate(
+    franchise = case_when(
+      # TODO: map season-teams to franchises
+      season_to <= 1978 & team_acronym == "Braves" ~ "Clippers",
+      season_from >= 2004 & season_to <= 2014 & team_acronym == "Bobcats" ~ "Hornets",
 
-trades <- e$trade.df_long |>
-  select(-season) |> # This one is wrong
+      TRUE ~ team_acronym # else
+    )
+  )
+```
+
+# Data on trades
+
+Process data on trades.
+
+- Assign trades to seasons using April 1 as a cut off date.
+- Columns with team names are called `franchise1` and `franchise2` as the names pretend to "season-invariant" names of the franchises, not actual team names used in during seasons.
+
+```{r}
+load(here::here("data-src", "data", "NBA_AnalysisData.RData"), 
+     envir=NBA_AnalysisData <- new.env())
+
+trades <- NBA_AnalysisData$trade.df_long |>
+  as_tibble() |>
+  select(-season) |>
+  rename(
+    franchise1 = P1,
+    franchise2 = P2
+  ) |>
   mutate(
@@
   )
+
+# Make sure of canonical order
+stopifnot(with(trades, all(franchise1 < franchise2)))
+```
+
+# Divisions
+
+From data on standings extract info on assignment to divisions
+
+- Copy 2018-2019 season to 2019-2020
+
+```{r}
+divisions <- readRDS(here::here("data", "standings.rds")) %>%
+  distinct(
+    season_from,
+    division = purrr::map_chr(strsplit(division, " "), 1),
+    team_acronym = purrr::map_chr(strsplit(full_team_name, " "), tail, 1)
+  )
+divisions <- bind_rows(
+  divisions,
+  divisions %>%
+    filter(season_from == 2018) %>%
+    mutate(season_from = 2019)
+)
 ```
@@
 graphlist <- lapply(
-  sort(unique(trades$season)),
+  sort(unique(trades$season_from)),
   function(s) {
-    g <- graph_from_data_frame( 
-      edb %>%
-        filter(season == s) %>%
-        select(P1, P2), 
-      directed=FALSE)
+    # Select trades
+    e <- trades %>%
+      filter(season_from == s) %>%
+      select(P1, P2)
+    # Nodes
+    n <- data.frame(
+      name = sort(unique(c(e$P1, e$P2)))
+    ) %>%
+      left_join(
+        divisions %>%
+          filter(season_from == s) %>%
+          rename(div = division),
+        by = c("name" = "team_acronym")
+      )
+    g <- graph_from_data_frame(e, vertices = n, directed=FALSE)
     E(g)$weight <- count_multiple(g)
     g <- simplify(g)
-    g$season <- s
+    E(g)$weight <- sqrt(E(g)$weight)
+    g$season <- paste0(s, "-", s+1)
     g

### 2021-11-29 make_data.Rmd
Finish data processing script

--- 2021-11-27 make_data.Rmd
@@
 ---
-title: "Make data files"
+title: "Data processing"
 output: 
@@
 
-Processing the files from `data-src` to produce data frames and igraph/network objects for analysis.
+Producing data frames and igraph objects for analysis.
 
-# Trades data
+# Team, franchises and divisions
 
-Load trades data from Zachary Richardson. Group trades into seasons using April 1 of each year.
+Seasonal team and franchise names along with division memberships based on data on standings from https://www.basketball-reference.com.
 
-```{r}
-e <- new.env()
-load(here::here("data-src", "data", "NBA_AnalysisData.RData"), envir=e)
+```{r teams-franchises-divisions}
+teams <- readRDS(here::here("data", "standings.rds")) %>%
+  distinct(
+    season_from,
+    season_to,
+    team = purrr::map_chr(strsplit(full_team_name, " "), tail, 1) |>
+      recode("Orleans" = "Hornets"),
+    division = gsub("Division", "", division) |> trimws()
+  )
 
-trades <- e$trade.df_long |>
-  select(-season) |> # This one is wrong
+teams <- teams |>
+  bind_rows(
+    filter(teams, season_from == 2018) |>
+      mutate(
+        season_from = 2019,
+        season_to = 2020
+      )
+  ) |>
+  mutate(
+    franchise = case_when(
+      # Map season-teams to franchises
+      season_from <= 1978 & team == "Braves" ~ "Clippers",
+      season_from >= 2004 & season_from <= 2013 & team == "Bobcats" ~ "Hornets",
+      team == "Bullets" & season_from <= 1996 ~ "Wizards",
+      team == "Hornets" & season_from >= 2002 & season_from <= 2012 ~ "Pelicans",
+      team == "SuperSonics" & season_from <= 2007 ~ "Thunder",
+      TRUE ~ team
+    ),
+    season = paste0(season_from, "-", season_to)
+  )
+```
+
+## Checking selected franchise histories
+
+Franchises in columns, time in rows, team names in cells.
+
+```{r selected-histories, R.options=list(knitr.kable.NA="")}
+teams |>
+  filter(franchise %in% c("Hornets", "Bobcats", "Pelicans", "Wizards", "Thunder", "Clippers")) |>
+  select(team, franchise, season_from) |>
+  tidyr::pivot_wider(names_from = franchise, values_from = team) |>
+  arrange(season_from) |>
+  knitr::kable()
+```
+
+# Data on trades
+
+Process data on trades.
+
+- Assign trades to seasons using April 1 as a cut off date.
+- Columns with team names are called `franchise1` and `franchise2` as the names pretend to "season-invariant" names of the franchises, not actual team names used in during seasons.
+- Fix some franchise names mislabeled in the original data.
+
+```{r trades}
+load(here::here("data-src", "data", "NBA_AnalysisData.RData"), 
+     envir=NBA_AnalysisData <- new.env())
+
+trades <- NBA_AnalysisData$trade.df_long |>
+  as_tibble() |>
+  select(-season) |>
   mutate(
@@
     season = paste0(season_from, "-", season_to)
+  ) |>
+  # Fix franchise/team names in trades data
+  mutate(
+    P1 = case_when(
+      P1 == "Pelicans" & season_from >= 1988 & season_from <= 2001 ~ "Hornets",
+      TRUE ~ P1
+    ),
+    P2 = case_when(
+      P2 == "Pelicans" & season_from >= 1988 & season_from <= 2001 ~ "Hornets",
+      TRUE ~ P2
     )
+  ) |>
+  mutate(
+    franchise1 = pmin(P1, P2),
+    franchise2 = pmax(P1, P2)
+  ) |>
+  select(-P1, -P2)
+
+# Make sure of canonical order
+stopifnot(with(trades, all(franchise1 < franchise2)))
+```
+
+## Checks
+
+Are there any franchise names in trade data that we can't match to standings on franchise name and season (expecting 0 rows in both)?
+
+```{r}
+trades %>%
+  distinct(season_from, franchise1) %>%
+  anti_join(teams, by = c("franchise1" = "franchise", "season_from")) %>%
+  print(n = Inf)
+```
+
+```{r}
+trades %>%
+  distinct(season_from, franchise2) %>%
+  anti_join(teams, by = c("franchise2" = "franchise", "season_from")) %>%
+  print(n = Inf)
 ```
@@
 graphlist <- lapply(
-  sort(unique(trades$season)),
+  sort(unique(trades$season_from)),
   function(s) {
-    g <- graph_from_data_frame( 
-      edb %>%
-        filter(season == s) %>%
-        select(P1, P2), 
-      directed=FALSE)
+    # Select trades
+    e <- trades %>%
+      filter(season_from == s) %>%
+      select(franchise1, franchise2)
+    # Nodes
+    n <- teams %>%
+      filter(season_from == s) %>%
+      select(franchise, team, division)
+    g <- graph_from_data_frame(e, vertices = n, directed=FALSE)
     E(g)$weight <- count_multiple(g)
-    g <- simplify(g)
-    g$season <- s
+    g <- simplify(g, edge.attr.comb = list(weight = "min"))
+    g$season <- paste0(s, "-", s+1)
     g
@@
 )
-```
 
-Since that was built from edges, non-trading teams are inadvertently excluded. Here, we're adding them back in.
-
-```{r graphlist-add-nodes}
-graphlist[[3]] <- add.vertices(graphlist[[3]], 1, name="Spurs", label="Spurs", div="Central")
-
-graphlist[[4]] <- add.vertices(graphlist[[4]], 1, name="Celtics", label="Celtics", div="Atlantic")
-graphlist[[4]] <- add.vertices(graphlist[[4]], 1, name="Warriors", label="Warriors", div="Pacific")
-
-graphlist[[5]] <- add.vertices(graphlist[[5]], 1, name="76ers", label="76ers", div="Atlantic")
-
-graphlist[[9]] <- add.vertices(graphlist[[9]], 1, name="Nets", label="Nets", div="Atlantic")
-graphlist[[9]] <- add.vertices(graphlist[[9]], 1, name="Suns", label="Suns", div="Pacific")
-
-graphlist[[10]] <- add.vertices(graphlist[[10]], 1, name="Nuggets", label="Nuggets", div="Midwest")
-graphlist[[10]] <- add.vertices(graphlist[[10]], 1, name="Knicks", label="Knicks", div="Atlantic")
-
-graphlist[[11]] <- add.vertices(graphlist[[11]], 1, name="Suns", label="Suns", div="Pacific")
-graphlist[[11]] <- add.vertices(graphlist[[11]], 1, name="Warriors", label="Warriors", div="Pacific")
-
-graphlist[[12]] <- add.vertices(graphlist[[12]], 1, name="Mavericks", label="Mavericks", div="Midwest")
-
-graphlist[[14]] <- add.vertices(graphlist[[14]], 1, name="Lakers", label="Lakers", div="Pacific")
-graphlist[[14]] <- add.vertices(graphlist[[14]], 1, name="Wizards", label="Bullets", div="Atlantic")
-graphlist[[14]] <- add.vertices(graphlist[[14]], 1, name="Jazz", label="Jazz", div="Midwest")
-
-graphlist[[15]] <- add.vertices(graphlist[[15]], 1, name="Pacers", label="Pacers", div="Central")
-graphlist[[15]] <- add.vertices(graphlist[[15]], 1, name="Celtics", label="Celtics", div="Atlantic")
-graphlist[[15]] <- add.vertices(graphlist[[15]], 1, name="Knicks", label="Knicks", div="Atlantic")
-graphlist[[15]] <- add.vertices(graphlist[[15]], 1, name="Cavaliers", label="Cavaliers", div="Central")
-
-graphlist[[16]] <- add.vertices(graphlist[[16]], 1, name="76ers", label="76ers", div="Atlantic")
-graphlist[[16]] <- add.vertices(graphlist[[16]], 1, name="Pacers", label="Pacers", div="Central")
-graphlist[[16]] <- add.vertices(graphlist[[16]], 1, name="Nets", label="Nets", div="Atlantic")
-graphlist[[16]] <- add.vertices(graphlist[[16]], 1, name="Magic", label="Magic", div="Atlantic")
-
-graphlist[[18]] <- add.vertices(graphlist[[18]], 1, name="Heat", label="Heat", div="Atlantic")
-graphlist[[18]] <- add.vertices(graphlist[[18]], 1, name="Celtics", label="Celtics", div="Atlantic")
-
-graphlist[[19]] <- add.vertices(graphlist[[19]], 1, name="Kings", label="Kings", div="Pacific")
-graphlist[[19]] <- add.vertices(graphlist[[19]], 1, name="Nets", label="Nets", div="Atlantic")
-graphlist[[19]] <- add.vertices(graphlist[[19]], 1, name="Bulls", label="Bulls", div="Central")
-graphlist[[19]] <- add.vertices(graphlist[[19]], 1, name="Cavaliers", label="Cavaliers", div="Central")
-
-graphlist[[20]] <- add.vertices(graphlist[[20]], 1, name="Pacers", label="Pacers", div="Central")
-graphlist[[20]] <- add.vertices(graphlist[[20]], 1, name="Rockets", label="Rockets", div="Midwest")
-graphlist[[20]] <- add.vertices(graphlist[[20]], 1, name="Jazz", label="Jazz", div="Midwest")
-graphlist[[20]] <- add.vertices(graphlist[[20]], 1, name="Mavericks", label="Mavericks", div="Midwest")
-
-graphlist[[21]] <- add.vertices(graphlist[[21]], 1, name="76ers", label="76ers", div="Atlantic")
-graphlist[[21]] <- add.vertices(graphlist[[21]], 1, name="Bulls", label="Bulls", div="Central")
-graphlist[[21]] <- add.vertices(graphlist[[21]], 1, name="Cavaliers", label="Cavaliers", div="Central")
-graphlist[[21]] <- add.vertices(graphlist[[21]], 1, name="Spurs", label="Spurs", div="Midwest")
-graphlist[[21]] <- add.vertices(graphlist[[21]], 1, name="Clippers", label="Clippers", div="Pacific")
-
-graphlist[[22]] <- add.vertices(graphlist[[22]], 1, name="Lakers", label="Lakers", div="Pacific")
-graphlist[[22]] <- add.vertices(graphlist[[22]], 1, name="Jazz", label="Jazz", div="Midwest")
-graphlist[[22]] <- add.vertices(graphlist[[22]], 1, name="Wizards", label="Wizards", div="Atlantic")
-graphlist[[22]] <- add.vertices(graphlist[[22]], 1, name="Spurs", label="Spurs", div="Midwest")
-
-graphlist[[23]] <- add.vertices(graphlist[[23]], 1, name="Heat", label="Heat", div="Atlantic")
-graphlist[[23]] <- add.vertices(graphlist[[23]], 1, name="Pacers", label="Pacers", div="Central")
-graphlist[[23]] <- add.vertices(graphlist[[23]], 1, name="Clippers", label="Clippers", div="Pacific")
-
-graphlist[[24]] <- add.vertices(graphlist[[24]], 1, name="Heat", label="Heat", div="Atlantic")
-graphlist[[24]] <- add.vertices(graphlist[[24]], 1, name="Jazz", label="Jazz", div="Midwest")
-graphlist[[24]] <- add.vertices(graphlist[[24]], 1, name="Knicks", label="Knicks", div="Atlantic")
-
-graphlist[[25]] <- add.vertices(graphlist[[25]], 1, name="Timberwolves", label="Timberwolves", div="Midwest")
-graphlist[[25]] <- add.vertices(graphlist[[25]], 1, name="Nets", label="Nets", div="Atlantic")
-
-graphlist[[26]] <- add.vertices(graphlist[[26]], 1, name="Jazz", label="Jazz", div="Midwest")
-
-graphlist[[27]] <- add.vertices(graphlist[[27]], 1, name="Heat", label="Heat", div="Atlantic")
-graphlist[[27]] <- add.vertices(graphlist[[27]], 1, name="Pacers", label="Pacers", div="Central")
-graphlist[[27]] <- add.vertices(graphlist[[27]], 1, name="Timberwolves", label="Timberwolves", div="Midwest")
-graphlist[[27]] <- add.vertices(graphlist[[27]], 1, name="Mavericks", label="Mavericks", div="Midwest")
-
-graphlist[[28]] <- add.vertices(graphlist[[28]], 1, name="Heat", label="Heat", div="Atlantic")
-graphlist[[28]] <- add.vertices(graphlist[[28]], 1, name="Lakers", label="Lakers", div="Pacific")
-
-graphlist[[29]] <- add.vertices(graphlist[[29]], 1, name="Timberwolves", label="Timberwolves", div="Northwest")
-
-graphlist[[30]] <- add.vertices(graphlist[[30]], 1, name="Mavericks", label="Mavericks", div="Southwest")
-graphlist[[30]] <- add.vertices(graphlist[[30]], 1, name="Spurs", label="Spurs", div="Southwest")
-graphlist[[30]] <- add.vertices(graphlist[[30]], 1, name="Warriors", label="Warriors", div="Pacific")
-
-graphlist[[31]] <- add.vertices(graphlist[[31]], 1, name="Heat", label="Heat", div="Southeast")
-graphlist[[31]] <- add.vertices(graphlist[[31]], 1, name="Clippers", label="Clippers", div="Pacific")
-graphlist[[31]] <- add.vertices(graphlist[[31]], 1, name="Kings", label="Kings", div="Pacific")
-graphlist[[31]] <- add.vertices(graphlist[[31]], 1, name="Wizards", label="Wizards", div="Southeast")
-graphlist[[31]] <- add.vertices(graphlist[[31]], 1, name="Knicks", label="Knicks", div="Atlantic")
-
-graphlist[[32]] <- add.vertices(graphlist[[32]], 1, name="Bucks", label="Bucks", div="Central")
-
-graphlist[[33]] <- add.vertices(graphlist[[33]], 1, name="Hawks", label="Hawks", div="Southeast")
-
-graphlist[[34]] <- add.vertices(graphlist[[34]], 1, name="Pacers", label="Pacers", div="Central")
-
-graphlist[[35]] <- add.vertices(graphlist[[35]], 1, name="Pistons", label="Pistons", div="Central")
-
-graphlist[[36]] <- add.vertices(graphlist[[36]], 1, name="Pistons", label="Pistons", div="Central")
-graphlist[[36]] <- add.vertices(graphlist[[36]], 1, name="Suns", label="Suns", div="Pacific")
-
-graphlist[[37]] <- add.vertices(graphlist[[37]], 1, name="Spurs", label="Spurs", div="Southwest")
-
-graphlist[[38]] <- add.vertices(graphlist[[38]], 1, name="Magic", label="Magic", div="Southeast")
-
-graphlist[[39]] <- add.vertices(graphlist[[39]], 1, name="Warriors", label="Warriors", div="Pacific")
-
-graphlist[[41]] <- add.vertices(graphlist[[41]], 1, name="Timberwolves", label="Timberwolves", div="Northwest")
-
-graphlist[[42]] <- add.vertices(graphlist[[42]], 1, name="Spurs", label="Spurs", div="Southwest")
-
-graphlist[[43]] <- add.vertices(graphlist[[43]], 1, name="Warriors", label="Warriors", div="Pacific")
-
-graphlist[[44]] <- add.vertices(graphlist[[44]], 1, name="Raptors", label="Raptors", div="Atlantic")
-```
-
-Save
-
-```{r graphlist-save}
-saveRDS(graphlist, file = "wgl.rds")
+saveRDS(graphlist, file = here::here("data", "igraph-list.rds"))
 ```
@@
 # Edge and node data frames
+
+Create node and edge data frames.
 
@@
 
-- `name` -- time-constant "franchise name"
-- `label` -- time-varying team name
-- `div` -- division name
+- `name` -- season-invariant "franchise name"
+- `team` -- team name
+- `division` -- division name
 - `season`, `season_from`, `season_to` -- season dates
@@
   )
+
 vdb
+
+saveRDS(vdb, file=here::here("data", "nodes.rds"))
 ```
@@
 
-- `from`, `to` -- franchise names
+- `from`, `to` -- trading franchises' names
 - `season`, `season_from`, `season_to` -- season dates
@@
   )
+
 edb
+
+saveRDS(edb, file = here::here("data", "edges.rds"))
 ```
@@
 
-## Save
-
-```{r save}
-saveRDS(vdb, file = here::here("data", "df-vdb.rds"))
-saveRDS(edb, file = here::here("data", "df-edb.rds"))
-```
-
 <!-- Include folded session info -->

### 2021-11-29 Manuscript.Rmd
Half-baked manuscript using Distill

@@
 title: "Do NBA Teams Avoid Trading within Their Own Division?^[We appreciate feedback we received on this paper from Skye Bender-deMoll, Pavel Krivitsky and David Schaefer.]"
-author: "jimi adams & Michał Bojanowski"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
-abstract: "Within US professional sports, trades within one's own division are often perceived to be disadvantageous.
+author: 
+  - first_name: Jimi
+    last_name: Adams
+    orcid_id: "0000-0001-5043-1149"
+    affiliation: University of Colorado
+  - first_name: Michał
+    last_name: Bojanowski
+    orcid_id: "0000-0001-7503-852X"
+    affiliation: Kozminski University
+date: November 29, 2021
+description: "Within US professional sports, trades within one's own division are often perceived to be disadvantageous.
 We ask how common this practice is.
@@
 output: 
-  bookdown::html_document2:
-    keep_md: true
+  distill::distill_article:
+    code_folding: true
 link-citations: true
@@
 requireNamespace("knitr")
+requireNamespace("igraph")
+requireNamespace("withr")
+library(tibble)
 library(dplyr)
+library(tidyr)
+library(purrr)
+library(ggplot2)
+
 knitr::opts_chunk$set(
@@
 
+```{r load-nodes-and-edges}
+vdb <- readRDS(here::here("data", "nodes.rds"))
+edb <- readRDS(here::here("data", "edges.rds"))
+```
+
 Here, we draw on a database of all player transactions in the NBA from the beginning of the 1976 season---when the NBA merged with the ABA uniting major professional basketball in the US under one league---through the completion of the 2018-19 season [@richardson2020].
@@
 
+<!--
 ![Figure 1. **Number of Trades by Year.** (normalized to a per team rate)](EDA_files/figure-gfm/trades-per-season-seasonwise-1.png)
+-->
+
+```{r trades-per-year, fig.cap="**Number of Trades by Year.** (normalized to a per team rate"}
+edb %>%
+  filter(season_to <= 2019) %>%
+  group_by(season) %>%
+  summarise(
+    n_trades = sum(weight),
+    .groups = "drop"
+  ) %>%
+  left_join(
+    count(vdb, season, name = "n_teams"),
+    by = "season"
+  ) %>%
+  transmute(
+    season,
+    trades_per_team = n_trades / n_teams
+  ) %>%
+  ggplot(aes(x=season, y = trades_per_team)) +
+  geom_bar(stat = "identity", alpha = 0.5) +
+  geom_hline(aes(yintercept = mean(trades_per_team)), color = "darkred") +
+  xlab("Season") +
+  ylab("Number of trades per team") +
+  theme(
+    axis.text.x = element_text(angle = 90, vjust = 0.5)
+  )
+```
 
@@
 
+<!--
 ![Figure 2. **Number of Trades by Team.** (normalized to a per season rate)](EDA_files/figure-gfm/trades-per-season-teamwise-1.png)
+-->
+
+```{r trades-per-team, fig.cap="**Number of Trades by Team.** (normalized to a per season rate)"}
+# Time-aggregated network
+g <- igraph::graph_from_data_frame(
+  edb %>%
+    filter(season_to <= 2019) %>%
+    group_by(from, to) %>%
+    summarise(
+      seasons = list(unique(season)),
+      weight = sum(weight),
+      .groups = "drop"
+    ),
+  directed = FALSE, 
+  vertices = vdb %>%
+    filter(season_to <= 2019) %>%
+    group_by(name) %>%
+    summarise(
+      seasons = list(unique(season)),
+      .groups = "drop"
+    )
+)
+
+igraph::strength(g) %>%
+  enframe(value = "n_trades") %>%
+  left_join(
+    vdb %>%
+      filter(season_to <= 2019) %>%
+      count(name, name = "n_seasons"),
+    by = "name"
+  ) %>%
+  mutate(
+    trades_per_season = n_trades / n_seasons
+  ) %>%
+  ggplot(aes(y=name, x=trades_per_season)) +
+  geom_vline(aes(xintercept = mean(trades_per_season)), color = "darkred") +
+  geom_col(alpha = 0.5) +
+  xlab("Number of trades per season") +
+  ylab("Team name")
+```
 
@@
 
-```{r, echo=FALSE}
+<!-- Manually wrap in distill HTML tags to get a proper caption. -->
+<div class="figure">
+```{r animation, echo=FALSE}
 htmltools::tags$iframe(src = 'NBA_Trades.html', width = '80%', height = '600px')
 ```
-
-Figure 3. **Dynamic Visualization of Annual Trade Networks** Note: Intra-division trades are highlighted in red. Node color and position refers to division.
+<p class="caption">Figure 3. **Dynamic Visualization of Annual Trade Networks** Note: Intra-division trades are highlighted in red. Node color and position refers to division.</p>
+</div>
 
@@
 ```{r mm-1976-2004}
-mm_1976_2004 <- readRDS("mm_1976_2004.rds")
+mm_1976_2004 <- edb %>%
+  filter(season_to <= 2004) %>%
+  left_join(
+    select(vdb, season, name, div_from = division),
+    by = c("season", "from" = "name")
+  ) %>%
+  left_join(
+    select(vdb, season, name, div_to = division),
+    by = c("season", "to" = "name")
+  ) %>%
+  mutate(
+    x = map2_chr(div_from, div_to, ~ paste(sort(c(.x, .y)), collapse="-"))
+  ) %>%
+  group_by(x) %>%
+  summarise(
+    n = sum(weight),
+    .groups = "drop"
+  ) %>%
+  separate(x, into=c("div_from", "div_to")) %>%
+  pivot_wider(names_from = div_to, values_from = n)
+
 withr::with_options(
@@
 ```{r mm-2005-2019}
-mm_2005_2019 <- readRDS("mm_2005_2019.rds")
+mm_2005_2019 <- edb %>%
+  filter(season_to >= 2005, season_to <= 2019) %>%
+  left_join(
+    select(vdb, season, name, div_from = division),
+    by = c("season", "from" = "name")
+  ) %>%
+  left_join(
+    select(vdb, season, name, div_to = division),
+    by = c("season", "to" = "name")
+  ) %>%
+  mutate(
+    x = map2_chr(div_from, div_to, ~ paste(sort(c(.x, .y)), collapse="-"))
+  ) %>%
+  group_by(x) %>%
+  summarise(
+    n = sum(weight),
+    .groups = "drop"
+  ) %>%
+  separate(x, into=c("div_from", "div_to")) %>%
+  pivot_wider(names_from = div_to, values_from = n)
+
 withr::with_options(
@@
 
+```{r load-graphlist}
+graphlist <- readRDS(here::here("data", "igraph-list.rds"))
+```
+
+<!--
 ![Figure 4. **Divisional Homophily in NBA Trades by Season.** (as computed by Moody's $\alpha$-index )](segregation_files/figure-gfm/or2-1.png)
+-->
+
+```{r alphas, eval=FALSE}
+nl <- lapply(graphlist, intergraph::asNetwork)
+
+alpha <- lapply(nl, OR2, "div") %>%
+  bind_rows(.id = "season") %>%
+  mutate(
+    season_start = as.numeric(substr(season, 1, 4))
+  ) %>%
+  filter(season_start < 2019)
+
+alpha %>%
+  ggplot(aes(x = season, y = or, ymin=ci.low, ymax=ci.high)) +
+  geom_hline(yintercept = 1, size = 0.5, color = "red") +
+  geom_pointrange() +
+  scale_y_continuous(trans = "log", breaks=c(1/8, 1/4, 1/2, 1, 2, 4)) +
+  xlab("Season") +
+  ylab("Within Division Trades (OR)") +
+  theme(
+    axis.text.x = element_text(angle = 90, vjust = 0.5)
+  )
+```
 
@@
 
+<!--
 ![Figure 5.
 **Homophily Coefficients and Standard Errors from Weighted Exponential Random Graph Models.** (including base-rate effects)](ergm-results_files/figure-gfm/B-nodematch-funnel-allinone-1.png)
+-->
+
+```{r seasonal-ergms, eval=FALSE, fig.cap="**Homophily Coefficients and Standard Errors from Weighted Exponential Random Graph Models.** (including base-rate effects)"}
+# TODO: include
+```
 
@@
 
-```{r model1-table-short}
+```{r model1-table-short, eval=FALSE}
  withr::with_options(
@@
 
-# Appendix
-
-## ERGM model fit
+## ERGM model fit {.appendix}
 
@@
 |*Homophily*           |  0.0099434| 0.0587577|(-0.11;  0.13) |   0.866|
-
-# References {-}

### 2021-11-30 codebox/sna.R
Add codebox with alpha()

@@ -1,0 +1,57 @@
+#' Calculate Moody's alpha with confidence intervals
+#'
+#' @param net object of class network
+#' @param att character; name of the vertex attribute
+#' @param conf.level numeric; value in [0,1], the confidence level
+#' 
+#' @return Numeric vector with elements:
+#' - `or` -- odds ratio for within-group ties
+#' - `se` -- standard error (square root of the sum of reciprocals of the b-g and w-g counts)
+#' - `ci.low`, `ci.high` -- confidence interval
+#' - `conf.level` - confidence level
+alpha <- function(net, att, conf.level = 0.95) {
+  box::use(
+    stats[
+      qnorm
+    ],
+    network[
+      set.vertex.attribute, 
+      get.vertex.attribute,
+      as.sociomatrix, 
+      network,
+      mixingmatrix,
+      is.directed
+    ]
+  )
+  stopifnot(inherits(net, "network"))
+  net.noties <- network(1 - as.sociomatrix(net))
+  set.vertex.attribute(
+    net.noties, 
+    att, 
+    get.vertex.attribute(net, att)
+  )
+  # Mixing matrices:
+  oT <- mixingmatrix(net, att)  # observed ties
+  nT <- mixingmatrix(net.noties, att)  # null ties
+  # Nos. of ties and noties between- and within-:
+  wg.ties <- sum(diag(oT))
+  bg.noties <- sum(nT) - sum(diag(nT))
+  wg.noties <- sum(diag(nT))
+  bg.ties <- sum(oT) - sum(diag(oT))
+  if(!is.directed(net)) {
+    # For undirected networks some of the counts need to be halved
+    bg.noties <- bg.noties / 2
+    bg.ties <- bg.ties / 2
+    wg.noties <- wg.noties / 2
+  }
+  OR <- (wg.ties / wg.noties) / (bg.ties / bg.noties)
+  SE <- sqrt(1 / wg.ties + 1 / bg.noties + 1 / wg.noties + 1 / bg.ties)
+  CI <- exp(log(OR) + qnorm(c((1 - conf.level)/2, 1 - (1 - conf.level)/2)) * SE)
+  names(CI) <- c("ci.low", "ci.high")
+  return(c(
+    or = OR, 
+    se = SE,
+    CI,
+    conf.level = conf.level
+  ))
+}

### 2021-11-30 Manuscript.Rmd
Re-add figure with alphas

@@
 ```{r setup, include=FALSE, cache=FALSE}
+requireNamespace("box")
 requireNamespace("knitr")
 requireNamespace("igraph")
+requireNamespace("intergraph")
 requireNamespace("withr")
@@
 
+# Functions from the 'codebox'
+box::use(
+  sna = codebox/sna[alpha]
+)
+
 knitr::opts_chunk$set(
@@
 )
+
+theme_set(theme_bw())
 ```
@@
 
-```{r load-nodes-and-edges}
+```{r load-data}
 vdb <- readRDS(here::here("data", "nodes.rds"))
 edb <- readRDS(here::here("data", "edges.rds"))
+graphlist <- readRDS(here::here("data", "igraph-list.rds"))
 ```
@@
 
-```{r load-graphlist}
-graphlist <- readRDS(here::here("data", "igraph-list.rds"))
-```
-
 <!--
@@
 
-```{r alphas, eval=FALSE}
-nl <- lapply(graphlist, intergraph::asNetwork)
+```{r alphas, fig.cap="**Divisional Homophily in NBA Trades by Season.** (as computed by Moody's $\alpha$-index)."}
+netlist <- lapply(graphlist, intergraph::asNetwork)
+names(netlist) <- vapply(graphlist, igraph::get.graph.attribute, character(1), "season")
 
-alpha <- lapply(nl, OR2, "div") %>%
+alpha <- lapply(netlist, alpha, "division") %>%
   bind_rows(.id = "season") %>%

### 2021-11-30 video.R
Add video-making script

@@ -1,0 +1,168 @@
+library(lubridate)
+library(networkDynamic)
+
+tl <- readRDS("data/trades.rds")
+tl$Date <- as.character(tl$Date)
+tl$P1 <- tl$franchise1
+tl$P2 <- tl$franchise2
+tl$franchise1 <- tl$franchise2 <- NULL
+
+# Removing trades from the incomplete 2019-20 season
+tl <- tl[which(as.Date(tl$Date) < as.Date("2019-06-01")),]
+
+# MB: This is already fixed in `data/trades.rds`
+# Correcting the misclassification of Hornets/Pelicans in ZR's recoding
+# tl$P1[which(tl$P1=="Pelicans" & as.Date(tl$Date) < as.Date("2002-05-01"))] <- "Hornets"
+# tl$P2[which(tl$P2=="Pelicans" & as.Date(tl$Date) < as.Date("2002-05-01"))] <- "Hornets"
+
+# Grabbing the "Franchise ID"
+fid <- unique(c(tl$P1, tl$P2))
+
+# Setting up the basic network structure
+  n <- network.initialize(length(fid), directed=F)
+  set.vertex.attribute(n, "franchise_name", fid)
+
+# Expansion Team 
+  activate.vertices(n, onset=1980, terminus=2019, v=which(get.vertex.attribute(n, "franchise_name")=="Mavericks"))
+  activate.vertices(n, onset=1988, terminus=2019, v=which(get.vertex.attribute(n, "franchise_name")=="Heat"))
+  activate.vertices(n, onset=1988, terminus=2019, v=which(get.vertex.attribute(n, "franchise_name")=="Hornets"))
+  activate.vertices(n, onset=1989, terminus=2019, v=which(get.vertex.attribute(n, "franchise_name")=="Timberwolves"))
+  activate.vertices(n, onset=1989, terminus=2019, v=which(get.vertex.attribute(n, "franchise_name")=="Magic"))
+  activate.vertices(n, onset=1995, terminus=2019, v=which(get.vertex.attribute(n, "franchise_name")=="Raptors"))
+  activate.vertices(n, onset=1995, terminus=2019, v=which(get.vertex.attribute(n, "franchise_name")=="Grizzlies"))
+  activate.vertices(n, onset=2002, terminus=2019, v=which(get.vertex.attribute(n, "franchise_name")=="Pelicans"))
+  
+# networkDynamic requires a numeric node ID, so quickly adding that
[…]

### 2021-12-01 video2.R
Simpler animation?

@@ -1,0 +1,25 @@
+requireNamespace("here")
+requireNamespace("igraph")
+library(networkDynamic)
+library(lubridate)
+
+glist <- readRDS(here::here("data", "igraph-list.rds"))
+netlist <- lapply(glist, intergraph::asNetwork)
+
+dn <- networkDynamic(network.list = netlist, vertex.pid = "vertex.names")
+activate.vertex.attribute(dn, )
+
+library(ndtv)
+xy <- compute.animation(dn)
+
+render.animation(xy)
+
+render.d3movie(
+  xy, 
+  file="video0.html",
+  tweenframes = 20,
+  d3.options = list(
+    enterExitAnimationFactor = 0.1
+  ),
+  displaylabels = TRUE
+)

### 2022-01-05 Manuscript.Rmd
Fix alpha in figure caption

@@
 
-```{r alphas, fig.cap="**Divisional Homophily in NBA Trades by Season.** (as computed by Moody's $\alpha$-index)."}
+```{r alphas, fig.cap="**Divisional Homophily in NBA Trades by Season.** (as computed by Moody's $\\alpha$-index)."}
 netlist <- lapply(graphlist, intergraph::asNetwork)

### 2022-01-05 data/standings.R
Make standings.R standalone
The script does not expect downloaded files, but runs wget to fetch
them.

@@ -12,10 +12,19 @@ years <- 1976:2019 # Season (year-1)-year
 urls <- glue("{base_url}/NBA_{years}.html")
 cat(urls, file=here::here("data", "fetchlist.txt"), sep="\n")
 
-# NOW execute `data/fetch.sh` to wget all the pages and save them to
-# `data/standings` folder. What follows assumes that they are there.
+# Use `wget` to fetch all the pages and save them to `data/standings` folder.
+system2(
+  "wget",
+  args = c(
+    paste0("-i ", here::here("data", "fetchlist.txt")),
+    "-w 6",
+    "--no-check-certificate",
+    "--user-agent=\"Mozilla/5.0 (X11; Linux x86_64; rv:60.0) Gecko/20100101 Firefox/60.0\"",
+    paste0("-P ", here::here("data", "standings"))
+  )
+)
 
-# List of downloaded files
+# Paths to downloaded files
 fnames <- list.files(
   here::here("data", "standings"),
   "^NBA_[0-9]{4}\\.html",

### 2022-01-05 ergm-pooled.R
Fix pooled ERGM estimation script
Parallel prociessing based on R_CORES envvar.
Use updated data file.
Vertex attribute is called 'division' not 'div'.

@@ -3,8 +3,14 @@ requireNamespace("igraph")
 library(ggplot2)
 library(ergm)
 library(ergm.count)
+box::use(
+  codebox/misc[get_r_cores]
+)
 
-load(here::here("data", "wgl.Rdata"))
+set.seed(666)
+r_cores <- get_r_cores()
+
+graphlist <- readRDS(here::here("data", "igraph-list.rds"))
 graphlist <- graphlist |>
   setNames(
     vapply(graphlist, igraph::graph_attr, character(1), name = "season")
@@ -24,7 +30,6 @@ season_from <- names(graphlist) |>
 bigg <- do.call(igraph::disjoint_union, graphlist)
 net <- intergraph::asNetwork(bigg)
 
-library(ergm.count)
 
 
 # Fit model 1 -------------------------------------------------------------
@@ -32,13 +37,13 @@ library(ergm.count)
 # All seasons/periods
 
 fit <- ergm(
-  net ~ nodefactor("season", levels = TRUE) + nodematch("div"),
+  net ~ nodefactor("season", levels = TRUE) + nodematch("division"),
   constraints = ~ blockdiag("season"),
   reference = ~ Poisson,
   response = "weight",
   verbose = TRUE,
   control = control.ergm(
-    parallel = 7
+    parallel = r_cores
   )
 )
 saveRDS(fit, file="ergm-pooled1.rds")
@@ -52,13 +57,13 @@ bigg2 <- do.call(igraph::disjoint_union, graphlist[season_from <= 2003])
 net2 <- intergraph::asNetwork(bigg2)
 
 fit2 <- ergm(
-  net2 ~ nodefactor("season", levels = TRUE) + nodematch("div"),
+  net2 ~ nodefactor("season", levels = TRUE) + nodematch("division"),
   constraints = ~ blockdiag("season"),
   reference = ~ Poisson,
   response = "weight",
   verbose = TRUE,
   control = control.ergm(
-    parallel = 7
+    parallel = r_cores
   )
 )
 
@@ -75,13 +80,13 @@ bigg3 <- do.call(igraph::disjoint_union, graphlist[season_from >= 2004])
 net3 <- intergraph::asNetwork(bigg3)
 
 fit3 <- ergm(
-  net3 ~ nodefactor("season", levels = TRUE) + nodematch("div"),
+  net3 ~ nodefactor("season", levels = TRUE) + nodematch("division"),
   constraints = ~ blockdiag("season"),
[…]

### 2022-01-05 codebox/misc.R
Add some misc funs

@@ -1,0 +1,13 @@
+#' Determine the number of cores to run on based on R_CORES
+#' 
+#' @param envar character; name of the variable to inspect
+#' 
+#' @return Integer number of cores. [control.ergm()] expects 0 for no parallel estimation.
+
+get_r_cores <- function(envar = "R_CORES") {
+  v <- Sys.getenv(envar)
+  if(v == "") return(0)
+  vn <- as.integer(v)
+  if(is.na(vn) || vn < 0 ) stop("Bad value of ", envar, ": ", v, call.=FALSE)
+  vn
+}

### 2022-01-05 codebox/gof.R
Functions

@@ -1,0 +1,185 @@
+#' GOF of valued ERGM: degree distribution
+#' 
+#' Calculate data necessary for degree distribution GOF plot of a valued ERGM.
+#' 
+#' @param fit fitted model object
+#' @param sims list of simulated networks from the `fit` (as returned by [ergm::simulate()])
+#' 
+#' @details Vertex degrees are calculated as the sum of weights.
+#' 
+#' @return A list with elements:
+#' - `observed` - tibble with columns `season`, `deg` (degree), and `n` (counts)
+#' - `simulated` - tibble with columns `sim`, `season`, `deg` and `n`
+#' 
+#' @note This implementation is not generic, i.e. it is tailored to NBA trades
+#' project.
+#' 
+#' @export
+gofplot_degrees <- function(fit, sims) {
+  box::use(
+    ergm[is.valued],
+    igraph[strength, degree],
+    intergraph[asIgraph],
+    network[`%v%`],
+    tibble[enframe],
+    dplyr[...],
+    stringi
+  )
+  # Argument checks
+  stopifnot(inherits(fit, "ergm"))
+  stopifnot(is.valued(fit))
+  stopifnot(inherits(sims, "network.list"))
+  
+  # Observed
+  obs <- fit$network |>
+    asIgraph() |>
+    strength() |>
+    tibble::enframe(value = "degree") |>
+    mutate(
+      vnames = fit$network %v% "vertex.names",
+      season = stringi::stri_extract_first_regex(vnames, "[0-9]{4}-[0-9]{4}"),
+      team = stringi::stri_extract_first_regex(vnames, "[0-9A-Za-z]+$")
+    ) |>
+    group_by(season, .drop = TRUE) |>
+    count(deg=degree) |>
+    ungroup()
+  
+  # Simulated
+  simed <- lapply(sims, function(n) {
+    strength(asIgraph(n))
+  }) |>
+    lapply(function(x) { tibble::tibble(deg = x, team = fit$network %v% "vertex.names")}) |>
+    dplyr::bind_rows(.id = "sim") |>
+    dplyr::mutate(
+      season = stringi::stri_extract_first_regex(team, "[0-9]{4}-[0-9]{4}"),
+      team = stringi::stri_extract_first_regex(team, "[0-9A-Za-z]+$")
+    ) |>
+    dplyr::group_by(sim, season, .drop = TRUE) |>
+    dplyr::count(deg) |>
+    ungroup()
+  
+  list(
+    observed = obs,
+    simulated = simed
+  )
+}
+
+
[…]

### 2022-01-05 codebox/misc.R

@@ -11,3 +11,84 @@ get_r_cores <- function(envar = "R_CORES") {
   if(is.na(vn) || vn < 0 ) stop("Bad value of ", envar, ": ", v, call.=FALSE)
   vn
 }
+
+
+
+
+
+#' Given an igraph object and vertex attribute create a list of subgraphs of
+#' within-group graphs
+#' 
+#' @export
+
+split_graph <- function(g, ...) UseMethod("split_graph")
+
+#' @export
+split_graph.igraph <- function(g, vattr) {
+  box::use(
+    igraph[vertex_attr, V, induced_subgraph]
+  )
+  v <- vertex_attr(g, vattr)
+  lapply(
+    sort(unique(v)),
+    function(x) {
+      induced_subgraph(g, v = which(v == x))
+    }
+  )
+}
+
+#' @export
+split_graph.network <- function(g, ...) {
+  box::use(
+    intergraph[asNetwork, asIgraph]
+  )
+  g |>
+    asIgraph() |>
+    split_graph.igraph(...) |>
+    lapply(asNetwork)
+}
+
+
+# -------------------------------------------------------------------------
+
+
+#' Converting arrays to tibbles with or without dimnames
+#' 
+#' @export
+asdf <- function(x, responseName="n", add_dimnames=FALSE,
+                 retval=c("tibble", "df"), ...) {
+  stopifnot(is.array(x))
+  retval <- match.arg(retval)
+  # construct df building call
+  ex <- quote(
+    data.frame(
+      do.call("expand.grid", c(
+        structure(
+          lapply(dim(x), seq),
+          names = paste0(".dim", seq(1, length(dim(x))))
+        ),
+        KEEP.OUT.ATTRS = FALSE,
+        stringsAsFactors=FALSE
+      ) ),
+      Freq = c(x) )
+  )
+  names(ex)[3L] <- responseName
+  rval <- eval(ex)
+  if(add_dimnames) {
+    ex <- quote(
+      data.frame(do.call("expand.grid", c(
+        dimnames(provideDimnames(x, ...)),
+        KEEP.OUT.ATTRS = FALSE,
+        stringsAsFactors=FALSE
+      ) ) )
+    )
+    dnames <- eval(ex)
+    rval <- cbind(rval, dnames)
+  }
+  
+  switch(retval,
+         tibble = tibble::tibble(rval),
+         df = rval
+  )
+}
+

### 2022-01-05 Manuscript.Rmd
Add table-making Rmd

@@
 
-```{r model1-table-short, eval=FALSE}
+```{r model1-table-short}
  withr::with_options(

### 2022-01-05 ergm-pooled-gof.Rmd
GOF of pooled ERGMs

---
title: "GoF of valued ERGMs fitted to pooled data"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, cache=FALSE}
library(ggplot2)
library(ergm.count)
requireNamespace("box")
box::use(
  funs.gof = codebox/gof,
  funs.utils = codebox/misc[...]
)

knitr::opts_chunk$set(
  echo = FALSE
  )

r_cores <- max(2, get_r_cores())
```

# Model 1

```{r model1-load}
fit <- readRDS("ergm-pooled1.rds")
```

Summary:

```{r model1-summary}
summary(fit)
```

```{r model1-gof-setup}
nsims <- 500
```

GOF is based on `r nsims` simulated networks.

```{r model1-simulate}
set.seed(666)
simnets <- simulate(fit, nsim = 500)
```

## Degree distribution

```{r model1-gof-degree, fig.width=7, fig.height=12}
dat <- funs.gof$gofplot_degrees(fit, simnets)

dat$simulated |>
  ggplot(aes(x=deg, y = n, group = deg)) +
  geom_boxplot(aes(color = "simulated"), key_glyph = draw_key_path) +
  geom_line(aes(x = deg, y = n, group = NULL, color = "observed"), 
            data = dat$observed, key_glyph = draw_key_path) +
  facet_wrap(~ season, ncol = 4) +
  xlab("Degree (weighted)") +
  ylab("Frequency") +
  scale_color_manual(
    name = "Simulated / Observed",
    values = c(simulated = "black", observed = "red"),
    labels = c("Simulated", "Observed")
  ) +
  theme(
    legend.position = "top"
  )
```

## Mixing matrix

```{r, eval=FALSE}
# Observed
obs <- funs.gof$nba_summary(fit$network ~ mm(div ~ div)) |>
  dplyr::mutate(
    div_pair = paste0(
      stringi::stri_extract_first_regex(name, "(?<=div=)[A-Za-z]+(?=,)") |>
        substr(1, 1),
      "-",
      stringi::stri_extract_first_regex(name, "(?<=div=)[A-Za-z]+(?=\\])") |>
        substr(1,1)
    )
  )

# Simulated

library(parallel)
cl <- makeCluster(r_cores)
clusterExport(cl, c("summary_formula", "funs.gof"))

l <- parLapply(
  cl, 
  simnets, 
  function(net) funs.gof$nba_summary(net ~ mm(div ~ div))
)

stopCluster(cl)

l |>
  dplyr::bind_rows(.id = "simid") |>
  dplyr::mutate(
    div_pair = paste0(
      stringi::stri_extract_first_regex(name, "(?<=div=)[A-Za-z]+(?=,)") |>
        substr(1, 1),
      "-",
      stringi::stri_extract_first_regex(name, "(?<=div=)[A-Za-z]+(?=\\])") |>
        substr(1,1)
    )
  ) -> d

d |>
  ggplot(aes(x=div_pair, y = value)) +
  geom_boxplot() +
  geom_point(aes(x=div_pair, y = value), color = "red", data = obs) +
  facet_wrap(~ season, scales = "free_x") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5)
  )

```

# Model 2

Seasons 1976-2004.

```{r model2-load}
fit <- readRDS("ergm-pooled2.rds")
```

Summary:

```{r model2-summary}
summary(fit)
```

```{r model2-gof-setup}
nsims <- 500
```

GOF is based on `r nsims` simulated networks.

```{r model2-simulate}
set.seed(666)
simnets <- simulate(fit, nsim = 500)
```

## Degree distribution

```{r model2-gof-degree, fig.width=7, fig.height=12}
dat <- funs.gof$gofplot_degrees(fit, simnets)

dat$simulated |>
  ggplot(aes(x=deg, y = n, group = deg)) +
  geom_boxplot(aes(color = "simulated"), key_glyph = draw_key_path) +
  geom_line(aes(x = deg, y = n, group = NULL, color = "observed"), 
            data = dat$observed, key_glyph = draw_key_path) +
  facet_wrap(~ season, ncol = 4) +
  xlab("Degree (weighted)") +
  ylab("Frequency") +
  scale_color_manual(
    name = "Simulated / Observed",
    values = c(simulated = "black", observed = "red"),
    labels = c("Simulated", "Observed")
  ) +
  theme(
    legend.position = "top"
  )
```

# Model 3

Seasons 2005-2019.

```{r model3-load}
fit <- readRDS("ergm-pooled3.rds")
```

Summary:

```{r model3-summary}
summary(fit)
```

```{r model3-gof-setup}
nsims <- 500
```

GOF is based on `r nsims` simulated networks.

```{r model3-simulate}
set.seed(666)
simnets <- simulate(fit, nsim = 500)
```

## Degree distribution

```{r model3-gof-degree, fig.width=7, fig.height=12}
dat <- funs.gof$gofplot_degrees(fit, simnets)

dat$simulated |>
  ggplot(aes(x=deg, y = n, group = deg)) +
  geom_boxplot(aes(color = "simulated"), key_glyph = draw_key_path) +
  geom_line(aes(x = deg, y = n, group = NULL, color = "observed"), 
            data = dat$observed, key_glyph = draw_key_path) +
  facet_wrap(~ season, ncol = 4) +
  xlab("Degree (weighted)") +
  ylab("Frequency") +
  scale_color_manual(
    name = "Simulated / Observed",
    values = c(simulated = "black", observed = "red"),
    labels = c("Simulated", "Observed")
  ) +
  theme(
    legend.position = "top"
  )
```

<!-- Include folded session info -->
<details>
<summary>Session info</summary>
```{r session-info}
sessioninfo::session_info()
```
</details>

### 2022-01-05 ergm-seasonal.R
Add seasonal ERGMs

@@ -1,0 +1,68 @@
+# Fitting weighted ERGMs to seasonal networks
+
+
+library(dplyr)
+library(purrr)
+library(rlang)
+library(ergm.count)
+library(parallel)
+library(doParallel)
+box::use(
+  codebox/misc[...]
+)
+
+r_cores <- max(2, get_r_cores())
+
+
+# Models ------------------------------------------------------------------
+
+models <- list(
+  A = net ~ sum + nodefactor("division"),
+  B = net ~ sum + nodefactor("division") + nodematch("division"),
+  C = net ~ sum + nodefactor("division") + nodematch("division", diff=TRUE),
+  D = net ~ sum + nodematch("division", diff=FALSE)
+) %>% 
+  tibble::enframe("model_name", "model_formula")
+
+
+# Data --------------------------------------------------------------------
+
+graphlist <- readRDS(here::here("data", "igraph-list.rds"))
+
+d <- tibble::tibble(
+  season = vapply(graphlist, igraph::get.graph.attribute, character(1), "season"),
+  graph = graphlist, # igraphs
+  net = lapply(graphlist, intergraph::asNetwork) # networks
+) %>%
+  tidyr::crossing(models) %>%
+  mutate(
+    model_formula = map2(net, model_formula, ~ {
+      environment(.y)$net <- .x
+      .y
+    }
+    )
+  )
+
+d$model_formula[[1]]
+ls(envir=environment(d$model_formula[[1]]))
+
+# Fit in parallel ---------------------------------------------------------
+
+cl <- makeCluster(r_cores)
+registerDoParallel(cl)
+do_fit <- function(i) {
+  net <- d$net[[i]]
+  frm <- d$model_formula[[i]]
+  environment(frm)$net <- net
+  fit <- try(ergm(frm, 
+       reference = ~ Poisson,
+       response = "weight"
+  ))
+  message("DONE ", i)
+  fit
+}
+clusterEvalQ(cl, library("ergm.count"))
+clusterExport(cl, list("do_fit", "d"))
+results <- parLapply(cl, seq(1, nrow(d)), do_fit)
+stopCluster(cl)
+saveRDS(results, file = here::here("ergm-seasonal.rds"))

### 2022-01-05 codebox/misc.R
Export get_r_cores()

@@ -2,7 +2,10 @@
 #' 
 #' @param envar character; name of the variable to inspect
 #' 
-#' @return Integer number of cores. [control.ergm()] expects 0 for no parallel estimation.
+#' @return Integer number of cores. [control.ergm()] expects 0 for no parallel
+#'   estimation.
+#' 
+#' @export
 
 get_r_cores <- function(envar = "R_CORES") {
   v <- Sys.getenv(envar)

### 2022-01-06 Manuscript.Rmd
Seasonal model results

@@
 requireNamespace("withr")
+requireNamespace("metafor")
 library(tibble)
@@
 
-```{r seasonal-ergms, eval=FALSE, fig.cap="**Homophily Coefficients and Standard Errors from Weighted Exponential Random Graph Models.** (including base-rate effects)"}
-# TODO: include
+```{r seasonal-ergms, fig.cap="**Homophily Coefficients and Standard Errors from Weighted Exponential Random Graph Models.** (including base-rate effects)"}
+with(
+  readRDS("ergm-seasonal-db.rds") %>%
+    mutate(
+      season_start = as.numeric(substr(season, 1, 4))
+    ) %>%
+    filter( 
+      model_name == "B", 
+      term == "nodematch.sum.division",
+      season_start < 2019
+    ),
+  metafor::funnel(estimate, vi = std.error^2, 
+                  col = c("black", "red")[(season_start > 2004) + 1]
+  )
+)
+legend(
+  "topright",
+  pch = 19,
+  col = c("black", "red"),
+  legend = c("1976 - 2004", "2004 - 2019")
+)
+
 ```

### 2022-01-06 ergm-seasonal-results.Rmd
Seasonal model results

---
title: "Valued ERGMs"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, cache=FALSE}
requireNamespace("metafor")
library(dplyr)
library(ergm.count)
knitr::opts_chunk$set(
  echo = FALSE
  )
```

```{r load-data}
graphlist <- readRDS(here::here("data", "igraph-list.rds"))
results <- readRDS("ergm-seasonal.rds")

models <- list(
  A = net ~ sum + nodefactor("division"),
  B = net ~ sum + nodefactor("division") + nodematch("division"),
  C = net ~ sum + nodefactor("division") + nodematch("division", diff=TRUE),
  D = net ~ sum + nodematch("division")
) %>% 
  tibble::enframe("model_name", "model_formula")

d <- tibble::tibble(
  season = vapply(graphlist, igraph::get.graph.attribute, character(1), "season"),
  graph = graphlist, # igraphs
  net = lapply(graphlist, intergraph::asNetwork) # networks
) %>%
  tidyr::crossing(models) %>%
  mutate(
    model_formula = purrr::map2(net, model_formula, ~ {
      environment(.y)$net <- .x
      .y
    }
    )
  )

op <- options(warn = -1)

db <- d %>%
  transmute(
    season, 
    season_start = as.numeric(substr(season, 1, 4)),
    model_name,
    results = lapply(results, broom::tidy)
  ) %>%
  tidyr::unnest(results)

options(op)

saveRDS(db, "ergm-seasonal-db.rds")
```

The models are:

```{r}
models %>%
  knitr::kable()
```

# All models

The specifications fit to each season separately. Each cell is `coefficient (standard error) p-value`.

```{r}
db %>%
  transmute(
    season,
    model_name,
    term,
    cell = glue::glue("{round(estimate, 3)} ({round(std.error, 3)}) {format.pval(p.value, 3)}")
  ) %>%
  tidyr::pivot_wider(
    one_of("season", "term"),
    names_from = "model_name", 
    values_from = "cell",
    values_fill = "-"
  ) %>%
  knitr::kable()
```

# Uniform homophily coefficients

I.e. `nodematch` terms from models B and D:

```{r B-D-nodematch}
db %>%
  filter(
    model_name %in% c("B", "D"),
    term == "nodematch.sum.division"
  ) %>%
  knitr::kable()

db %>%
  filter(model_name == "B", term == "nodematch.sum.division") %>%
  pull(estimate) %>%
  hist(main = "Values of nodematch from models B")
```

## Funnel plots

### Model B

```{r B-nodematch-funnel-1976-2004}
with(
  db %>%
    mutate(
      season_start = as.numeric(substr(season, 1, 4))
    ) %>%
    filter( 
      model_name == "B", 
      term == "nodematch.sum.division",
      season_start < 2004
    ),
  metafor::funnel(estimate, vi = std.error^2, main = "1976 - 2004")
)  
```

```{r B-nodematch-funnel-2004-2019}
with(
  db %>%
    mutate(
      season_start = as.numeric(substr(season, 1, 4))
    ) %>%
    filter( 
      model_name == "B", 
      term == "nodematch.sum.division",
      season_start >= 2004,
      season_start < 2019
    ),
  metafor::funnel(estimate, vi = std.error^2, main = "2004 - 2019")
)  
```

```{r B-nodematch-funnel-allinone}
with(
  db %>%
    mutate(
      season_start = as.numeric(substr(season, 1, 4))
    ) %>%
    filter( 
      model_name == "B", 
      term == "nodematch.sum.division",
      season_start < 2019
    ),
  metafor::funnel(estimate, vi = std.error^2, 
                  col = c("black", "red")[(season_start > 2004) + 1]
  )
)
legend(
  "topright",
  pch = 19,
  col = c("black", "red"),
  legend = c("1976 - 2004", "2004 - 2019")
)
```

### Model D

```{r D-nodematch-funnel-1976-2004}
with(
  db %>%
    mutate(
      season_start = as.numeric(substr(season, 1, 4))
    ) %>%
    filter( 
      model_name == "D", 
      term == "nodematch.sum.division",
      season_start < 2004
    ),
  metafor::funnel(estimate, vi = std.error^2, main = "1976 - 2004")
)  
```

```{r D-nodematch-funnel-2004-2019}
with(
  db %>%
    mutate(
      season_start = as.numeric(substr(season, 1, 4))
    ) %>%
    filter( 
      model_name == "D", 
      term == "nodematch.sum.division",
      season_start >= 2004,
      season_start < 2019
    ),
  metafor::funnel(estimate, vi = std.error^2, main = "2004 - 2019")
)  
```

```{r D-nodematch-funnel-allinone}
with(
  db %>%
    mutate(
      season_start = as.numeric(substr(season, 1, 4))
    ) %>%
    filter( 
      model_name == "D", 
      term == "nodematch.sum.division",
      season_start < 2019
    ),
  metafor::funnel(estimate, vi = std.error^2, 
                  col = c("black", "red")[(season_start > 2004) + 1]
  )
)
legend(
  "topright",
  pch = 19,
  col = c("black", "red"),
  legend = c("1976 - 2004", "2004 - 2019")
)
```

### 2022-01-06 Manuscript.Rmd
Fix year in funnel plot legend

@@
   col = c("black", "red"),
-  legend = c("1976 - 2004", "2004 - 2019")
+  legend = c("1976 - 2004", "2005 - 2019")
 )
-
 ```

### 2022-01-06 ergm-seasonal.R
Set seed for seasonal ERGMs

@@ -11,6 +11,7 @@ box::use(
   codebox/misc[...]
 )
 
+set.seed(6666)
 r_cores <- max(2, get_r_cores())
 
 

### 2022-01-06 Manuscript.Rmd
Generate table in the appendix

@@
   distill::distill_article:
-    code_folding: true
+    code_folding: false
 link-citations: true
@@
 
-Table: Full results for the ERG model fit to pooled data.
-
-|Effect                |   Estimate|        SE|95% CI         | p-value|
-|:---------------------|----------:|---------:|:--------------|-------:|
-|*Seasonal base rates* |           |          |               |        |
-|1976-1977             | -0.9026534| 0.0792530|(-1.06; -0.75) |   0.000|
-|1977-1978             | -0.8427925| 0.0724760|(-0.98; -0.70) |   0.000|
-|1978-1979             | -0.9263500| 0.0796106|(-1.08; -0.77) |   0.000|
-|1979-1980             | -0.8762352| 0.0771426|(-1.03; -0.73) |   0.000|
-|1980-1981             | -0.9910653| 0.0854415|(-1.16; -0.82) |   0.000|
-|1981-1982             | -0.9050884| 0.0754043|(-1.05; -0.76) |   0.000|
-|1982-1983             | -0.8069531| 0.0682278|(-0.94; -0.67) |   0.000|
-|1983-1984             | -0.9092292| 0.0766022|(-1.06; -0.76) |   0.000|
-|1984-1985             | -1.0526877| 0.0872369|(-1.22; -0.88) |   0.000|
-|1985-1986             | -1.0782139| 0.0930195|(-1.26; -0.90) |   0.000|
-|1986-1987             | -0.9198185| 0.0757693|(-1.07; -0.77) |   0.000|
-|1987-1988             | -0.9403157| 0.0781448|(-1.09; -0.79) |   0.000|
-|1988-1989             | -1.1192392| 0.0854069|(-1.29; -0.95) |   0.000|
-|1989-1990             | -1.1674130| 0.0852806|(-1.33; -1.00) |   0.000|
-|1990-1991             | -1.1015000| 0.0789942|(-1.26; -0.95) |   0.000|
-|1991-1992             | -1.3314738| 0.1014767|(-1.53; -1.13) |   0.000|
-|1992-1993             | -1.2131200| 0.0885597|(-1.39; -1.04) |   0.000|
-|1993-1994             | -1.2949074| 0.0953019|(-1.48; -1.11) |   0.000|
-|1994-1995             | -1.3952511| 0.1080338|(-1.61; -1.18) |   0.000|
-|1995-1996             | -1.2254082| 0.0852894|(-1.39; -1.06) |   0.000|
-|1996-1997             | -1.2243585| 0.0837553|(-1.39; -1.06) |   0.000|
-|1997-1998             | -1.1148373| 0.0727527|(-1.26; -0.97) |   0.000|
-|1998-1999             | -1.2492779| 0.0861211|(-1.42; -1.08) |   0.000|
-|1999-2000             | -1.2789135| 0.0904522|(-1.46; -1.10) |   0.000|
-|2000-2001             | -1.0430409| 0.0700097|(-1.18; -0.91) |   0.000|
-|2001-2002             | -1.0608807| 0.0696335|(-1.20; -0.92) |   0.000|
-|2002-2003             | -1.2653868| 0.0888219|(-1.44; -1.09) |   0.000|
-|2003-2004             | -1.0915922| 0.0733866|(-1.24; -0.95) |   0.000|
-|2004-2005             | -1.0507651| 0.0685301|(-1.19; -0.92) |   0.000|
-|2005-2006             | -1.0769347| 0.0703190|(-1.21; -0.94) |   0.000|
-|2006-2007             | -1.1925306| 0.0774736|(-1.34; -1.04) |   0.000|
-|2007-2008             | -1.1603188| 0.0759557|(-1.31; -1.01) |   0.000|
-|2008-2009             | -0.9697841| 0.0623303|(-1.09; -0.85) |   0.000|
-|2009-2010             | -0.9605862| 0.0613118|(-1.08; -0.84) |   0.000|
-|2010-2011             | -0.9335487| 0.0588953|(-1.05; -0.82) |   0.000|
-|2011-2012             | -1.1465080| 0.0721213|(-1.29; -1.01) |   0.000|
-|2012-2013             | -0.9791587| 0.0622563|(-1.10; -0.86) |   0.000|
-|2013-2014             | -0.9789426| 0.0643977|(-1.11; -0.85) |   0.000|
-|2014-2015             | -0.8316791| 0.0542273|(-0.94; -0.73) |   0.000|
-|2015-2016             | -1.0257114| 0.0647017|(-1.15; -0.90) |   0.000|
-|2016-2017             | -1.0875715| 0.0680398|(-1.22; -0.95) |   0.000|
-|2017-2018             | -1.0105469| 0.0640282|(-1.14; -0.89) |   0.000|
-|2018-2019             | -0.9172200| 0.0586872|(-1.03; -0.80) |   0.000|
-|*Homophily*           |  0.0099434| 0.0587577|(-0.11;  0.13) |   0.866|
+```{r}
+withr::with_options(
+  list(knitr.kable.NA = ""), {
+    knitr::kable(
+      readRDS(file = "model1-table.rds"), 
+      caption = "Full results for the ERG model fit to pooled data.",
+      align = "lrrrr")
+  })
+```

### 2022-01-06 Manuscript.Rmd
Make table/figure references dynamic
Closes #20.

@@
 Using these dates, there were no ambiguous trades that we were unable to attribute to a particular season.] The number of teams within each annual slice changes over time as the league gradually expanded from 22 to 30 teams.^[Unless otherwise specifically noted, all historical accounts in the text were pulled from Wikipedia and confirmed on nba.com.].
-The number of trades observed within each slice varies from year to year (range 22-86; see Figure 1), and teams exhibit differing rates of trading (see Figure 2).
+The number of trades observed within each slice varies from year to year (range 22-86; see Figure \@ref(fig:trades-per-year)), and teams exhibit differing rates of trading (see Figure \@ref(fig:trades-per-team)).
 
@@
 
-Figure 1 presents the trend of number of trades per season.
+Figure \@ref(fig:trades-per-year) presents the trend of number of trades per season.
 Since, the number of teams varies across the window, we standardize these values to represent the number of trades *per team* in each season.
@@
 Notably, this low-point came *after* the introduction of unrestricted free-agency (in 1988), allowing players to sign with any team after their contract expires.
-As a corollary, Figure 2 presents the frequency distribution of trades per team across the full observed window.
+As a corollary, Figure \@ref(fig:trades-per-team) presents the frequency distribution of trades per team across the full observed window.
 There are some teams (e.g., the Spurs) who consistently trade less frequently, and others who are much more trade active (e.g., the Mavericks).
@@
 Since this is a homophily statistic, testing for nodes within the *same* division, a significantly *negative* $\alpha$-index would be consistent with teams following the proposed prohibition.
-Figure 4 presents this $\alpha$-index for each trade season individually.
+Figure \@ref(fig:alphas) presents this $\alpha$-index for each trade season individually.
 What can be seen here is that in none of the observed trade years does the $\alpha$-index differ significantly from random expectation (OR=1).
@@
 Once properly conditioning the test for each of these necessary caveats, we find that in none of the seasons are teams more likely to avoid trading with other teams from within their division.
-As can be verified in the "funnel plot" presented in Figure 5, each the seasonal homophily coefficients fall within the white cone of statistical *in*significance.
+As can be verified in the "funnel plot" presented in Figure \@ref(fig:seasonal-ergms), each the seasonal homophily coefficients fall within the white cone of statistical *in*significance.
 
@@
 
-To summarize the results from Figure 5 in a single model, we also fit a pooled ERGM across all seasons, including a base-rate effect for each year's trade volume.
+To summarize the results from Figure \@ref(fig:seasonal-ergms) in a single model, we also fit a pooled ERGM across all seasons, including a base-rate effect for each year's trade volume.
 These results are presented in Table \@ref(tab:model1-table-short).
@@
      knitr::kable(
-       caption = "**Valued ERGM fitted to pooled seasonal data**. The table excludes seasonal constants, complete table is presented in the Appendix.",
+       caption = "**Valued ERGM fitted to pooled seasonal data**. The table excludes seasonal constants, complete table is presented in Table \\@ref(tab:model1-table) the Appendix.",
+
        align = "lrrrr"
@@
 
-```{r}
+```{r model1-table}
 withr::with_options(

### 2022-01-06 Manuscript.Rmd
Add table of contents and acknowledgements

@@
   distill::distill_article:
+    toc: true
     code_folding: false
@@
 ```
+
+## Acknowledgements {.appendix}
+
+We appreciate feedback we received on this paper from Skye Bender-deMoll, Pavel Krivitsky and David Schaefer.

### 2022-01-06 Manuscript.Rmd
Add missing parenthesis

@@
 
-```{r trades-per-year, fig.cap="**Number of Trades by Year.** (normalized to a per team rate"}
+```{r trades-per-year, fig.cap="**Number of Trades by Year.** (normalized to a per team rate)"}
 edb %>%

### 2022-01-06 Manuscript.Rmd
Add URL to GitHub

@@
 link-citations: true
+repository_url: https://github.com/mbojan/nba-trades
 bibliography: nba_trades.bib

### 2022-01-06 Manuscript.Rmd
Drop footnote to title

@@
 ---
-title: "Do NBA Teams Avoid Trading within Their Own Division?^[We appreciate feedback we received on this paper from Skye Bender-deMoll, Pavel Krivitsky and David Schaefer.]"
+title: "Do NBA Teams Avoid Trading within Their Own Division?"
 author: 

### 2022-01-06 Manuscript.Rmd
Add a note that it's a draft under review

@@
     affiliation: Kozminski University
-date: November 29, 2021
+date: "November 29, 2021"
 description: "Within US professional sports, trades within one's own division are often perceived to be disadvantageous.
@@
 In addition to the empirical question, this analysis serves to demonstrate the necessity and difficulty of constructing the proper baseline for statistical comparison.
-We find limited-to-no support for the popular perception."
+We find limited-to-no support for the popular perception.<br/><br/>**Draft under review**."
 output: 

### 2022-01-06 Manuscript.Rmd
Addresses

@@
     orcid_id: "0000-0001-5043-1149"
-    affiliation: University of Colorado
+    affiliation: University of Colorado, Denver, USA
   - first_name: Michał
@@
     orcid_id: "0000-0001-7503-852X"
-    affiliation: Kozminski University
+    affiliation: Kozminski University, Warsaw, Poland
 date: "November 29, 2021"

### 2022-05-10 Manuscript.Rmd
Back-implement in Rmd the changes jimi made to HTML

--- 2022-01-06 Manuscript.Rmd
@@
 author: 
-  - first_name: Jimi
-    last_name: Adams
+  - first_name: jimi
+    last_name: adams
     orcid_id: "0000-0001-5043-1149"
-    affiliation: University of Colorado
+    affiliation: University of Colorado Denver, USA
   - first_name: Michał
@@
     orcid_id: "0000-0001-7503-852X"
-    affiliation: Kozminski University
-date: November 29, 2021
+    affiliation: Kozminski University, Warsaw, Poland
+date: "November 29, 2021"
 description: "Within US professional sports, trades within one's own division are often perceived to be disadvantageous.
@@
     code_folding: false
+    keep_md: true
 link-citations: true
@@
 ```
-<p class="caption">Figure 3. **Dynamic Visualization of Annual Trade Networks** Note: Intra-division trades are highlighted in red. Node color and position refers to division.</p>
+<p class="caption">Video 1. **Dynamic Visualization of Annual Trade Networks** Note: Intra-division trades are highlighted in red. Node color and position refers to division.</p>
 </div>
@@
 The second window began in 2005 and runs through the end of our analytic period at the end of the 2018-19 season, during which the teams were separated into six divisions.
-To provide a visual representation of the changes in these division-specific trade patterns across the examined period, Figure 3 presents a dynamic visualization of each year's trade network.^[The appendices also include full replication code for all analyzes conducted and the generation of this visualization.]
+To provide a visual representation of the changes in these division-specific trade patterns across the examined period, Video 1 presents a dynamic visualization of each year's trade network.^[The appendices also include full replication code for all analyzes conducted and the generation of this visualization.]
 
@@
 
-It is worth mentioning that the models we have estimated in the results for Figure 5 are actually no different from what would have been possible through a generalized linear model predicting the (weighted) number of ties observed between pairs of teams.
+It is worth mentioning that the models we have estimated in the results for Figure \@ref(fig:seasonal-ergms) are actually no different from what would have been possible through a generalized linear model predicting the (weighted) number of ties observed between pairs of teams.
 However, the value of introducing the weighted ERGM approach with these data is that if subsequent researchers want to take the next step to provide an explanatory account that further examines which teams are likely to trade with which others, this modeling framework would allow for the simultaneous estimation of predictive factors that variously operate at the node level (e.g., is a team that finished lower in the previous year's standings more likely to trade), the dyad level (e.g., are teams with higher total salaries more likely to trade with other teams who have lower total salaries), or structural features above the dyad (e.g., the tendency to form trade-triples or larger configurations through multi-team deals).

### 2022-05-10 Manuscript.Rmd
Set the date to May 5 2022

@@
     affiliation: Kozminski University, Warsaw, Poland
-date: "November 29, 2021"
+date: "May 10, 2022"
 description: "Within US professional sports, trades within one's own division are often perceived to be disadvantageous.

### 2022-05-10 Manuscript.Rmd
Updated the text reference to Figure 1
Added range in units actually used in the Figure (annual per-team
numbero of trades).

@@
 Using these dates, there were no ambiguous trades that we were unable to attribute to a particular season.] The number of teams within each annual slice changes over time as the league gradually expanded from 22 to 30 teams.^[Unless otherwise specifically noted, all historical accounts in the text were pulled from Wikipedia and confirmed on nba.com.].
-The number of trades observed within each slice varies from year to year (range 22-86; see Figure \@ref(fig:trades-per-year)), and teams exhibit differing rates of trading (see Figure \@ref(fig:trades-per-team)).
+The number of trades observed within each slice varies from year to year (range 22-86 or 0.2-2.9 if expressed in per-team averages; see Figure \@ref(fig:trades-per-year)), and teams exhibit differing rates of trading (see Figure \@ref(fig:trades-per-team)).
 

### 2022-05-10 Manuscript.Rmd
Fix typo: it's 0.8, not 0.2

--- 2022-05-10 Manuscript.Rmd
@@
 Using these dates, there were no ambiguous trades that we were unable to attribute to a particular season.] The number of teams within each annual slice changes over time as the league gradually expanded from 22 to 30 teams.^[Unless otherwise specifically noted, all historical accounts in the text were pulled from Wikipedia and confirmed on nba.com.].
-The number of trades observed within each slice varies from year to year (range 22-86; see Figure \@ref(fig:trades-per-year)), and teams exhibit differing rates of trading (see Figure \@ref(fig:trades-per-team)).
+The number of trades observed within each slice varies from year to year (range 22-86 or 0.8-2.9 if expressed in per-team averages; see Figure \@ref(fig:trades-per-year)), and teams exhibit differing rates of trading (see Figure \@ref(fig:trades-per-team)).
 

### 2023-02-01 Manuscript.Rmd
Added pooled binary ERGM results

@@
 bibliography: nba_trades.bib
+editor_options: 
+  chunk_output_type: console
 ---
@@
 
-Beyond the popular perception of the avoidance of intra-division trades, there is also ample scholarly literature that would lead us to the same expectation, particularly in professional sports [@Stewart1999].
+Beyond the popular perception of the avoidance of intra-division trades, there is also ample scholarly literature that would lead us to the same expectation, particularly in professional sports [@Stewart1999], both arising from literature on organizational competition broadly, and management strategy more speficically.
+
+A primary basis for such an expectation arises from organizations avoiding collaborating with thier (perceived) competitors.
 For example, within a field where the actors are (or are perceived to be) competitors with one another, they may avoid cooperating with one another because it is perceived to be a competitive disadvantage [@Hoffmann2018], though that assumption has been questioned [@Bengtsson1999; @peng12].
 Instead of cooperating directly with one's competitors, actors may therefore seek out means of collaboration with those outside the competitive field [@Soda_2017].
-For example, in the case of NBA player development, this could lead to a variety of strategic patterns.
-Teams may prioritize trading with teams they do not perceive to be direct competitors [@Barman_2002], such as those outside their own division.
-Alternatively, general managers may develop strong stable relationships as resources [@Elfenbein2014], for example with particular player agents to provide comparative advantages in access to players on the free agent market (thus reducing their reliance on trades).
-Teams also could carve out recruitment niches that avoid direct competition [@Soltis_2010; @Barman_2002].
-For example, recent shifts in player "apprentice" opportunities through the developmental league and foreign partnerships has opened opportunities for teams to gain contractual rights to players outside of the trade system [@Keiper_2020].
-Each of these possibilities could bolster teams' ability to avoid cooperating through trading with their direct competitors.
+Or they may develop more narrow difinitions of who their competitors are [@Barman_2002].
+In combination, these could facilitate the avoidance of intra-divistion trades through prioritizing trades with teams who are not perceived as their direct competitors--i.e., from outside their own division.
 
-In sum, we therefore investigate how strongly division shapes trade partners among National Basketball Association (NBA) franchises, focused especially on whether teams tend to avoid trading with other teams within their own division.
+Additionally, the avoidance of competitor collaborations could be facilitated by the availability or pursuit of alternative strategies for developing competitive advantages.
+In this particular case, such alternatives could focus on strategies to enhanceplayer development (vs.
+player "acquisition" through trades).
+By shifting the focus to alternative strategies of player acquisition and/or development, teams coudl further avoid necessitating collaboration (through trades) within their division.
+This could take the form of enhancing other means of recruitment such as stable pipelines for talent acquisition [@brymer14; @Elfenbein2014], wherein general managers could develop stable relationships with particular agents to provide comparative advantages in access to players on the free agent market.
+Organizations also carve out recruitment niches that avoid direct competition [@Soltis_2010; @Barman_2002].
+For example, NBA teams have recently developed improved "apprentice" opportunities through the "Developmental League," while other teams have focused on foreign partnerships to open opportunities to gain contractual rights to players outside of the trade system [@Keiper_2020].
+While we do empirically test the development of these alternative opportunities here, each should facilitate teams' ability to avoid engaging in cooperating with their direct competitors--i.e., by avoiding intra-division trades in concert with the popular perception reported above.
+
+In sum, we therefore investigate how strongly division shapes trade partners among National Basketball Association (NBA) franchises.
+Drawing both on the popular refrain and supporting scholarly literature, this investigation begins with the expectatoin that teams may tend to avoid trading with other teams within their own division.
 
@@
 ```
-<p class="caption">Video 1. **Dynamic Visualization of Annual Trade Networks** Note: Intra-division trades are highlighted in red. Node color and position refers to division.</p>
+<p class="caption">Video 1.
+**Dynamic Visualization of Annual Trade Networks** Note: Intra-division trades are highlighted in red.
+Node color and position refers to division.
+The upper-right menu allows the animation speed to be adjusted while visualizing the entire sequence, or the foward/back buttons at the bottom can be used to animate the visualization one annual slice (trade season) at a time.</p>
 </div>
@@
 
+Adequate modeling a weighted network via an ERGM requires adapting model specification to the distribution of trade counts across dyads by choosing the reference measure [@krivit2012, sec.
+5.2].
+While the typical approach involves assuming that the counts are Poisson-distributed, in practice the data is distributed differently with overprevalence of 0s (zero-inflation) or with greater variation than the Poisson distribution assumes (overdispersion).
+Table \@ref(tab:dyadic-trade-counts) summarises that trading is relatively rare in general and it happens even more rearely that a pair of teams trades more than once in the same season.
+
+```{r dyadic-trade-counts}
+# MB: This is pasted from the calculation in privaterepo/overdispersion.Rmd
+structure(
+  list(
+    trades = c(0, 1, 2, 3, 4), 
+    n = c(13720L, 1683L,  118L, 11L, 1L), 
+    pct = c(88.3, 10.8, 0.8, 0.1, 0)
+  ), 
+  row.names = c(NA,  -5L), 
+  class = c("tbl_df", "tbl", "data.frame")
+) %>%
+  rename(
+    `Number of trades` = trades,
+    N = n,
+    `%` = pct
+  ) %>%
+  knitr::kable(
+    caption = "**Distribution of dyadic trade counts**. Frequency of number of trades in a season in pairs of teams.",    # Sounds convoluted...
+    align = "rrr"
+  )
+```
+
+At this point one could either choose a non-Poissonian reference measure or simplify data by dychotomizing the counts (any number of trades vs no trades at all) and proceeding with a binary ERGM.
+To make our conclusions more robust and for illustrative purposes with provide both types of results.
+
 Once properly conditioning the test for each of these necessary caveats, we find that in none of the seasons are teams more likely to avoid trading with other teams from within their division.
@@
 
-## ERGM model fit {.appendix}
+## Binary ERGM model fit {.appendix}
+
+```{r model-binary-table}
+structure(list(Effect = c("*Seasonal base rates*", "1977-1978",  "1978-1979", "1979-1980", "1980-1981", "1981-1982", "1982-1983",  "1983-1984", "1984-1985", "1985-1986", "1986-1987", "1987-1988",  "1988-1989", "1989-1990", "1990-1991", "1991-1992", "1992-1993",  "1993-1994", "1994-1995", "1995-1996", "1996-1997", "1997-1998",  "1998-1999", "1999-2000", "2000-2001", "2001-2002", "2002-2003",  "2003-2004", "2004-2005", "2005-2006", "2006-2007", "2007-2008",  "2008-2009", "2009-2010", "2010-2011", "2011-2012", "2012-2013",  "2013-2014", "2014-2015", "2015-2016", "2016-2017", "2017-2018",  "2018-2019", "Atlantic division", "Central division", "Midwest division",  "Northwest division", "Pacific division", "Southeast division",  "Southwest division", "*Homophily*"), Estimate = c(NA, 0.0644405709809631,  0.00191183013592781, 0.00191183013587444, -0.0510792301762593,  -0.00301475598725536, 0.11151751406511, 0.0273160881984102, -0.120873378851605,  -0.241797069713423, -0.034689299944754, -0.102736047536669, -0.217649092501239,  -0.30417868994459, -0.221517869773192, -0.423638103944277, -0.306766674453116,  -0.38207146557124, -0.492148757048376, -0.368437716751424, -0.351709490768931,  -0.21888283796225, -0.335437019204636, -0.38565088841675, -0.109253481582016,  -0.15570959635153, -0.351709490778105, -0.192860596693431, -0.187248236068849,  -0.22034655957727, -0.373121806940009, -0.33128353096099, -0.115852230243608,  -0.0873886627773311, -0.06903205140053, -0.304889898702513, -0.106234994930531,  -0.106234994930528, 0.0163922227340902, -0.18724823606885, -0.243383258589619,  -0.125604821725264, -0.0600276821843626, -0.801604197626109,  -0.866780460779051, -0.869801998775562, -0.71676022773136, -0.912537765260793,  -0.861071612165831, -0.665077455419437, 0.0122759266186398),      SE = c(NA, 0.127030461420487, 0.129814083679627, 0.129814092249876,      0.129291411629841, 0.12715500781249, 0.122756935416747, 0.125898331758024,      0.132729817906103, 0.139724669892161, 0.128542414167505,      0.131796473203285, 0.132010767728565, 0.131420860718578,      0.127137016042545, 0.138568763749447, 0.131454151198191,      0.135887273703477, 0.143380183694574, 0.130029601322226,      0.129122811845738, 0.122743181611443, 0.128263950828193,      0.130988252937235, 0.118460375222509, 0.120175658066723,      0.129122966384354, 0.12165157379233, 0.121780823545216, 0.122921411086859,      0.129126929340387, 0.127263407924174, 0.119539607693635,      0.118724498302347, 0.118221222501676, 0.126153907222097,      0.119259391114893, 0.119259390545633, 0.11609615644012, 0.121780823683373,      0.123755575999076, 0.119828869972253, 0.117980604919478,      0.0989169031985135, 0.0981917740131896, 0.100979355062305,      0.119436809946451, 0.0992411986882728, 0.12109213666063,      0.118908017386702, 0.0651683403015203), "95% CI" = c(NA,      "(-0.18;  0.31)", "(-0.25;  0.26)", "(-0.25;  0.26)", "(-0.30;  0.20)",      "(-0.25;  0.25)", "(-0.13;  0.35)", "(-0.22;  0.27)", "(-0.38;  0.14)",      "(-0.52;  0.03)", "(-0.29;  0.22)", "(-0.36;  0.16)", "(-0.48;  0.04)",      "(-0.56; -0.05)", "(-0.47;  0.03)", "(-0.70; -0.15)", "(-0.56; -0.05)",      "(-0.65; -0.12)", "(-0.77; -0.21)", "(-0.62; -0.11)", "(-0.60; -0.10)",      "(-0.46;  0.02)", "(-0.59; -0.08)", "(-0.64; -0.13)", "(-0.34;  0.12)",      "(-0.39;  0.08)", "(-0.60; -0.10)", "(-0.43;  0.05)", "(-0.43;  0.05)",      "(-0.46;  0.02)", "(-0.63; -0.12)", "(-0.58; -0.08)", "(-0.35;  0.12)",      "(-0.32;  0.15)", "(-0.30;  0.16)", "(-0.55; -0.06)", "(-0.34;  0.13)",      "(-0.34;  0.13)", "(-0.21;  0.24)", "(-0.43;  0.05)", "(-0.49;  0.00)",      "(-0.36;  0.11)", "(-0.29;  0.17)", "(-1.00; -0.61)", "(-1.06; -0.67)",      "(-1.07; -0.67)", "(-0.95; -0.48)", "(-1.11; -0.72)", "(-1.10; -0.62)",      "(-0.90; -0.43)", "(-0.12;  0.14)"), "p-value" = c(NA, 0.612,      0.988, 0.988, 0.693, 0.981, 0.364, 0.828, 0.362, 0.084, 0.787,      0.436, 0.099, 0.021, 0.081, 0.002, 0.02, 0.005, 0.001, 0.005,      0.006, 0.075, 0.009, 0.003, 0.356, 0.195, 0.006, 0.113, 0.124,      0.073, 0.004, 0.009, 0.332, 0.462, 0.559, 0.016, 0.373, 0.373,      0.888, 0.124, 0.049, 0.295, 0.611, 0, 0, 0, 0, 0, 0, 0, 0.851     )), class = c("tbl_df", "tbl", "data.frame"), row.names = c(NA,  -51L)) %>%
+  knitr::kable(
+    caption = "Full results for the binary ERG model fit to pooled data.",
+    align = "lrrrr"
+  )
+```
+
+## Weighted ERGM model fit {.appendix}
 

### 2023-02-01 Manuscript.Rmd
Transient manuscript updates

@@
   class = c("tbl_df", "tbl", "data.frame")
-) %>%
+) %>%(standard error of 0.059, z-statistic of 0.169)
   rename(
@@
 At this point one could either choose a non-Poissonian reference measure or simplify data by dychotomizing the counts (any number of trades vs no trades at all) and proceeding with a binary ERGM.
-To make our conclusions more robust and for illustrative purposes with provide both types of results.
+The homophily coefficient in the binary ERGM fit to all seasonal trade networks simultaneously, including season- and division-specific effects, is equal to 0.0122759 (standard error of 0.065, z-statistic of 0.188) showing no avoidance patterns.
+The complete model is presented in Table \@ref(model-binary-table) in the Appendix.
+
+A valued ERGM assuming Conway-Maxwell-Poisson distribution for the trade counts substantively leads to the same conclusion.
 
@@
 ```{r model-binary-table}
+# Model: privaterepo/estimation/ergm-pooled-binary-4
 structure(list(Effect = c("*Seasonal base rates*", "1977-1978",  "1978-1979", "1979-1980", "1980-1981", "1981-1982", "1982-1983",  "1983-1984", "1984-1985", "1985-1986", "1986-1987", "1987-1988",  "1988-1989", "1989-1990", "1990-1991", "1991-1992", "1992-1993",  "1993-1994", "1994-1995", "1995-1996", "1996-1997", "1997-1998",  "1998-1999", "1999-2000", "2000-2001", "2001-2002", "2002-2003",  "2003-2004", "2004-2005", "2005-2006", "2006-2007", "2007-2008",  "2008-2009", "2009-2010", "2010-2011", "2011-2012", "2012-2013",  "2013-2014", "2014-2015", "2015-2016", "2016-2017", "2017-2018",  "2018-2019", "Atlantic division", "Central division", "Midwest division",  "Northwest division", "Pacific division", "Southeast division",  "Southwest division", "*Homophily*"), Estimate = c(NA, 0.0644405709809631,  0.00191183013592781, 0.00191183013587444, -0.0510792301762593,  -0.00301475598725536, 0.11151751406511, 0.0273160881984102, -0.120873378851605,  -0.241797069713423, -0.034689299944754, -0.102736047536669, -0.217649092501239,  -0.30417868994459, -0.221517869773192, -0.423638103944277, -0.306766674453116,  -0.38207146557124, -0.492148757048376, -0.368437716751424, -0.351709490768931,  -0.21888283796225, -0.335437019204636, -0.38565088841675, -0.109253481582016,  -0.15570959635153, -0.351709490778105, -0.192860596693431, -0.187248236068849,  -0.22034655957727, -0.373121806940009, -0.33128353096099, -0.115852230243608,  -0.0873886627773311, -0.06903205140053, -0.304889898702513, -0.106234994930531,  -0.106234994930528, 0.0163922227340902, -0.18724823606885, -0.243383258589619,  -0.125604821725264, -0.0600276821843626, -0.801604197626109,  -0.866780460779051, -0.869801998775562, -0.71676022773136, -0.912537765260793,  -0.861071612165831, -0.665077455419437, 0.0122759266186398),      SE = c(NA, 0.127030461420487, 0.129814083679627, 0.129814092249876,      0.129291411629841, 0.12715500781249, 0.122756935416747, 0.125898331758024,      0.132729817906103, 0.139724669892161, 0.128542414167505,      0.131796473203285, 0.132010767728565, 0.131420860718578,      0.127137016042545, 0.138568763749447, 0.131454151198191,      0.135887273703477, 0.143380183694574, 0.130029601322226,      0.129122811845738, 0.122743181611443, 0.128263950828193,      0.130988252937235, 0.118460375222509, 0.120175658066723,      0.129122966384354, 0.12165157379233, 0.121780823545216, 0.122921411086859,      0.129126929340387, 0.127263407924174, 0.119539607693635,      0.118724498302347, 0.118221222501676, 0.126153907222097,      0.119259391114893, 0.119259390545633, 0.11609615644012, 0.121780823683373,      0.123755575999076, 0.119828869972253, 0.117980604919478,      0.0989169031985135, 0.0981917740131896, 0.100979355062305,      0.119436809946451, 0.0992411986882728, 0.12109213666063,      0.118908017386702, 0.0651683403015203), "95% CI" = c(NA,      "(-0.18;  0.31)", "(-0.25;  0.26)", "(-0.25;  0.26)", "(-0.30;  0.20)",      "(-0.25;  0.25)", "(-0.13;  0.35)", "(-0.22;  0.27)", "(-0.38;  0.14)",      "(-0.52;  0.03)", "(-0.29;  0.22)", "(-0.36;  0.16)", "(-0.48;  0.04)",      "(-0.56; -0.05)", "(-0.47;  0.03)", "(-0.70; -0.15)", "(-0.56; -0.05)",      "(-0.65; -0.12)", "(-0.77; -0.21)", "(-0.62; -0.11)", "(-0.60; -0.10)",      "(-0.46;  0.02)", "(-0.59; -0.08)", "(-0.64; -0.13)", "(-0.34;  0.12)",      "(-0.39;  0.08)", "(-0.60; -0.10)", "(-0.43;  0.05)", "(-0.43;  0.05)",      "(-0.46;  0.02)", "(-0.63; -0.12)", "(-0.58; -0.08)", "(-0.35;  0.12)",      "(-0.32;  0.15)", "(-0.30;  0.16)", "(-0.55; -0.06)", "(-0.34;  0.13)",      "(-0.34;  0.13)", "(-0.21;  0.24)", "(-0.43;  0.05)", "(-0.49;  0.00)",      "(-0.36;  0.11)", "(-0.29;  0.17)", "(-1.00; -0.61)", "(-1.06; -0.67)",      "(-1.07; -0.67)", "(-0.95; -0.48)", "(-1.11; -0.72)", "(-1.10; -0.62)",      "(-0.90; -0.43)", "(-0.12;  0.14)"), "p-value" = c(NA, 0.612,      0.988, 0.988, 0.693, 0.981, 0.364, 0.828, 0.362, 0.084, 0.787,      0.436, 0.099, 0.021, 0.081, 0.002, 0.02, 0.005, 0.001, 0.005,      0.006, 0.075, 0.009, 0.003, 0.356, 0.195, 0.006, 0.113, 0.124,      0.073, 0.004, 0.009, 0.332, 0.462, 0.559, 0.016, 0.373, 0.373,      0.888, 0.124, 0.049, 0.295, 0.611, 0, 0, 0, 0, 0, 0, 0, 0.851     )), class = c("tbl_df", "tbl", "data.frame"), row.names = c(NA,  -51L)) %>%

### 2023-02-01 Manuscript.Rmd
Manuscript updates:
Added binary and weighted CMP ERGM results.
Added table with distribution of trade counts.
Added text explaining the above.

--- 2023-02-01 Manuscript.Rmd
@@
 Table \@ref(tab:dyadic-trade-counts) summarises that trading is relatively rare in general and it happens even more rearely that a pair of teams trades more than once in the same season.
+Formal tests diagnoze overdispersion in this distribution [@cameron1990regression;@rpkg-AER].
 
@@
 At this point one could either choose a non-Poissonian reference measure or simplify data by dychotomizing the counts (any number of trades vs no trades at all) and proceeding with a binary ERGM.
-To make our conclusions more robust and for illustrative purposes with provide both types of results.
+The homophily coefficient in the binary ERGM fit to all seasonal trade networks simultaneously, including season- and division-specific effects, is equal to 0.0122759 (standard error of 0.065, z-statistic of 0.188) showing no avoidance patterns.
+The complete model is presented in Table \@ref(tab:model-binary-table) in the Appendix.
 
@@
 
-To summarize the results from Figure \@ref(fig:seasonal-ergms) in a single model, we also fit a pooled ERGM across all seasons, including a base-rate effect for each year's trade volume.
+To summarize the results from Figure \@ref(fig:seasonal-ergms) in a single model, we also fit a pooled ERGM across all seasons, including a base-rate effect for each year's trade volume, similarly to the binary ERGM mentioned above.
 These results are presented in Table \@ref(tab:model1-table-short).
@@
 
-We find an aggregate same-division homophily effect of 0.0099 (standard error of 0.059, z-statistic of 0.169), again showing no avoidance pattern.
+We find an aggregate same-division homophily effect of 0.0060264 (standard error of 0.059, z-statistic of 0.102), again showing no avoidance pattern.
 Fitting similar models to the periods 1976-2004 and 2004-2019 separately leads to almost identical results.[^periodModels]
@@
 ```{r model-binary-table}
+# Model: privaterepo/estimation/ergm-pooled-binary-4
+withr::with_options(
+  list(knitr.kable.NA = ""), {
     structure(list(Effect = c("*Seasonal base rates*", "1977-1978",  "1978-1979", "1979-1980", "1980-1981", "1981-1982", "1982-1983",  "1983-1984", "1984-1985", "1985-1986", "1986-1987", "1987-1988",  "1988-1989", "1989-1990", "1990-1991", "1991-1992", "1992-1993",  "1993-1994", "1994-1995", "1995-1996", "1996-1997", "1997-1998",  "1998-1999", "1999-2000", "2000-2001", "2001-2002", "2002-2003",  "2003-2004", "2004-2005", "2005-2006", "2006-2007", "2007-2008",  "2008-2009", "2009-2010", "2010-2011", "2011-2012", "2012-2013",  "2013-2014", "2014-2015", "2015-2016", "2016-2017", "2017-2018",  "2018-2019", "Atlantic division", "Central division", "Midwest division",  "Northwest division", "Pacific division", "Southeast division",  "Southwest division", "*Homophily*"), Estimate = c(NA, 0.0644405709809631,  0.00191183013592781, 0.00191183013587444, -0.0510792301762593,  -0.00301475598725536, 0.11151751406511, 0.0273160881984102, -0.120873378851605,  -0.241797069713423, -0.034689299944754, -0.102736047536669, -0.217649092501239,  -0.30417868994459, -0.221517869773192, -0.423638103944277, -0.306766674453116,  -0.38207146557124, -0.492148757048376, -0.368437716751424, -0.351709490768931,  -0.21888283796225, -0.335437019204636, -0.38565088841675, -0.109253481582016,  -0.15570959635153, -0.351709490778105, -0.192860596693431, -0.187248236068849,  -0.22034655957727, -0.373121806940009, -0.33128353096099, -0.115852230243608,  -0.0873886627773311, -0.06903205140053, -0.304889898702513, -0.106234994930531,  -0.106234994930528, 0.0163922227340902, -0.18724823606885, -0.243383258589619,  -0.125604821725264, -0.0600276821843626, -0.801604197626109,  -0.866780460779051, -0.869801998775562, -0.71676022773136, -0.912537765260793,  -0.861071612165831, -0.665077455419437, 0.0122759266186398),      SE = c(NA, 0.127030461420487, 0.129814083679627, 0.129814092249876,      0.129291411629841, 0.12715500781249, 0.122756935416747, 0.125898331758024,      0.132729817906103, 0.139724669892161, 0.128542414167505,      0.131796473203285, 0.132010767728565, 0.131420860718578,      0.127137016042545, 0.138568763749447, 0.131454151198191,      0.135887273703477, 0.143380183694574, 0.130029601322226,      0.129122811845738, 0.122743181611443, 0.128263950828193,      0.130988252937235, 0.118460375222509, 0.120175658066723,      0.129122966384354, 0.12165157379233, 0.121780823545216, 0.122921411086859,      0.129126929340387, 0.127263407924174, 0.119539607693635,      0.118724498302347, 0.118221222501676, 0.126153907222097,      0.119259391114893, 0.119259390545633, 0.11609615644012, 0.121780823683373,      0.123755575999076, 0.119828869972253, 0.117980604919478,      0.0989169031985135, 0.0981917740131896, 0.100979355062305,      0.119436809946451, 0.0992411986882728, 0.12109213666063,      0.118908017386702, 0.0651683403015203), "95% CI" = c(NA,      "(-0.18;  0.31)", "(-0.25;  0.26)", "(-0.25;  0.26)", "(-0.30;  0.20)",      "(-0.25;  0.25)", "(-0.13;  0.35)", "(-0.22;  0.27)", "(-0.38;  0.14)",      "(-0.52;  0.03)", "(-0.29;  0.22)", "(-0.36;  0.16)", "(-0.48;  0.04)",      "(-0.56; -0.05)", "(-0.47;  0.03)", "(-0.70; -0.15)", "(-0.56; -0.05)",      "(-0.65; -0.12)", "(-0.77; -0.21)", "(-0.62; -0.11)", "(-0.60; -0.10)",      "(-0.46;  0.02)", "(-0.59; -0.08)", "(-0.64; -0.13)", "(-0.34;  0.12)",      "(-0.39;  0.08)", "(-0.60; -0.10)", "(-0.43;  0.05)", "(-0.43;  0.05)",      "(-0.46;  0.02)", "(-0.63; -0.12)", "(-0.58; -0.08)", "(-0.35;  0.12)",      "(-0.32;  0.15)", "(-0.30;  0.16)", "(-0.55; -0.06)", "(-0.34;  0.13)",      "(-0.34;  0.13)", "(-0.21;  0.24)", "(-0.43;  0.05)", "(-0.49;  0.00)",      "(-0.36;  0.11)", "(-0.29;  0.17)", "(-1.00; -0.61)", "(-1.06; -0.67)",      "(-1.07; -0.67)", "(-0.95; -0.48)", "(-1.11; -0.72)", "(-1.10; -0.62)",      "(-0.90; -0.43)", "(-0.12;  0.14)"), "p-value" = c(NA, 0.612,      0.988, 0.988, 0.693, 0.981, 0.364, 0.828, 0.362, 0.084, 0.787,      0.436, 0.099, 0.021, 0.081, 0.002, 0.02, 0.005, 0.001, 0.005,      0.006, 0.075, 0.009, 0.003, 0.356, 0.195, 0.006, 0.113, 0.124,      0.073, 0.004, 0.009, 0.332, 0.462, 0.559, 0.016, 0.373, 0.373,      0.888, 0.124, 0.049, 0.295, 0.611, 0, 0, 0, 0, 0, 0, 0, 0.851     )), class = c("tbl_df", "tbl", "data.frame"), row.names = c(NA,  -51L)) %>%
@@
         align = "lrrrr"
+      )
+  }
 )
@@
   list(knitr.kable.NA = ""), {
+    structure(list(Effect = c("CMP", "1976-1977", "1977-1978", "1978-1979",  "1979-1980", "1980-1981", "1981-1982", "1982-1983", "1983-1984",  "1984-1985", "1985-1986", "1986-1987", "1987-1988", "1988-1989",  "1989-1990", "1990-1991", "1991-1992", "1992-1993", "1993-1994",  "1994-1995", "1995-1996", "1996-1997", "1997-1998", "1998-1999",  "1999-2000", "2000-2001", "2001-2002", "2002-2003", "2003-2004",  "2004-2005", "2005-2006", "2006-2007", "2007-2008", "2008-2009",  "2009-2010", "2010-2011", "2011-2012", "2012-2013", "2013-2014",  "2014-2015", "2015-2016", "2016-2017", "2017-2018", "2018-2019",  "*Homophily*"), Estimate = c(0.231584299590497, -0.890763342957112,  -0.83536651253216, -0.916362759460457, -0.871853243700118, -0.987841290853209,  -0.900902605442105, -0.800007214754898, -0.899560461864575, -1.0483260921955,  -1.07751539263132, -0.913653500983868, -0.937433706117321, -1.11513540560325,  -1.16287289821496, -1.09836829500166, -1.32117548878016, -1.20705371653326,  -1.28771790721094, -1.39250323356208, -1.22250163107761, -1.22000987502278,  -1.11156405118103, -1.25007055148395, -1.27514745462814, -1.03847047606852,  -1.05882132141435, -1.26269742772427, -1.08833699875934, -1.0469834704777,  -1.07289680849128, -1.1929728670501, -1.1562207254889, -0.961040860349278,  -0.956076839003664, -0.925824485618014, -1.14194414800718, -0.971701230551129,  -0.971414447588866, -0.82982643075161, -1.01829169839617, -1.07774381829707,  -0.999993419528592, -0.912020178092013, 0.00602638250203989),      SE = c(0.12203512128938, 0.0786659348166571, 0.0749602274310118,      0.0789621568846716, 0.0760022797170869, 0.0849915193088094,      0.0765183664206582, 0.069324091908689, 0.0754443838436976,      0.0885417986944042, 0.0913748344030373, 0.0767335950520064,      0.081705154479656, 0.0874539689668034, 0.0838134623779419,      0.0794202098916715, 0.103689830592027, 0.0889138426500533,      0.0980046218251252, 0.105920044602825, 0.0836029172158815,      0.080937063964129, 0.0737605330112484, 0.0872142023820389,      0.0904111796177795, 0.0712261498735629, 0.0719333510361581,      0.0885909751979331, 0.0733621462658981, 0.067206171621419,      0.0674913276175189, 0.0767037480828984, 0.0768718796054209,      0.0627954612107514, 0.06156685684603, 0.0602325458811699,      0.0751034750003797, 0.0636099818174729, 0.0642021297111841,      0.0536773940981808, 0.0669347619251306, 0.0717645838325605,      0.0660829455575422, 0.0579577186366169, 0.0593486673413606     ), "95% CI" = c("(-0.01;  0.47)", "(-1.04; -0.74)", "(-0.98; -0.69)",      "(-1.07; -0.76)", "(-1.02; -0.72)", "(-1.15; -0.82)", "(-1.05; -0.75)",      "(-0.94; -0.66)", "(-1.05; -0.75)", "(-1.22; -0.87)", "(-1.26; -0.90)",      "(-1.06; -0.76)", "(-1.10; -0.78)", "(-1.29; -0.94)", "(-1.33; -1.00)",      "(-1.25; -0.94)", "(-1.52; -1.12)", "(-1.38; -1.03)", "(-1.48; -1.10)",      "(-1.60; -1.18)", "(-1.39; -1.06)", "(-1.38; -1.06)", "(-1.26; -0.97)",      "(-1.42; -1.08)", "(-1.45; -1.10)", "(-1.18; -0.90)", "(-1.20; -0.92)",      "(-1.44; -1.09)", "(-1.23; -0.94)", "(-1.18; -0.92)", "(-1.21; -0.94)",      "(-1.34; -1.04)", "(-1.31; -1.01)", "(-1.08; -0.84)", "(-1.08; -0.84)",      "(-1.04; -0.81)", "(-1.29; -0.99)", "(-1.10; -0.85)", "(-1.10; -0.85)",      "(-0.94; -0.72)", "(-1.15; -0.89)", "(-1.22; -0.94)", "(-1.13; -0.87)",      "(-1.03; -0.80)", "(-0.11;  0.12)"), "p-value" = c(0.058,      0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,      0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,      0, 0, 0, 0, 0, 0.919)), row.names = c(NA, -45L), class = "data.frame") |>
     knitr::kable(
-      readRDS(file = "model1-table.rds"), 
-      caption = "Full results for the ERG model fit to pooled data.",
+      caption = "Full results for the weighted ERG model (Conway-Maxwell-Poisson reference) fit to pooled data.",
       align = "lrrrr")

### 2023-02-01 Manuscript.Rmd
Added a comment

--- 2023-02-01 Manuscript.Rmd
@@
 ```{r model1-table}
+# The model from privaterepo/estimation/ergm-pooled-cmp-1
 withr::with_options(
   list(knitr.kable.NA = ""), {
+    structure(list(Effect = c("CMP", "1976-1977", "1977-1978", "1978-1979",  "1979-1980", "1980-1981", "1981-1982", "1982-1983", "1983-1984",  "1984-1985", "1985-1986", "1986-1987", "1987-1988", "1988-1989",  "1989-1990", "1990-1991", "1991-1992", "1992-1993", "1993-1994",  "1994-1995", "1995-1996", "1996-1997", "1997-1998", "1998-1999",  "1999-2000", "2000-2001", "2001-2002", "2002-2003", "2003-2004",  "2004-2005", "2005-2006", "2006-2007", "2007-2008", "2008-2009",  "2009-2010", "2010-2011", "2011-2012", "2012-2013", "2013-2014",  "2014-2015", "2015-2016", "2016-2017", "2017-2018", "2018-2019",  "*Homophily*"), Estimate = c(0.231584299590497, -0.890763342957112,  -0.83536651253216, -0.916362759460457, -0.871853243700118, -0.987841290853209,  -0.900902605442105, -0.800007214754898, -0.899560461864575, -1.0483260921955,  -1.07751539263132, -0.913653500983868, -0.937433706117321, -1.11513540560325,  -1.16287289821496, -1.09836829500166, -1.32117548878016, -1.20705371653326,  -1.28771790721094, -1.39250323356208, -1.22250163107761, -1.22000987502278,  -1.11156405118103, -1.25007055148395, -1.27514745462814, -1.03847047606852,  -1.05882132141435, -1.26269742772427, -1.08833699875934, -1.0469834704777,  -1.07289680849128, -1.1929728670501, -1.1562207254889, -0.961040860349278,  -0.956076839003664, -0.925824485618014, -1.14194414800718, -0.971701230551129,  -0.971414447588866, -0.82982643075161, -1.01829169839617, -1.07774381829707,  -0.999993419528592, -0.912020178092013, 0.00602638250203989),      SE = c(0.12203512128938, 0.0786659348166571, 0.0749602274310118,      0.0789621568846716, 0.0760022797170869, 0.0849915193088094,      0.0765183664206582, 0.069324091908689, 0.0754443838436976,      0.0885417986944042, 0.0913748344030373, 0.0767335950520064,      0.081705154479656, 0.0874539689668034, 0.0838134623779419,      0.0794202098916715, 0.103689830592027, 0.0889138426500533,      0.0980046218251252, 0.105920044602825, 0.0836029172158815,      0.080937063964129, 0.0737605330112484, 0.0872142023820389,      0.0904111796177795, 0.0712261498735629, 0.0719333510361581,      0.0885909751979331, 0.0733621462658981, 0.067206171621419,      0.0674913276175189, 0.0767037480828984, 0.0768718796054209,      0.0627954612107514, 0.06156685684603, 0.0602325458811699,      0.0751034750003797, 0.0636099818174729, 0.0642021297111841,      0.0536773940981808, 0.0669347619251306, 0.0717645838325605,      0.0660829455575422, 0.0579577186366169, 0.0593486673413606     ), "95% CI" = c("(-0.01;  0.47)", "(-1.04; -0.74)", "(-0.98; -0.69)",      "(-1.07; -0.76)", "(-1.02; -0.72)", "(-1.15; -0.82)", "(-1.05; -0.75)",      "(-0.94; -0.66)", "(-1.05; -0.75)", "(-1.22; -0.87)", "(-1.26; -0.90)",      "(-1.06; -0.76)", "(-1.10; -0.78)", "(-1.29; -0.94)", "(-1.33; -1.00)",      "(-1.25; -0.94)", "(-1.52; -1.12)", "(-1.38; -1.03)", "(-1.48; -1.10)",      "(-1.60; -1.18)", "(-1.39; -1.06)", "(-1.38; -1.06)", "(-1.26; -0.97)",      "(-1.42; -1.08)", "(-1.45; -1.10)", "(-1.18; -0.90)", "(-1.20; -0.92)",      "(-1.44; -1.09)", "(-1.23; -0.94)", "(-1.18; -0.92)", "(-1.21; -0.94)",      "(-1.34; -1.04)", "(-1.31; -1.01)", "(-1.08; -0.84)", "(-1.08; -0.84)",      "(-1.04; -0.81)", "(-1.29; -0.99)", "(-1.10; -0.85)", "(-1.10; -0.85)",      "(-0.94; -0.72)", "(-1.15; -0.89)", "(-1.22; -0.94)", "(-1.13; -0.87)",      "(-1.03; -0.80)", "(-0.11;  0.12)"), "p-value" = c(0.058,      0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,      0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,      0, 0, 0, 0, 0, 0.919)), row.names = c(NA, -45L), class = "data.frame") |>
     knitr::kable(
-      readRDS(file = "model1-table.rds"), 
-      caption = "Full results for the ERG model fit to pooled data.",
+      caption = "Full results for the weighted ERG model (Conway-Maxwell-Poisson reference) fit to pooled data.",
       align = "lrrrr")

### 2023-02-01 Manuscript.Rmd
revisions for clarity

@@
 A common refrain among sports commentators is that professional sports teams avoid trading players with other teams from within their own division [@wong2017; @simmons].
-Some offered conjectures on why this should be avoided include: not wanting to improve the competitiveness of a team's direct rivals [@ley2017], especially in leagues where games within the division are more common than other match-ups [@bates2015], or wanting to avoid fans being reminded of players they gave up [@fs2015], especially if they turn out to play better for their new teams [@ley2017].
+Some conjectures on why this should be avoided include: not wanting to improve the competitiveness of a team's direct rivals [@ley2017], especially in leagues where games within the division are more common than other match-ups [@bates2015], or wanting to avoid fans being reminded of players they gave up [@fs2015], especially if they turn out to play better for their new teams [@ley2017].
 Despite the frequency of this speculation, attempts to quantify the actual frequency of such a prohibition has been rare [for one limited exception, see @ahr2018].
 
-Beyond the popular perception of the avoidance of intra-division trades, there is also ample scholarly literature that would lead us to the same expectation, particularly in professional sports [@Stewart1999], both arising from literature on organizational competition broadly, and management strategy more speficically.
+Beyond the popular perception of the avoidance of intra-division trades, there is also ample scholarly literature that would lead us to the same expectation, particularly in professional sports [@Stewart1999].
+These expectations have theoretical roots in literature on organizational competition broadly, and management strategy more specifically.
 
@@
 Or they may develop more narrow difinitions of who their competitors are [@Barman_2002].
-In combination, these could facilitate the avoidance of intra-divistion trades through prioritizing trades with teams who are not perceived as their direct competitors--i.e., from outside their own division.
+In combination, these could facilitate the avoidance of intra-division trades through prioritizing trades with teams who are not perceived as their direct competitors--i.e., from outside their own division.
 
 Additionally, the avoidance of competitor collaborations could be facilitated by the availability or pursuit of alternative strategies for developing competitive advantages.
-In this particular case, such alternatives could focus on strategies to enhanceplayer development (vs.
+In this particular case, such alternatives could focus on strategies to enhance player development (vs.
 player "acquisition" through trades).
-By shifting the focus to alternative strategies of player acquisition and/or development, teams coudl further avoid necessitating collaboration (through trades) within their division.
-This could take the form of enhancing other means of recruitment such as stable pipelines for talent acquisition [@brymer14; @Elfenbein2014], wherein general managers could develop stable relationships with particular agents to provide comparative advantages in access to players on the free agent market.
+This could take the form of enhancing other means of recruitment such as stable pipelines for talent acquisition [@brymer14; @Elfenbein2014], wherein general managers could develop stable relationships with particular representatives to provide comparative advantages in access to players on the free agent market.
+By shifting the focus to alternative strategies of player acquisition and/or development in these ways, teams could further avoid necessitating collaboration (through trades) within their division.
+
 Organizations also carve out recruitment niches that avoid direct competition [@Soltis_2010; @Barman_2002].
-For example, NBA teams have recently developed improved "apprentice" opportunities through the "Developmental League," while other teams have focused on foreign partnerships to open opportunities to gain contractual rights to players outside of the trade system [@Keiper_2020].
-While we do empirically test the development of these alternative opportunities here, each should facilitate teams' ability to avoid engaging in cooperating with their direct competitors--i.e., by avoiding intra-division trades in concert with the popular perception reported above.
+For example, NBA teams have recently developed improved "apprentice" opportunities through the "Developmental League," while some teams have focused on foreign partnerships to open opportunities to gain contractual rights to players outside of the trade system [@Keiper_2020].
+While we do not empirically test the development of these alternative mechanisms here, each should facilitate teams' ability to avoid engaging in cooperating with their direct competitors--i.e., by avoiding intra-division trades--directly paralleling the expectation from the popular perception as reported above.
 
 In sum, we therefore investigate how strongly division shapes trade partners among National Basketball Association (NBA) franchises.
-Drawing both on the popular refrain and supporting scholarly literature, this investigation begins with the expectatoin that teams may tend to avoid trading with other teams within their own division.
+Drawing both on the popular refrain and supporting scholarly literature, this investigation begins with the expectation that teams may tend to avoid trading with other teams within their own division.
 
@@
 
-Here, we draw on a database of all player transactions in the NBA from the beginning of the 1976 season---when the NBA merged with the ABA uniting major professional basketball in the US under one league---through the completion of the 2018-19 season [@richardson2020].
+To address this question, we draw on a database of all player transactions in the NBA from the beginning of the 1976 season---when the NBA merged with the ABA uniting major professional basketball in the US under one league---through the completion of the 2018-19 season [@richardson2020].
 We compile this list of 1,977 trades into 43 annual trade networks, with the nodes representing teams, and each edge representing a unique trade between those teams.
@@
 Using these dates, there were no ambiguous trades that we were unable to attribute to a particular season.] The number of teams within each annual slice changes over time as the league gradually expanded from 22 to 30 teams.^[Unless otherwise specifically noted, all historical accounts in the text were pulled from Wikipedia and confirmed on nba.com.].
-The number of trades observed within each slice varies from year to year (range 22-86 or 0.8-2.9 if expressed in per-team averages; see Figure \@ref(fig:trades-per-year)), and teams exhibit differing rates of trading (see Figure \@ref(fig:trades-per-team)).
+The number of trades observed within each slice varies from year to year (range 22-86 or 0.8-2.9 when stated as per-team averages; see Figure \@ref(fig:trades-per-year)), and teams exhibit differing rates of trading (see Figure \@ref(fig:trades-per-team)).
 
 <!--
-![Figure 1. **Number of Trades by Year.** (normalized to a per team rate)](EDA_files/figure-gfm/trades-per-season-seasonwise-1.png)
+![Figure 1. **Average Trades by Year.** (standardized to a per team rate)](EDA_files/figure-gfm/trades-per-season-seasonwise-1.png)
 -->
 
-```{r trades-per-year, fig.cap="**Number of Trades by Year.** (normalized to a per team rate)"}
+```{r trades-per-year, fig.cap="**Number of Trades by Year.** (standardized to a per team rate)"}
 edb %>%
@@
 
-Figure \@ref(fig:trades-per-year) presents the trend of number of trades per season.
+Figure \@ref(fig:trades-per-year) presents the trend of the average trades per season.
 Since, the number of teams varies across the window, we standardize these values to represent the number of trades *per team* in each season.
 This pattern shows a roughly U-shaped trend, with the highest rates of trades (nearing 2 per team) occurring early and late in the period and the lowest rates occurring in the early 1990s.
-Notably, this low-point came *after* the introduction of unrestricted free-agency (in 1988), allowing players to sign with any team after their contract expires.
+Notably, this low-point came *after* the introduction of unrestricted free-agency (in 1988), which allowed players to sign with any team after their contract expires.
 As a corollary, Figure \@ref(fig:trades-per-team) presents the frequency distribution of trades per team across the full observed window.
@@
 <!--
-![Figure 2. **Number of Trades by Team.** (normalized to a per season rate)](EDA_files/figure-gfm/trades-per-season-teamwise-1.png)
+![Figure 2. **Number of Trades by Team.** (standardized to a per season rate)](EDA_files/figure-gfm/trades-per-season-teamwise-1.png)
 -->
 
-```{r trades-per-team, fig.cap="**Number of Trades by Team.** (normalized to a per season rate)"}
+```{r trades-per-team, fig.cap="**Average Trades by Team.** (standardized to a per season rate)"}
 # Time-aggregated network
@@
 
-To address our research question, we needed to supplement these trade data by compiling a list of each team's division membership.
+To address our main research question, we needed to supplement these trade data by compiling a list of each team's division membership.
 From the 1976-77 season through 2003-04, there were four divisions (Atlantic, Central, Midwest, and Pacific).
@@
 Secondarily, these steps also allow us to illustrate the proper way to statistically condition a question such as this, and to explain the need for doing so.
-This combination motivates our usage of a relatively recent development for statistically modeling weighted network data.
+This combination further motivates our usage of a relatively recent development for statistically modeling weighted network data.
 
@@
 **Dynamic Visualization of Annual Trade Networks** Note: Intra-division trades are highlighted in red.
-Node color and position refers to division.
-The upper-right menu allows the animation speed to be adjusted while visualizing the entire sequence, or the foward/back buttons at the bottom can be used to animate the visualization one annual slice (trade season) at a time.</p>
+Node color and position indicates division.
+The upper-right menu allows the animation speed to be adjusted while visualizing the entire sequence, or the foward/back buttons at the bottom can be used to step-wise animate the visualization one annual slice (trade season) at a time.</p>
 </div>
@@
 Our first step is to tabulate the number of trades observed both within and across divisions.
-This tabulation is presented in the form of what is referred to as a mixing matrix, which shows---at the level of divisions---how many trades occur between teams of the same division and between teams of differing divisions.
+This tabulation is presented in the form of a mixing matrix, which shows---at the level of divisions---how many trades occur between teams of the same division (diagonal) and between teams of differing divisions (off-diagonal).
 We present these mixing matrices in two aggregations.
-The first represents all years prior to division expansion, when there were four divisions, a period which spans from 1976-2004.
+The first represents when there were four divisions, i.e., all years prior to division expansion, a period which spans from 1976-2004.
 The second window began in 2005 and runs through the end of our analytic period at the end of the 2018-19 season, during which the teams were separated into six divisions.
-To provide a visual representation of the changes in these division-specific trade patterns across the examined period, Video 1 presents a dynamic visualization of each year's trade network.^[The appendices also include full replication code for all analyzes conducted and the generation of this visualization.]
+To provide a visual representation of the changes in these division-specific trade patterns across the examined period, Video 1 presents a dynamic visualization of each year's trade network.^[The appendices also include full replication code for all analyzes conducted and the generation of this dynamic visualization.]
 
@@
 As can be seen in Tables \@ref(tab:mm-1976-2004) and \@ref(tab:mm-2005-2019), these tabulations would appear to suggest that there is an avoidance of trading with members of one's own division.
-For example in the pre-expansion era only 22% (231/1062) of trades were within division, a ratio of nearly 3.6 trades across divisions for every one trade within division.
+For example in the pre-expansion era only 22% (233/1077) of trades were within division, a ratio of approximately 3.6 trades across divisions for every one trade within division.
 
 While these mixing matrices may appear to support the prohibition against teams trading within their division, there are two caveats for interpreting these patterns which we address in each of our subsequent analytic steps.
-In order to test for whether teams avoid trading within their division we must first account for each team's tendency to trade at all, before we can assess with whom they conduct those trades.
+First, in order to test for whether teams avoid trading within their division we must first account for each team's tendency to trade at all, before we can assess with whom they conduct those trades.
 For example, in the 1976-77 trade season, there were 35 trades, of which 7 were within division and 28 across divisions.
-If we account for how many trades each team made, yet assumed that they made those trades without respect to division of their trade partner, random expectation would be that on average 7.6 trades (sd 2.2) would be within division and 27.4 (sd=2.2) would be across divisions.
+If we account for how many trades each team made, yet assumed that they made those trades without respect to division of their trade partner, random expectation would be that on average 7.6 trades (sd 2.2) would be within division and 27.4 (sd 2.2) would be across divisions.
 That is, this observed mixing ratio of 4.0 trades across division for every one within is actually consistent with the random expectations, and thus does not conform to a within division avoidance pattern.
@@
 Our second analytic step therefore addresses these conditional distribution assumptions that are not apparent from the mixing matrices alone.
-To accomplish these properly conditioned versions of the test we use what is known as the $\alpha$-index for computing segregation in networks [@moody2001].
+To properly condition a test to account for baseline expectations, we use what is known as the $\alpha$-index for computing segregation in networks [@moody2001].
 Essentially, what the $\alpha$-index does is allow the estimates of frequency of ties between teams from the same division to be estimated as an odds ratio, which is properly conditioned to account for the number of trade partners one would expect to be within and across division, given the observed base rates of trading by each team, *if* they chose with whom to trade without respect to division [@Bojanowski:2014aa].
-Since this is a homophily statistic, testing for nodes within the *same* division, a significantly *negative* $\alpha$-index would be consistent with teams following the proposed prohibition.
+Since this is a homophily statistic, testing for ties between teams within the *same* division, a significantly *negative* $\alpha$-index would be consistent with teams following the proposed prohibition.
 Figure \@ref(fig:alphas) presents this $\alpha$-index for each trade season individually.
-What can be seen here is that in none of the observed trade years does the $\alpha$-index differ significantly from random expectation (OR=1).
+What can be seen here is that in *none* of the observed trade years does the $\alpha$-index differ significantly from random expectation (OR=1).
 In other words this improved conditional form of the test suggests that there is no observed prohibition against teams trading within division.
@@
 ERGM is a statistical model for explaining the structure of a network by means of various local tendencies for ties to be present or absent quantified by model terms such as density, degree, homophily, transitivity and so on.
-The magnitude of effect of each term is measured by an associated coefficient.
-In the player trade networks we are interested in verifying whether there is any divisional homophily effect, i.e.
+The magnitude of effect corresponding with each term is measured by an associated coefficient.
+In the player trade networks we are interested in verifying whether there is a (negative) divisional homophily effect, i.e.
 analogously to the above presented $\alpha$.
-This would indicated if (multiple) trades are more or less likely in pairs of teams belonging to the same division compared to to pairs in different divisions.
-A positive value of the coefficient would suggest homophily while negative values would indicate heterophily (the hypothesized avoidance of within-division trading).
+This would indicated if (multiple) trades are more or less likely in pairs of teams belonging to the same division compared to team pairs in different divisions.
+A positive value of the coefficient would suggest homophily, while negative values would indicate heterophily (the hypothesized avoidance of within-division trading).
 
-Adequate modeling a weighted network via an ERGM requires adapting model specification to the distribution of trade counts across dyads by choosing the reference measure [@krivit2012, sec.
+Adequate modeling a weighted network via an ERGM also requires adapting model specification to the distribution of trade counts across dyads by choosing an appropriate reference measure [@krivit2012, sec.
 5.2].
-While the typical approach involves assuming that the counts are Poisson-distributed, in practice the data is distributed differently with overprevalence of 0s (zero-inflation) or with greater variation than the Poisson distribution assumes (overdispersion).
-Table \@ref(tab:dyadic-trade-counts) summarises that trading is relatively rare in general and it happens even more rearely that a pair of teams trades more than once in the same season.
-Formal tests diagnoze overdispersion in this distribution [@cameron1990regression;@rpkg-AER].
+While the typical approach involves assuming that the counts are Poisson-distributed, in practice empirical data distributions often exhibit an over-abundance of 0s (zero-inflation) and/or  greater variation than the Poisson distribution assumes (overdispersion).
+Table \@ref(tab:dyadic-trade-counts) summarises that trading is relatively rare in general and it happens even more rarely that a pair of teams trades more than once in the same season.
+Formal tests of this distribution do indicate overdispersion [@cameron1990regression;@rpkg-AER].
 
@@
 
-At this point one could either choose a non-Poissonian reference measure or simplify data by dychotomizing the counts (any number of trades vs no trades at all) and proceeding with a binary ERGM.
-The homophily coefficient in the binary ERGM fit to all seasonal trade networks simultaneously, including season- and division-specific effects, is equal to 0.0122759 (standard error of 0.065, z-statistic of 0.188) showing no avoidance patterns.
+At this point we could either choose a non-Poissonian reference measure or simplify data by dychotomizing the counts (any number of trades vs no trades at all) and proceeding with a binary ERGM.
+We ran both versions, reaching the same conclusion.
+The homophily coefficient in the **binary** ERGM fit to all seasonal trade networks simultaneously, including season- and division-specific effects, is equal to 0.012 (standard error of 0.065, z-statistic of 0.188) again showing no avoidance patterns.
 The complete model is presented in Table \@ref(tab:model-binary-table) in the Appendix.
 
-Once properly conditioning the test for each of these necessary caveats, we find that in none of the seasons are teams more likely to avoid trading with other teams from within their division.
-As can be verified in the "funnel plot" presented in Figure \@ref(fig:seasonal-ergms), each the seasonal homophily coefficients fall within the white cone of statistical *in*significance.
+Applying the **valued** ERGM--to account for the caveats raised above, and with the properly slected reference distribution, we again find that in *none* of the seasons do teams exhibit a tendency to avoid trading with other teams from within their division.
+This is summarized with the "funnel plot" presented in Figure \@ref(fig:seasonal-ergms), that shows each of homophily coefficients from season-specific weighted ERGMs fall within the white cone of statistical *in*significance.
 
@@
 These results are presented in Table \@ref(tab:model1-table-short).
+In this final specificatio, we find an aggregate same-division homophily effect (coefficient 0.006, standard error of 0.059, z-statistic of 0.102), that again shows no avoidance pattern.
+Fitting similar models to the periods 1976-2004 and 2004-2019 separately leads to almost identical results.[^periodModels]
 
@@
 
-We find an aggregate same-division homophily effect of 0.0060264 (standard error of 0.059, z-statistic of 0.102), again showing no avoidance pattern.
-Fitting similar models to the periods 1976-2004 and 2004-2019 separately leads to almost identical results.[^periodModels]
-
 [^periodModels]: Results can be obtained from the authors upon request.
@@
 The GLM framework would not be able to estimate each of these types of features in such a model without violating model assumptions (e.g., some of these are explicitly modeling the *dependency* between multiple trades).
-As such, we hope that the illustration if how the weighted ERGM could be beneficial serves as a useful introduction to this modeling framework, which could be extended by future researchers using the data we provide.
+As such, in addition to our primary empirical conclusions, we hope that the illustration if how the weighted ERGM could be beneficial serves as a useful introduction to this modeling framework, which could be extended by future researchers using the data we provide.
 

### 2023-02-01 Manuscript.Rmd
typo

@@
 As can be seen in Tables \@ref(tab:mm-1976-2004) and \@ref(tab:mm-2005-2019), these tabulations would appear to suggest that there is an avoidance of trading with members of one's own division.
-For example in the pre-expansion era only 22% (233/1077) of trades were within division, a ratio of approximately 3.6 trades across divisions for every one trade within division.
+For example in the pre-expansion era (1976-2004) only 22% (233/1077) of trades were within division, a ratio of approximately 3.6 trades across divisions for every one trade within division.
 

### 2023-04-14 Manuscript.Rmd
condional acceptance revisions

@@
 
-A primary basis for such an expectation arises from organizations avoiding collaborating with thier (perceived) competitors.
+A primary basis for such an expectation arises from organizations avoiding collaborating with their (perceived) competitors.
 For example, within a field where the actors are (or are perceived to be) competitors with one another, they may avoid cooperating with one another because it is perceived to be a competitive disadvantage [@Hoffmann2018], though that assumption has been questioned [@Bengtsson1999; @peng12].
 Instead of cooperating directly with one's competitors, actors may therefore seek out means of collaboration with those outside the competitive field [@Soda_2017].
-Or they may develop more narrow difinitions of who their competitors are [@Barman_2002].
+Or they may develop more narrow definitions of who their competitors are [@Barman_2002].
 In combination, these could facilitate the avoidance of intra-division trades through prioritizing trades with teams who are not perceived as their direct competitors--i.e., from outside their own division.
@@
 Node color and position indicates division.
-The upper-right menu allows the animation speed to be adjusted while visualizing the entire sequence, or the foward/back buttons at the bottom can be used to step-wise animate the visualization one annual slice (trade season) at a time.</p>
+The upper-right menu allows the animation speed to be adjusted while visualizing the entire sequence, or the forward/back buttons at the bottom can be used to step-wise animate the visualization one annual slice (trade season) at a time.</p>
 </div>
@@
 While the typical approach involves assuming that the counts are Poisson-distributed, in practice empirical data distributions often exhibit an over-abundance of 0s (zero-inflation) and/or  greater variation than the Poisson distribution assumes (overdispersion).
-Table \@ref(tab:dyadic-trade-counts) summarises that trading is relatively rare in general and it happens even more rarely that a pair of teams trades more than once in the same season.
+Table \@ref(tab:dyadic-trade-counts) summarizes that trading is relatively rare in general and it happens even more rarely that a pair of teams trades more than once in the same season.
 Formal tests of this distribution do indicate overdispersion [@cameron1990regression;@rpkg-AER].
@@
 
-At this point we could either choose a non-Poissonian reference measure or simplify data by dychotomizing the counts (any number of trades vs no trades at all) and proceeding with a binary ERGM.
+At this point we could either choose a non-Poissonian reference measure or simplify data by dichotomizing the counts (any number of trades vs no trades at all) and proceeding with a binary ERGM.
 We ran both versions, reaching the same conclusion.
@@
 
-Applying the **valued** ERGM--to account for the caveats raised above, and with the properly slected reference distribution, we again find that in *none* of the seasons do teams exhibit a tendency to avoid trading with other teams from within their division.
+Applying the **valued** ERGM--to account for the caveats raised above, and with the properly selected reference distribution, we again find that in *none* of the seasons do teams exhibit a tendency to avoid trading with other teams from within their division.
 This is summarized with the "funnel plot" presented in Figure \@ref(fig:seasonal-ergms), that shows each of homophily coefficients from season-specific weighted ERGMs fall within the white cone of statistical *in*significance.
@@
 These results are presented in Table \@ref(tab:model1-table-short).
-In this final specificatio, we find an aggregate same-division homophily effect (coefficient 0.006, standard error of 0.059, z-statistic of 0.102), that again shows no avoidance pattern.
+In this final specification, we find an aggregate same-division homophily effect (coefficient 0.006, standard error of 0.059, z-statistic of 0.102), that again shows no avoidance pattern.
 Fitting similar models to the periods 1976-2004 and 2004-2019 separately leads to almost identical results.[^periodModels]
@@
 The lingering weak associations are even further reduced once including the weighting that is possible from multiple trades between partners in a single season.
+We are reticent to suggest that these findings necessarily alter the theoretical claims within organizational literature that competitors may avoid cooperating with one another [@Hoffmann2018, @Stewart1999]; however they do suggest that future scholarship may want to examine potential variation between the contexts where such prohibitions exist compared to others like the NBA case, where they seem to be violated.
 
@@
 However, the value of introducing the weighted ERGM approach with these data is that if subsequent researchers want to take the next step to provide an explanatory account that further examines which teams are likely to trade with which others, this modeling framework would allow for the simultaneous estimation of predictive factors that variously operate at the node level (e.g., is a team that finished lower in the previous year's standings more likely to trade), the dyad level (e.g., are teams with higher total salaries more likely to trade with other teams who have lower total salaries), or structural features above the dyad (e.g., the tendency to form trade-triples or larger configurations through multi-team deals).
-The GLM framework would not be able to estimate each of these types of features in such a model without violating model assumptions (e.g., some of these are explicitly modeling the *dependency* between multiple trades).
-As such, in addition to our primary empirical conclusions, we hope that the illustration if how the weighted ERGM could be beneficial serves as a useful introduction to this modeling framework, which could be extended by future researchers using the data we provide.
+The GLM framework would not be able to estimate each of these types of features in such a model without violating model assumptions (e.g., some of these are explicitly modeling the *dependency* between multiple trades).^[Additionally, to test a non-parametric version of our primary question here, we could employ a QAP, but that approach would not allow for a straightforward means to account for the valued relationships, nor the ready possibility to incorporate the types of predictors in potential extensions of this work as described above.] As such, in addition to our primary empirical conclusions, we hope that the illustration if how the weighted ERGM could be beneficial serves as a useful introduction to this modeling framework, which could be extended by future researchers using the data we provide.
 

### 2023-04-14 Manuscript.Rmd
fix incorrect QAP fn

@@
 However, the value of introducing the weighted ERGM approach with these data is that if subsequent researchers want to take the next step to provide an explanatory account that further examines which teams are likely to trade with which others, this modeling framework would allow for the simultaneous estimation of predictive factors that variously operate at the node level (e.g., is a team that finished lower in the previous year's standings more likely to trade), the dyad level (e.g., are teams with higher total salaries more likely to trade with other teams who have lower total salaries), or structural features above the dyad (e.g., the tendency to form trade-triples or larger configurations through multi-team deals).
-The GLM framework would not be able to estimate each of these types of features in such a model without violating model assumptions (e.g., some of these are explicitly modeling the *dependency* between multiple trades).^[Additionally, to test a non-parametric version of our primary question here, we could employ a QAP, but that approach would not allow for a straightforward means to account for the valued relationships, nor the ready possibility to incorporate the types of predictors in potential extensions of this work as described above.] As such, in addition to our primary empirical conclusions, we hope that the illustration if how the weighted ERGM could be beneficial serves as a useful introduction to this modeling framework, which could be extended by future researchers using the data we provide.
+The GLM framework would not be able to estimate each of these types of features in such a model without violating model assumptions (e.g., some of these are explicitly modeling the *dependency* between multiple trades).^[Additionally, the MRQAP framework would allow us to test the primary form of our question as a properly conditioned test for a weighted outcome, and to allow extensions that include testing nodal and dyadic covariates as described above.
+However, if structural features such as multi-team configurations or deals are part of those predictions, MRQAP proves more difficult for including such "supra-dyadic" structural features.] As such, in addition to our primary empirical conclusions, we hope that the illustration if how the weighted ERGM could be beneficial serves as a useful introduction to this modeling framework, which could be extended by future researchers using the data we provide.
 

### 2023-05-30 Manuscript.Rmd
Revising

@@
   ggplot(aes(x = season, y = or, ymin=ci.low, ymax=ci.high)) +
-  geom_hline(yintercept = 1, size = 0.5, color = "red") +
+  geom_hline(yintercept = 1, linewidth = 0.5, color = "red") +
   geom_pointrange() +
@@
 
+$$P( Y = y ) = \frac{\exp (
+\sum_{s=2}^S \theta_s g_1 (y, s) +
+\sum_{d=1}^S \theta_d g_1 (y, d) +
+\theta_h g_2 (y))}{\kappa(\theta_s, \theta_d, \theta_h)}$$
+
+where the sufficient statistics are:
+
+- $g_1(y, s)$ -- number of trade participations in season $s$
+- $g_1(y, d)$ -- number of trade participations by teams in division $d$
+- $g_2(y)$ -- number of trades involving teams from the same divisions
+
+The model is constrained to within-season dyads only.
+
 ```{r model-binary-table}
@@
 withr::with_options(
-  list(knitr.kable.NA = ""), {
-    structure(list(Effect = c("*Seasonal base rates*", "1977-1978",  "1978-1979", "1979-1980", "1980-1981", "1981-1982", "1982-1983",  "1983-1984", "1984-1985", "1985-1986", "1986-1987", "1987-1988",  "1988-1989", "1989-1990", "1990-1991", "1991-1992", "1992-1993",  "1993-1994", "1994-1995", "1995-1996", "1996-1997", "1997-1998",  "1998-1999", "1999-2000", "2000-2001", "2001-2002", "2002-2003",  "2003-2004", "2004-2005", "2005-2006", "2006-2007", "2007-2008",  "2008-2009", "2009-2010", "2010-2011", "2011-2012", "2012-2013",  "2013-2014", "2014-2015", "2015-2016", "2016-2017", "2017-2018",  "2018-2019", "Atlantic division", "Central division", "Midwest division",  "Northwest division", "Pacific division", "Southeast division",  "Southwest division", "*Homophily*"), Estimate = c(NA, 0.0644405709809631,  0.00191183013592781, 0.00191183013587444, -0.0510792301762593,  -0.00301475598725536, 0.11151751406511, 0.0273160881984102, -0.120873378851605,  -0.241797069713423, -0.034689299944754, -0.102736047536669, -0.217649092501239,  -0.30417868994459, -0.221517869773192, -0.423638103944277, -0.306766674453116,  -0.38207146557124, -0.492148757048376, -0.368437716751424, -0.351709490768931,  -0.21888283796225, -0.335437019204636, -0.38565088841675, -0.109253481582016,  -0.15570959635153, -0.351709490778105, -0.192860596693431, -0.187248236068849,  -0.22034655957727, -0.373121806940009, -0.33128353096099, -0.115852230243608,  -0.0873886627773311, -0.06903205140053, -0.304889898702513, -0.106234994930531,  -0.106234994930528, 0.0163922227340902, -0.18724823606885, -0.243383258589619,  -0.125604821725264, -0.0600276821843626, -0.801604197626109,  -0.866780460779051, -0.869801998775562, -0.71676022773136, -0.912537765260793,  -0.861071612165831, -0.665077455419437, 0.0122759266186398),      SE = c(NA, 0.127030461420487, 0.129814083679627, 0.129814092249876,      0.129291411629841, 0.12715500781249, 0.122756935416747, 0.125898331758024,      0.132729817906103, 0.139724669892161, 0.128542414167505,      0.131796473203285, 0.132010767728565, 0.131420860718578,      0.127137016042545, 0.138568763749447, 0.131454151198191,      0.135887273703477, 0.143380183694574, 0.130029601322226,      0.129122811845738, 0.122743181611443, 0.128263950828193,      0.130988252937235, 0.118460375222509, 0.120175658066723,      0.129122966384354, 0.12165157379233, 0.121780823545216, 0.122921411086859,      0.129126929340387, 0.127263407924174, 0.119539607693635,      0.118724498302347, 0.118221222501676, 0.126153907222097,      0.119259391114893, 0.119259390545633, 0.11609615644012, 0.121780823683373,      0.123755575999076, 0.119828869972253, 0.117980604919478,      0.0989169031985135, 0.0981917740131896, 0.100979355062305,      0.119436809946451, 0.0992411986882728, 0.12109213666063,      0.118908017386702, 0.0651683403015203), "95% CI" = c(NA,      "(-0.18;  0.31)", "(-0.25;  0.26)", "(-0.25;  0.26)", "(-0.30;  0.20)",      "(-0.25;  0.25)", "(-0.13;  0.35)", "(-0.22;  0.27)", "(-0.38;  0.14)",      "(-0.52;  0.03)", "(-0.29;  0.22)", "(-0.36;  0.16)", "(-0.48;  0.04)",      "(-0.56; -0.05)", "(-0.47;  0.03)", "(-0.70; -0.15)", "(-0.56; -0.05)",      "(-0.65; -0.12)", "(-0.77; -0.21)", "(-0.62; -0.11)", "(-0.60; -0.10)",      "(-0.46;  0.02)", "(-0.59; -0.08)", "(-0.64; -0.13)", "(-0.34;  0.12)",      "(-0.39;  0.08)", "(-0.60; -0.10)", "(-0.43;  0.05)", "(-0.43;  0.05)",      "(-0.46;  0.02)", "(-0.63; -0.12)", "(-0.58; -0.08)", "(-0.35;  0.12)",      "(-0.32;  0.15)", "(-0.30;  0.16)", "(-0.55; -0.06)", "(-0.34;  0.13)",      "(-0.34;  0.13)", "(-0.21;  0.24)", "(-0.43;  0.05)", "(-0.49;  0.00)",      "(-0.36;  0.11)", "(-0.29;  0.17)", "(-1.00; -0.61)", "(-1.06; -0.67)",      "(-1.07; -0.67)", "(-0.95; -0.48)", "(-1.11; -0.72)", "(-1.10; -0.62)",      "(-0.90; -0.43)", "(-0.12;  0.14)"), "p-value" = c(NA, 0.612,      0.988, 0.988, 0.693, 0.981, 0.364, 0.828, 0.362, 0.084, 0.787,      0.436, 0.099, 0.021, 0.081, 0.002, 0.02, 0.005, 0.001, 0.005,      0.006, 0.075, 0.009, 0.003, 0.356, 0.195, 0.006, 0.113, 0.124,      0.073, 0.004, 0.009, 0.332, 0.462, 0.559, 0.016, 0.373, 0.373,      0.888, 0.124, 0.049, 0.295, 0.611, 0, 0, 0, 0, 0, 0, 0, 0.851     )), class = c("tbl_df", "tbl", "data.frame"), row.names = c(NA,  -51L)) %>%
+  list(knitr.kable.NA = ""), 
+  {
+    structure(list(Effect = c("*Seasonal base rates*", "1977-1978",  "1978-1979", "1979-1980", "1980-1981", "1981-1982", "1982-1983",  "1983-1984", "1984-1985", "1985-1986", "1986-1987", "1987-1988",  "1988-1989", "1989-1990", "1990-1991", "1991-1992", "1992-1993",  "1993-1994", "1994-1995", "1995-1996", "1996-1997", "1997-1998",  "1998-1999", "1999-2000", "2000-2001", "2001-2002", "2002-2003",  "2003-2004", "2004-2005", "2005-2006", "2006-2007", "2007-2008",  "2008-2009", "2009-2010", "2010-2011", "2011-2012", "2012-2013",  "2013-2014", "2014-2015", "2015-2016", "2016-2017", "2017-2018",  "2018-2019", "Atlantic division", "Central division", "Midwest division",  "Northwest division", "Pacific division", "Southeast division",  "Southwest division", "*Homophily*"), Estimate = c(NA, 0.0644405709809631,  0.00191183013592781, 0.00191183013587444, -0.0510792301762593,  -0.00301475598725536, 0.11151751406511, 0.0273160881984102, -0.120873378851605,  -0.241797069713423, -0.034689299944754, -0.102736047536669, -0.217649092501239,  -0.30417868994459, -0.221517869773192, -0.423638103944277, -0.306766674453116,  -0.38207146557124, -0.492148757048376, -0.368437716751424, -0.351709490768931,  -0.21888283796225, -0.335437019204636, -0.38565088841675, -0.109253481582016,  -0.15570959635153, -0.351709490778105, -0.192860596693431, -0.187248236068849,  -0.22034655957727, -0.373121806940009, -0.33128353096099, -0.115852230243608,  -0.0873886627773311, -0.06903205140053, -0.304889898702513, -0.106234994930531,  -0.106234994930528, 0.0163922227340902, -0.18724823606885, -0.243383258589619,  -0.125604821725264, -0.0600276821843626, -0.801604197626109,  -0.866780460779051, -0.869801998775562, -0.71676022773136, -0.912537765260793,  -0.861071612165831, -0.665077455419437, 0.0122759266186398),      SE = c(NA, 0.127030461420487, 0.129814083679627, 0.129814092249876,      0.129291411629841, 0.12715500781249, 0.122756935416747, 0.125898331758024,      0.132729817906103, 0.139724669892161, 0.128542414167505,      0.131796473203285, 0.132010767728565, 0.131420860718578,      0.127137016042545, 0.138568763749447, 0.131454151198191,      0.135887273703477, 0.143380183694574, 0.130029601322226,      0.129122811845738, 0.122743181611443, 0.128263950828193,      0.130988252937235, 0.118460375222509, 0.120175658066723,      0.129122966384354, 0.12165157379233, 0.121780823545216, 0.122921411086859,      0.129126929340387, 0.127263407924174, 0.119539607693635,      0.118724498302347, 0.118221222501676, 0.126153907222097,      0.119259391114893, 0.119259390545633, 0.11609615644012, 0.121780823683373,      0.123755575999076, 0.119828869972253, 0.117980604919478,      0.0989169031985135, 0.0981917740131896, 0.100979355062305,      0.119436809946451, 0.0992411986882728, 0.12109213666063,      0.118908017386702, 0.0651683403015203), "95% CI" = c(NA,      "(-0.18;  0.31)", "(-0.25;  0.26)", "(-0.25;  0.26)", "(-0.30;  0.20)",      "(-0.25;  0.25)", "(-0.13;  0.35)", "(-0.22;  0.27)", "(-0.38;  0.14)",      "(-0.52;  0.03)", "(-0.29;  0.22)", "(-0.36;  0.16)", "(-0.48;  0.04)",      "(-0.56; -0.05)", "(-0.47;  0.03)", "(-0.70; -0.15)", "(-0.56; -0.05)",      "(-0.65; -0.12)", "(-0.77; -0.21)", "(-0.62; -0.11)", "(-0.60; -0.10)",      "(-0.46;  0.02)", "(-0.59; -0.08)", "(-0.64; -0.13)", "(-0.34;  0.12)",      "(-0.39;  0.08)", "(-0.60; -0.10)", "(-0.43;  0.05)", "(-0.43;  0.05)",      "(-0.46;  0.02)", "(-0.63; -0.12)", "(-0.58; -0.08)", "(-0.35;  0.12)",      "(-0.32;  0.15)", "(-0.30;  0.16)", "(-0.55; -0.06)", "(-0.34;  0.13)",      "(-0.34;  0.13)", "(-0.21;  0.24)", "(-0.43;  0.05)", "(-0.49;  0.00)",      "(-0.36;  0.11)", "(-0.29;  0.17)", "(-1.00; -0.61)", "(-1.06; -0.67)",      "(-1.07; -0.67)", "(-0.95; -0.48)", "(-1.11; -0.72)", "(-1.10; -0.62)",      "(-0.90; -0.43)", "(-0.12;  0.14)"), "p-value" = c(NA, 0.612,      0.988, 0.988, 0.693, 0.981, 0.364, 0.828, 0.362, 0.084, 0.787,      0.436, 0.099, 0.021, 0.081, 0.002, 0.02, 0.005, 0.001, 0.005,      0.006, 0.075, 0.009, 0.003, 0.356, 0.195, 0.006, 0.113, 0.124,      0.073, 0.004, 0.009, 0.332, 0.462, 0.559, 0.016, 0.373, 0.373,      0.888, 0.124, 0.049, 0.295, 0.611, 0, 0, 0, 0, 0, 0, 0, 0.851     )), class = c("tbl_df", "tbl", "data.frame"), row.names = c(NA,  -51L)) |>
       knitr::kable(
@@
       )
-  }
-)
+  })
 ```
@@
 ## Weighted ERGM model fit {.appendix}
+
+```{r, eval=FALSE, include=FALSE}
+net ~ CMP + nodefactor("season", levels = NULL, form = "sum") + 
+    nodematch("div")
+```
+
+$$P( Y = y ) = \frac{\exp (
+\text{CMP} 
+\sum_{s=2}^S \theta_s g_1 (y, s) +
+\theta_h g_2 (y))}{\kappa(\theta_s, \theta_d, \theta_h)}$$
+
+where the sufficient statistics are:
+
+-
+
+The model is constrained to within-season dyads only.
 

### 2023-05-31 Manuscript.Rmd
Add Appendix with model specification details

@@
 
+## Details of model specification {.appendix}
+
+Basic notation:
+
+- $N$ -- number of unique team-season pairs
+- $i, j \in \{1, ..., N\}$ -- team playing in a season (a team-season pair)
+- $\{s_1, ..., s_i, ..., s_N\}$ -- season of each team-season pair
+- $s_i \in \{1, ..., s, ..., S\}$ -- seasons
+- $\{d_1, ..., d_i, ..., d_N\}$ -- division of each team-season pair
+- $d_i \in \{1, ..., d, ..., D\}$ -- divisions
+- $[y_{ij}]_{N \times N}$ -- A block-diagonal adjacency matrix of trades between teams in  seasons.
+For simplicity, but abusing notation slightly, $y_{ij}$ is the number of trades in a season between teams $i$ and $j$ in case of valued ERGMs and, in the case of binary ERGM, trade count dichotomised to $\{0,1\}$ (no trade vs any number of trades).
+
+In words, each node of the network corresponds to a team playing in the particular season.
+Nodes are grouped into seasonal blocks such that trades can happen only within seasons (ties within blocks) but cannot happen between seasons (no ties between blocks).
+
+The models use the following sufficient statistics:
+
+- *Season-specific propensity to trade* represented by the number of trade participations by teams in season $s$:
+
+$$g_1(y, s) = \sum_{i,j:\; s_i=s \,\vee\, s_j=s} y_{ij}$$
+
+- *Division-specific propensity to trade* represented by the number of trade participations by teams in division $d$:
+
+$$g_1(y, d) = \sum_{i,j:\; d_i = d \,\vee\, d_j=d} y_{ij}$$
+
+- *Propensity towards within-division trading* represented by the number of trades involving teams from the same division:
+
+$$g_2(y) = \sum_{i,j:\; d_i = d_j} y_{ij}$$
+
+- *Over-/under-dispersion* statistic which modifies the reference measure into a Conway-Maxwell-Poisson-reference ERGM [c.f. @krivit2012, sec. 5.2.3]:
+
+$$g_3(y) = \sum_{i,j} \log(y_{ij}!)$$
+
+Pooled binary ERGM is specified as:
+
+$$P( Y = y ) = \frac{\exp \left(
+\sum_{s=2}^S \theta_s g_1 (y, s) +
+\sum_{d=1}^D \theta_d g_1 (y, d) +
+\theta_2 g_2 (y) \right)}{\kappa}$$
+
+Pooled valued ERGM  is specified as:
+
+$$P( Y = y ) = \frac{\exp \left(
+\theta_3 g_3 (y) + 
+\sum_{s=1}^S \theta_s g_1 (y, s) +
+\theta_2 g_2 (y) \right)}{\kappa}$$
+
 ## Binary ERGM model fit {.appendix}
-
-$$P( Y = y ) = \frac{\exp (
-\sum_{s=2}^S \theta_s g_1 (y, s) +
-\sum_{d=1}^S \theta_d g_1 (y, d) +
-\theta_h g_2 (y))}{\kappa(\theta_s, \theta_d, \theta_h)}$$
-
-where the sufficient statistics are:
-
-- $g_1(y, s)$ -- number of trade participations in season $s$
-- $g_1(y, d)$ -- number of trade participations by teams in division $d$
-- $g_2(y)$ -- number of trades involving teams from the same divisions
-
-The model is constrained to within-season dyads only.
 
@@
 
-$$P( Y = y ) = \frac{\exp (
-\text{CMP} 
-\sum_{s=2}^S \theta_s g_1 (y, s) +
-\theta_h g_2 (y))}{\kappa(\theta_s, \theta_d, \theta_h)}$$
-
-where the sufficient statistics are:
-
--
-
-The model is constrained to within-season dyads only.
-
 ```{r model1-table}

### 2023-05-31 Manuscript.Rmd
Addressing pooling&specification comments...
... and adding rendered manuscript.

@@
 
-Applying the **valued** ERGM--to account for the caveats raised above, and with the properly selected reference distribution, we again find that in *none* of the seasons do teams exhibit a tendency to avoid trading with other teams from within their division.
+Applying the **valued** ERGM--to account for the caveats raised above, and with the Conway-Maxwell-Poisson reference measure, we again find that in *none* of the seasons do teams exhibit a tendency to avoid trading with other teams from within their division.
 This is summarized with the "funnel plot" presented in Figure \@ref(fig:seasonal-ergms), that shows each of homophily coefficients from season-specific weighted ERGMs fall within the white cone of statistical *in*significance.
@@
 
-To summarize the results from Figure \@ref(fig:seasonal-ergms) in a single model, we also fit a pooled ERGM across all seasons, including a base-rate effect for each year's trade volume, similarly to the binary ERGM mentioned above.
-These results are presented in Table \@ref(tab:model1-table-short).
+To summarize the results from Figure \@ref(fig:seasonal-ergms) in a single model, we also fit a pooled ERGM across all seasons, including a base-rate effect for each year's trade volume, similarly to the binary ERGM mentioned above.[^modelSpecification] These results are presented in Table \@ref(tab:model1-table-short).
 In this final specification, we find an aggregate same-division homophily effect (coefficient 0.006, standard error of 0.059, z-statistic of 0.102), that again shows no avoidance pattern.
@@
 [^periodModels]: Results can be obtained from the authors upon request.
+
+[^modelSpecification]: Pooling was performed by merging all seasonal networks into a single block-diagonal adjacency matrix and fitting a single model with block-diagonal sample space contraints.
+Further model specification details are presented in the Appendix.
 

### 2023-05-31 Manuscript.Rmd
ms updates according to memo

@@
 To examine this question, we construct a date-stamped network of all trades in the NBA between June 1976 and May 2019.
-We then use yearly weighted exponential random graph models to estimate the likelihood of teams avoiding within-division trade partners, and whether that pattern changes through time.
+We then use season-specific weighted exponential random graph models to estimate the likelihood of teams avoiding within-division trade partners, and how consistent that pattern is across the observed period.
 In addition to the empirical question, this analysis serves to demonstrate the necessity and difficulty of constructing the proper baseline for statistical comparison.
@@
 Each trade is assigned to the "trade season" which runs from the end of the previous season through the corresponding season's trade deadline.^[The opening of the trade season begins either (1) with the conclusion of the previous season's playoffs (early in the observed period), or on a specified date each summer (later in the observed window).
-Using these dates, there were no ambiguous trades that we were unable to attribute to a particular season.] The number of teams within each annual slice changes over time as the league gradually expanded from 22 to 30 teams.^[Unless otherwise specifically noted, all historical accounts in the text were pulled from Wikipedia and confirmed on nba.com.].
-The number of trades observed within each slice varies from year to year (range 22-86 or 0.8-2.9 when stated as per-team averages; see Figure \@ref(fig:trades-per-year)), and teams exhibit differing rates of trading (see Figure \@ref(fig:trades-per-team)).
+Using these dates, there were no ambiguous trades that we were unable to attribute to a particular season.] The number of teams within each annual slice changes over time as the league gradually expanded from 22 to 30 teams.^[Unless otherwise specifically noted, all historical accounts in the text were pulled from Wikipedia and confirmed on nba.com.] The number of trades observed within each slice varies from season to season (range 22-86 or 0.8-2.9 when stated as per-team averages; see Figure \@ref(fig:trades-per-year)), and teams exhibit differing rates of trading (see Figure \@ref(fig:trades-per-team)).
 
 <!--
-![Figure 1. **Average Trades by Year.** (standardized to a per team rate)](EDA_files/figure-gfm/trades-per-season-seasonwise-1.png)
+![Figure 1. **Average Trades by Season** (standardized to a per team rate)](EDA_files/figure-gfm/trades-per-season-seasonwise-1.png)
 -->
 
-```{r trades-per-year, fig.cap="**Number of Trades by Year.** (standardized to a per team rate)"}
+```{r trades-per-year, fig.cap="**Number of Trades by Season.** (standardized to a per team rate)"}
 edb %>%
@@
 As a corollary, Figure \@ref(fig:trades-per-team) presents the frequency distribution of trades per team across the full observed window.
-There are some teams (e.g., the Spurs) who consistently trade less frequently, and others who are much more trade active (e.g., the Mavericks).
+There are some teams (e.g., the Spurs) who consistently trade less frequently, and others who are much more trade active (e.g., the 76ers).
 
@@
 We present these mixing matrices in two aggregations.
-The first represents when there were four divisions, i.e., all years prior to division expansion, a period which spans from 1976-2004.
+The first represents when there were four divisions, i.e., all seasons prior to division expansion, a period which spans from 1976-2004.
 The second window began in 2005 and runs through the end of our analytic period at the end of the 2018-19 season, during which the teams were separated into six divisions.
-To provide a visual representation of the changes in these division-specific trade patterns across the examined period, Video 1 presents a dynamic visualization of each year's trade network.^[The appendices also include full replication code for all analyzes conducted and the generation of this dynamic visualization.]
+To provide a visual representation of the changes in these division-specific trade patterns across the examined period, Video 1 presents a dynamic visualization of each season's trade network.^[The appendices also include full replication code for all analyzes conducted and the generation of this dynamic visualization.]
 
@@
 Essentially, what the $\alpha$-index does is allow the estimates of frequency of ties between teams from the same division to be estimated as an odds ratio, which is properly conditioned to account for the number of trade partners one would expect to be within and across division, given the observed base rates of trading by each team, *if* they chose with whom to trade without respect to division [@Bojanowski:2014aa].
-Since this is a homophily statistic, testing for ties between teams within the *same* division, a significantly *negative* $\alpha$-index would be consistent with teams following the proposed prohibition.
-Figure \@ref(fig:alphas) presents this $\alpha$-index for each trade season individually.
-What can be seen here is that in *none* of the observed trade years does the $\alpha$-index differ significantly from random expectation (OR=1).
+Since this is a homophily statistic, testing for ties between teams within the *same* division, an $\alpha$-index significantly *less than* one would be consistent with teams following the proposed prohibition.
+Figure \@ref(fig:alphas) presents this $\alpha$-index (with 95% confidence interval) for each trade season individually.
+What can be seen here is that in *none* of the observed trade seasons does the $\alpha$-index differ significantly from random expectation (OR=1).
 In other words this improved conditional form of the test suggests that there is no observed prohibition against teams trading within division.
@@
   xlab("Season") +
-  ylab("Within Division Trades (OR)") +
+  ylab("Within Division Trades (OR w/ 95% CI)") +
   theme(
@@
 analogously to the above presented $\alpha$.
-This would indicated if (multiple) trades are more or less likely in pairs of teams belonging to the same division compared to team pairs in different divisions.
+This would indicate if (multiple) trades are more or less likely in pairs of teams belonging to the same division compared to team pairs in different divisions.
 A positive value of the coefficient would suggest homophily, while negative values would indicate heterophily (the hypothesized avoidance of within-division trading).
 
-Adequate modeling a weighted network via an ERGM also requires adapting model specification to the distribution of trade counts across dyads by choosing an appropriate reference measure [@krivit2012, sec.
+Appropriate modeling for a weighted network via an ERGM also requires adapting model specification to the distribution of trade counts across dyads by choosing an appropriate reference measure [@krivit2012, sec.
 5.2].
@@
 
-To summarize the results from Figure \@ref(fig:seasonal-ergms) in a single model, we also fit a pooled ERGM across all seasons, including a base-rate effect for each year's trade volume, similarly to the binary ERGM mentioned above.[^modelSpecification] These results are presented in Table \@ref(tab:model1-table-short).
+To summarize the results from Figure \@ref(fig:seasonal-ergms) in a single model, we also fit a pooled ERGM across all seasons, including a base-rate effect for each season's trade volume, similarly to the binary ERGM mentioned above.[^modelSpecification] These results are presented in Table \@ref(tab:model1-table-short).
 In this final specification, we find an aggregate same-division homophily effect (coefficient 0.006, standard error of 0.059, z-statistic of 0.102), that again shows no avoidance pattern.
@@
 
-[^modelSpecification]: Pooling was performed by merging all seasonal networks into a single block-diagonal adjacency matrix and fitting a single model with block-diagonal sample space contraints.
+[^modelSpecification]: Pooling was performed by merging all seasonal networks into a single block-diagonal adjacency matrix and fitting a single model with block-diagonal sample space constraints.
 Further model specification details are presented in the Appendix.
@@
 It is worth mentioning that the models we have estimated in the results for Figure \@ref(fig:seasonal-ergms) are actually no different from what would have been possible through a generalized linear model predicting the (weighted) number of ties observed between pairs of teams.
-However, the value of introducing the weighted ERGM approach with these data is that if subsequent researchers want to take the next step to provide an explanatory account that further examines which teams are likely to trade with which others, this modeling framework would allow for the simultaneous estimation of predictive factors that variously operate at the node level (e.g., is a team that finished lower in the previous year's standings more likely to trade), the dyad level (e.g., are teams with higher total salaries more likely to trade with other teams who have lower total salaries), or structural features above the dyad (e.g., the tendency to form trade-triples or larger configurations through multi-team deals).
+However, the value of introducing the weighted ERGM approach with these data is that if subsequent researchers want to take the next step to provide an explanatory account that further examines which teams are likely to trade with which others, this modeling framework would allow for the simultaneous estimation of predictive factors that variously operate at the node level (e.g., is a team that finished lower in the previous season's standings more likely to trade), the dyad level (e.g., are teams with higher total salaries more likely to trade with other teams who have lower total salaries), or structural features above the dyad (e.g., the tendency to form trade-triples or larger configurations through multi-team deals).
 The GLM framework would not be able to estimate each of these types of features in such a model without violating model assumptions (e.g., some of these are explicitly modeling the *dependency* between multiple trades).^[Additionally, the MRQAP framework would allow us to test the primary form of our question as a properly conditioned test for a weighted outcome, and to allow extensions that include testing nodal and dyadic covariates as described above.
-However, if structural features such as multi-team configurations or deals are part of those predictions, MRQAP proves more difficult for including such "supra-dyadic" structural features.] As such, in addition to our primary empirical conclusions, we hope that the illustration if how the weighted ERGM could be beneficial serves as a useful introduction to this modeling framework, which could be extended by future researchers using the data we provide.
+However, if structural features such as multi-team configurations or deals are part of those predictions, MRQAP proves more difficult for including such "supra-dyadic" structural features.] As such, in addition to our primary empirical conclusions--that within division trades are *not* avoided by NBA franchises--we hope that the illustration of how the weighted ERGM could be beneficial serves as a useful introduction to this modeling framework, and may be put to further examination by future researchers using the data and code we provide.
 
@@
 
-We appreciate feedback we received on this paper from Skye Bender-deMoll, Pavel Krivitsky and David Schaefer.
+We appreciate feedback we received on previous iterations of this work from Skye Bender-deMoll, Pavel Krivitsky and David Schaefer.

### 2023-06-01 Manuscript.Rmd
Changes for DOCX rendering
Added 'bookdown::workd_document2' output format.
Renamed 'description' to 'abstract' for DOCX rendering.

@@
 date: "May 10, 2022"
-description: "Within US professional sports, trades within one's own division are often perceived to be disadvantageous.
+abstract: "Within US professional sports, trades within one's own division are often perceived to be disadvantageous.
 We ask how common this practice is.
@@
     keep_md: true
+  bookdown::word_document2: default
 link-citations: true

### 2023-08-02 Manuscript.Rmd
Add PDF rendering
Uses bookdown::pdf_document2

@@
   bookdown::word_document2: default
+  bookdown::pdf_document2:
+    keep_tex: true
 link-citations: true

### 2023-08-07 Manuscript.Rmd
Rename bib file and update metadata
Underscore in Bib file name trips the rendering.
Update author and affiliation metadata.

--- 2023-06-01 Manuscript.Rmd
@@
   bookdown::word_document2: default
+  bookdown::pdf_document2:
+    keep_tex: true
 link-citations: true
 repository_url: https://github.com/mbojan/nba-trades
-bibliography: nba_trades.bib
+bibliography: nba-trades.bib
 editor_options: 

### 2023-08-07 Manuscript.qmd
Rename bib file and update metadata
Underscore in Bib file name trips the rendering.
Update author and affiliation metadata.

--- 2023-06-01 Manuscript.Rmd
@@
 author: 
-  - first_name: jimi
-    last_name: adams
-    orcid_id: "0000-0001-5043-1149"
-    affiliation: University of Colorado Denver, USA
-  - first_name: Michał
-    last_name: Bojanowski
-    orcid_id: "0000-0001-7503-852X"
-    affiliation: Kozminski University, Warsaw, Poland
+  - name: jimi adams
+    orcid: "0000-0001-5043-1149"
+    affiliations: "University of Colorado Denver, USA"
+  - name: Michał Bojanowski
+    orcid: "0000-0001-7503-852X"
+    affiliations:
+      - name: "Autonomous University of Barcelona"
+        city: Barcelona
+        country: Spain
+      - name: "Kozminski University"
+        city: Warsaw
+        country: Poland
 date: "May 10, 2022"
@@
 We find limited-to-no support for the popular perception.<br/><br/>**Draft under review**."
-output: 
-  distill::distill_article:
-    toc: true
-    code_folding: false
-    keep_md: true
-  bookdown::word_document2: default
-link-citations: true
+format: 
+  cambridge-medium-pdf:
+    keep-tex: true
+    cite-method: biblatex
 repository_url: https://github.com/mbojan/nba-trades
-bibliography: nba_trades.bib
+bibliography: nba-trades.bib
 editor_options: 
@@
 The lingering weak associations are even further reduced once including the weighting that is possible from multiple trades between partners in a single season.
-We are reticent to suggest that these findings necessarily alter the theoretical claims within organizational literature that competitors may avoid cooperating with one another [@Hoffmann2018, @Stewart1999]; however they do suggest that future scholarship may want to examine potential variation between the contexts where such prohibitions exist compared to others like the NBA case, where they seem to be violated.
+We are reticent to suggest that these findings necessarily alter the theoretical claims within organizational literature that competitors may avoid cooperating with one another [@Hoffmann2018; @Stewart1999]; however they do suggest that future scholarship may want to examine potential variation between the contexts where such prohibitions exist compared to others like the NBA case, where they seem to be violated.
 

### 2023-08-07 Manuscript.qmd
Drop "draft" from abstract.

@@
 In addition to the empirical question, this analysis serves to demonstrate the necessity and difficulty of constructing the proper baseline for statistical comparison.
-We find limited-to-no support for the popular perception.<br/><br/>**Draft under review**."
+We find limited-to-no support for the popular perception."
 format: 

### 2023-08-07 Manuscript.qmd
Fix affiliations and journal metadata

@@
     affiliations:
-      - name: "Autonomous University of Barcelona"
-        city: Barcelona
-        country: Spain
-      - name: "Kozminski University"
-        city: Warsaw
-        country: Poland
+      - name: "Autonomous University of Barcelona, Barcelona, Spain"
+      - name: "Kozminski University, Warsaw, Poland"
 date: "May 10, 2022"
@@
     cite-method: biblatex
+    journal: nws
+    year: 2023
+    volume: NA
 repository_url: https://github.com/mbojan/nba-trades

### 2023-08-07 Manuscript.qmd
Fix figure and table references

@@
 Each trade is assigned to the "trade season" which runs from the end of the previous season through the corresponding season's trade deadline.^[The opening of the trade season begins either (1) with the conclusion of the previous season's playoffs (early in the observed period), or on a specified date each summer (later in the observed window).
-Using these dates, there were no ambiguous trades that we were unable to attribute to a particular season.] The number of teams within each annual slice changes over time as the league gradually expanded from 22 to 30 teams.^[Unless otherwise specifically noted, all historical accounts in the text were pulled from Wikipedia and confirmed on nba.com.] The number of trades observed within each slice varies from season to season (range 22-86 or 0.8-2.9 when stated as per-team averages; see Figure \@ref(fig:trades-per-year)), and teams exhibit differing rates of trading (see Figure \@ref(fig:trades-per-team)).
+Using these dates, there were no ambiguous trades that we were unable to attribute to a particular season.] The number of teams within each annual slice changes over time as the league gradually expanded from 22 to 30 teams.^[Unless otherwise specifically noted, all historical accounts in the text were pulled from Wikipedia and confirmed on nba.com.] The number of trades observed within each slice varies from season to season (range 22-86 or 0.8-2.9 when stated as per-team averages; see @fig-trades-per-year), and teams exhibit differing rates of trading (see @fig-trades-per-team).
 
@@
 
-```{r trades-per-year, fig.cap="**Number of Trades by Season.** (standardized to a per team rate)"}
+```{r fig-trades-per-year, fig.cap="**Number of Trades by Season.** (standardized to a per team rate)"}
 edb %>%
@@
 
-Figure \@ref(fig:trades-per-year) presents the trend of the average trades per season.
+@fig-trades-per-year presents the trend of the average trades per season.
 Since, the number of teams varies across the window, we standardize these values to represent the number of trades *per team* in each season.
@@
 Notably, this low-point came *after* the introduction of unrestricted free-agency (in 1988), which allowed players to sign with any team after their contract expires.
-As a corollary, Figure \@ref(fig:trades-per-team) presents the frequency distribution of trades per team across the full observed window.
+As a corollary, @fig-trades-per-team presents the frequency distribution of trades per team across the full observed window.
 There are some teams (e.g., the Spurs) who consistently trade less frequently, and others who are much more trade active (e.g., the 76ers).
@@
 
-```{r trades-per-team, fig.cap="**Average Trades by Team.** (standardized to a per season rate)"}
+```{r fig-trades-per-team, fig.cap="**Average Trades by Team.** (standardized to a per season rate)"}
 # Time-aggregated network
@@
 
-```{r mm-1976-2004}
+```{r tbl-mm-1976-2004}
+#| tbl-cap: "**Mixing Matrix for the period 1976 - 2004**"
 mm_1976_2004 <- edb %>%
@@
     rename(`**1976 - 2004**` = div_from) %>%
-    knitr::kable(
-      caption = "**Mixing Matrix for the period 1976 - 2004**"
-    )
+    knitr::kable()
 )
@@
 
-```{r mm-2005-2019}
+```{r tbl-mm-2005-2019}
+#| tbl-cap: "**Mixing Matrix for the period 2005 - 2019**"
 mm_2005_2019 <- edb %>%
@@
     rename(`**2005 - 2019**` = div_from) %>%
-    knitr::kable(
-      caption = "**Mixing Matrix for the period 2005 - 2019**"
-    )
+    knitr::kable()
 )
@@
 
-As can be seen in Tables \@ref(tab:mm-1976-2004) and \@ref(tab:mm-2005-2019), these tabulations would appear to suggest that there is an avoidance of trading with members of one's own division.
+As can be seen in Tables [-@tbl-mm-1976-2004] and [-@tbl-mm-2005-2019], these tabulations would appear to suggest that there is an avoidance of trading with members of one's own division.
 For example in the pre-expansion era (1976-2004) only 22% (233/1077) of trades were within division, a ratio of approximately 3.6 trades across divisions for every one trade within division.
@@
 Since this is a homophily statistic, testing for ties between teams within the *same* division, an $\alpha$-index significantly *less than* one would be consistent with teams following the proposed prohibition.
-Figure \@ref(fig:alphas) presents this $\alpha$-index (with 95% confidence interval) for each trade season individually.
+Figure [-@fig-alphas] presents this $\alpha$-index (with 95% confidence interval) for each trade season individually.
 What can be seen here is that in *none* of the observed trade seasons does the $\alpha$-index differ significantly from random expectation (OR=1).
@@
 
-```{r alphas, fig.cap="**Divisional Homophily in NBA Trades by Season.** (as computed by Moody's $\\alpha$-index)."}
+```{r fig-alphas, fig.cap="**Divisional Homophily in NBA Trades by Season.** (as computed by Moody's $\\alpha$-index)."}
 netlist <- lapply(graphlist, intergraph::asNetwork)
@@
 While the typical approach involves assuming that the counts are Poisson-distributed, in practice empirical data distributions often exhibit an over-abundance of 0s (zero-inflation) and/or  greater variation than the Poisson distribution assumes (overdispersion).
-Table \@ref(tab:dyadic-trade-counts) summarizes that trading is relatively rare in general and it happens even more rarely that a pair of teams trades more than once in the same season.
+Table [-@tbl-dyadic-trade-counts] summarizes that trading is relatively rare in general and it happens even more rarely that a pair of teams trades more than once in the same season.
 Formal tests of this distribution do indicate overdispersion [@cameron1990regression;@rpkg-AER].
 
-```{r dyadic-trade-counts}
+```{r tbl-dyadic-trade-counts}
 # MB: This is pasted from the calculation in privaterepo/overdispersion.Rmd
@@
 Applying the **valued** ERGM--to account for the caveats raised above, and with the Conway-Maxwell-Poisson reference measure, we again find that in *none* of the seasons do teams exhibit a tendency to avoid trading with other teams from within their division.
-This is summarized with the "funnel plot" presented in Figure \@ref(fig:seasonal-ergms), that shows each of homophily coefficients from season-specific weighted ERGMs fall within the white cone of statistical *in*significance.
+This is summarized with the "funnel plot" presented in Figure [-@fig-seasonal-ergms], that shows each of homophily coefficients from season-specific weighted ERGMs fall within the white cone of statistical *in*significance.
 
@@
 
-```{r seasonal-ergms, fig.cap="**Homophily Coefficients and Standard Errors from Weighted Exponential Random Graph Models.** (including base-rate effects)"}
+```{r fig-seasonal-ergms, fig.cap="**Homophily Coefficients and Standard Errors from Weighted Exponential Random Graph Models.** (including base-rate effects)"}
 with(
@@
 
-To summarize the results from Figure \@ref(fig:seasonal-ergms) in a single model, we also fit a pooled ERGM across all seasons, including a base-rate effect for each season's trade volume, similarly to the binary ERGM mentioned above.[^modelSpecification] These results are presented in Table \@ref(tab:model1-table-short).
+To summarize the results from Figure [-@fig-seasonal-ergms] in a single model, we also fit a pooled ERGM across all seasons, including a base-rate effect for each season's trade volume, similarly to the binary ERGM mentioned above.[^modelSpecification] These results are presented in Table [-@tbl-model1-table-short].
 In this final specification, we find an aggregate same-division homophily effect (coefficient 0.006, standard error of 0.059, z-statistic of 0.102), that again shows no avoidance pattern.
@@
 
-```{r model1-table-short}
+```{r tbl-model1-table-short}
  withr::with_options(
@@
      knitr::kable(
-       caption = "**Valued ERGM fitted to pooled seasonal data**. The table excludes seasonal constants, complete table is presented in Table \\@ref(tab:model1-table) the Appendix.",
+       caption = "**Valued ERGM fitted to pooled seasonal data**. The table excludes seasonal constants, complete table is presented in Table [-@tbl-model1-table] the Appendix.",
 
@@
 
-```{r model-binary-table}
+```{r tbl-model-binary-table}
 # Model: privaterepo/estimation/ergm-pooled-binary-4
@@
 
-```{r model1-table}
+```{r tbl-model1-table}
 # The model from privaterepo/estimation/ergm-pooled-cmp-1

### 2023-08-07 Manuscript.qmd
Fix table captions and references

@@
 ```{r tbl-dyadic-trade-counts}
+#| tbl-cap: "**Distribution of dyadic trade counts**. Frequency of number of trades in a season in pairs of teams."
 # MB: This is pasted from the calculation in privaterepo/overdispersion.Rmd
@@
   ) %>%
-  knitr::kable(
-    caption = "**Distribution of dyadic trade counts**. Frequency of number of trades in a season in pairs of teams.",    # Sounds convoluted...
-    align = "rrr"
-  )
+  knitr::kable(align = "rrr")
 ```
@@
 The homophily coefficient in the **binary** ERGM fit to all seasonal trade networks simultaneously, including season- and division-specific effects, is equal to 0.012 (standard error of 0.065, z-statistic of 0.188) again showing no avoidance patterns.
-The complete model is presented in Table \@ref(tab:model-binary-table) in the Appendix.
+The complete model is presented in Table [-@tbl-model-binary-table] in the Appendix.
 
@@
 ```{r tbl-model1-table-short}
+#| tbl-cap: "**Valued ERGM fitted to pooled seasonal data**. The table excludes seasonal constants, complete table is presented in Table [-@tbl-model1-table] the Appendix."
 withr::with_options(
@@
  readRDS("model1-table-short.rds") |>
-     knitr::kable(
-       caption = "**Valued ERGM fitted to pooled seasonal data**. The table excludes seasonal constants, complete table is presented in Table [-@tbl-model1-table] the Appendix.",
-
-       align = "lrrrr"
-     )
+   knitr::kable(align = "lrrrr")
 )
@@
 
-It is worth mentioning that the models we have estimated in the results for Figure \@ref(fig:seasonal-ergms) are actually no different from what would have been possible through a generalized linear model predicting the (weighted) number of ties observed between pairs of teams.
+It is worth mentioning that the models we have estimated in the results for Figure [-@fig-seasonal-ergms] are actually no different from what would have been possible through a generalized linear model predicting the (weighted) number of ties observed between pairs of teams.
 However, the value of introducing the weighted ERGM approach with these data is that if subsequent researchers want to take the next step to provide an explanatory account that further examines which teams are likely to trade with which others, this modeling framework would allow for the simultaneous estimation of predictive factors that variously operate at the node level (e.g., is a team that finished lower in the previous season's standings more likely to trade), the dyad level (e.g., are teams with higher total salaries more likely to trade with other teams who have lower total salaries), or structural features above the dyad (e.g., the tendency to form trade-triples or larger configurations through multi-team deals).
@@
 ```{r tbl-model-binary-table}
+#| tbl-cap: "Full results for the binary ERG model fit to pooled data."
+
 # Model: privaterepo/estimation/ergm-pooled-binary-4
@@
     structure(list(Effect = c("*Seasonal base rates*", "1977-1978",  "1978-1979", "1979-1980", "1980-1981", "1981-1982", "1982-1983",  "1983-1984", "1984-1985", "1985-1986", "1986-1987", "1987-1988",  "1988-1989", "1989-1990", "1990-1991", "1991-1992", "1992-1993",  "1993-1994", "1994-1995", "1995-1996", "1996-1997", "1997-1998",  "1998-1999", "1999-2000", "2000-2001", "2001-2002", "2002-2003",  "2003-2004", "2004-2005", "2005-2006", "2006-2007", "2007-2008",  "2008-2009", "2009-2010", "2010-2011", "2011-2012", "2012-2013",  "2013-2014", "2014-2015", "2015-2016", "2016-2017", "2017-2018",  "2018-2019", "Atlantic division", "Central division", "Midwest division",  "Northwest division", "Pacific division", "Southeast division",  "Southwest division", "*Homophily*"), Estimate = c(NA, 0.0644405709809631,  0.00191183013592781, 0.00191183013587444, -0.0510792301762593,  -0.00301475598725536, 0.11151751406511, 0.0273160881984102, -0.120873378851605,  -0.241797069713423, -0.034689299944754, -0.102736047536669, -0.217649092501239,  -0.30417868994459, -0.221517869773192, -0.423638103944277, -0.306766674453116,  -0.38207146557124, -0.492148757048376, -0.368437716751424, -0.351709490768931,  -0.21888283796225, -0.335437019204636, -0.38565088841675, -0.109253481582016,  -0.15570959635153, -0.351709490778105, -0.192860596693431, -0.187248236068849,  -0.22034655957727, -0.373121806940009, -0.33128353096099, -0.115852230243608,  -0.0873886627773311, -0.06903205140053, -0.304889898702513, -0.106234994930531,  -0.106234994930528, 0.0163922227340902, -0.18724823606885, -0.243383258589619,  -0.125604821725264, -0.0600276821843626, -0.801604197626109,  -0.866780460779051, -0.869801998775562, -0.71676022773136, -0.912537765260793,  -0.861071612165831, -0.665077455419437, 0.0122759266186398),      SE = c(NA, 0.127030461420487, 0.129814083679627, 0.129814092249876,      0.129291411629841, 0.12715500781249, 0.122756935416747, 0.125898331758024,      0.132729817906103, 0.139724669892161, 0.128542414167505,      0.131796473203285, 0.132010767728565, 0.131420860718578,      0.127137016042545, 0.138568763749447, 0.131454151198191,      0.135887273703477, 0.143380183694574, 0.130029601322226,      0.129122811845738, 0.122743181611443, 0.128263950828193,      0.130988252937235, 0.118460375222509, 0.120175658066723,      0.129122966384354, 0.12165157379233, 0.121780823545216, 0.122921411086859,      0.129126929340387, 0.127263407924174, 0.119539607693635,      0.118724498302347, 0.118221222501676, 0.126153907222097,      0.119259391114893, 0.119259390545633, 0.11609615644012, 0.121780823683373,      0.123755575999076, 0.119828869972253, 0.117980604919478,      0.0989169031985135, 0.0981917740131896, 0.100979355062305,      0.119436809946451, 0.0992411986882728, 0.12109213666063,      0.118908017386702, 0.0651683403015203), "95% CI" = c(NA,      "(-0.18;  0.31)", "(-0.25;  0.26)", "(-0.25;  0.26)", "(-0.30;  0.20)",      "(-0.25;  0.25)", "(-0.13;  0.35)", "(-0.22;  0.27)", "(-0.38;  0.14)",      "(-0.52;  0.03)", "(-0.29;  0.22)", "(-0.36;  0.16)", "(-0.48;  0.04)",      "(-0.56; -0.05)", "(-0.47;  0.03)", "(-0.70; -0.15)", "(-0.56; -0.05)",      "(-0.65; -0.12)", "(-0.77; -0.21)", "(-0.62; -0.11)", "(-0.60; -0.10)",      "(-0.46;  0.02)", "(-0.59; -0.08)", "(-0.64; -0.13)", "(-0.34;  0.12)",      "(-0.39;  0.08)", "(-0.60; -0.10)", "(-0.43;  0.05)", "(-0.43;  0.05)",      "(-0.46;  0.02)", "(-0.63; -0.12)", "(-0.58; -0.08)", "(-0.35;  0.12)",      "(-0.32;  0.15)", "(-0.30;  0.16)", "(-0.55; -0.06)", "(-0.34;  0.13)",      "(-0.34;  0.13)", "(-0.21;  0.24)", "(-0.43;  0.05)", "(-0.49;  0.00)",      "(-0.36;  0.11)", "(-0.29;  0.17)", "(-1.00; -0.61)", "(-1.06; -0.67)",      "(-1.07; -0.67)", "(-0.95; -0.48)", "(-1.11; -0.72)", "(-1.10; -0.62)",      "(-0.90; -0.43)", "(-0.12;  0.14)"), "p-value" = c(NA, 0.612,      0.988, 0.988, 0.693, 0.981, 0.364, 0.828, 0.362, 0.084, 0.787,      0.436, 0.099, 0.021, 0.081, 0.002, 0.02, 0.005, 0.001, 0.005,      0.006, 0.075, 0.009, 0.003, 0.356, 0.195, 0.006, 0.113, 0.124,      0.073, 0.004, 0.009, 0.332, 0.462, 0.559, 0.016, 0.373, 0.373,      0.888, 0.124, 0.049, 0.295, 0.611, 0, 0, 0, 0, 0, 0, 0, 0.851     )), class = c("tbl_df", "tbl", "data.frame"), row.names = c(NA,  -51L)) |>
-      knitr::kable(
-        caption = "Full results for the binary ERG model fit to pooled data.",
-        align = "lrrrr"
-      )
+      knitr::kable(align = "lrrrr")
   })
@@
 ```{r tbl-model1-table}
+#| tbl-cap: "Full results for the weighted ERG model (Conway-Maxwell-Poisson reference) fit to pooled data."
+
 # The model from privaterepo/estimation/ergm-pooled-cmp-1
@@
     structure(list(Effect = c("CMP", "1976-1977", "1977-1978", "1978-1979",  "1979-1980", "1980-1981", "1981-1982", "1982-1983", "1983-1984",  "1984-1985", "1985-1986", "1986-1987", "1987-1988", "1988-1989",  "1989-1990", "1990-1991", "1991-1992", "1992-1993", "1993-1994",  "1994-1995", "1995-1996", "1996-1997", "1997-1998", "1998-1999",  "1999-2000", "2000-2001", "2001-2002", "2002-2003", "2003-2004",  "2004-2005", "2005-2006", "2006-2007", "2007-2008", "2008-2009",  "2009-2010", "2010-2011", "2011-2012", "2012-2013", "2013-2014",  "2014-2015", "2015-2016", "2016-2017", "2017-2018", "2018-2019",  "*Homophily*"), Estimate = c(0.231584299590497, -0.890763342957112,  -0.83536651253216, -0.916362759460457, -0.871853243700118, -0.987841290853209,  -0.900902605442105, -0.800007214754898, -0.899560461864575, -1.0483260921955,  -1.07751539263132, -0.913653500983868, -0.937433706117321, -1.11513540560325,  -1.16287289821496, -1.09836829500166, -1.32117548878016, -1.20705371653326,  -1.28771790721094, -1.39250323356208, -1.22250163107761, -1.22000987502278,  -1.11156405118103, -1.25007055148395, -1.27514745462814, -1.03847047606852,  -1.05882132141435, -1.26269742772427, -1.08833699875934, -1.0469834704777,  -1.07289680849128, -1.1929728670501, -1.1562207254889, -0.961040860349278,  -0.956076839003664, -0.925824485618014, -1.14194414800718, -0.971701230551129,  -0.971414447588866, -0.82982643075161, -1.01829169839617, -1.07774381829707,  -0.999993419528592, -0.912020178092013, 0.00602638250203989),      SE = c(0.12203512128938, 0.0786659348166571, 0.0749602274310118,      0.0789621568846716, 0.0760022797170869, 0.0849915193088094,      0.0765183664206582, 0.069324091908689, 0.0754443838436976,      0.0885417986944042, 0.0913748344030373, 0.0767335950520064,      0.081705154479656, 0.0874539689668034, 0.0838134623779419,      0.0794202098916715, 0.103689830592027, 0.0889138426500533,      0.0980046218251252, 0.105920044602825, 0.0836029172158815,      0.080937063964129, 0.0737605330112484, 0.0872142023820389,      0.0904111796177795, 0.0712261498735629, 0.0719333510361581,      0.0885909751979331, 0.0733621462658981, 0.067206171621419,      0.0674913276175189, 0.0767037480828984, 0.0768718796054209,      0.0627954612107514, 0.06156685684603, 0.0602325458811699,      0.0751034750003797, 0.0636099818174729, 0.0642021297111841,      0.0536773940981808, 0.0669347619251306, 0.0717645838325605,      0.0660829455575422, 0.0579577186366169, 0.0593486673413606     ), "95% CI" = c("(-0.01;  0.47)", "(-1.04; -0.74)", "(-0.98; -0.69)",      "(-1.07; -0.76)", "(-1.02; -0.72)", "(-1.15; -0.82)", "(-1.05; -0.75)",      "(-0.94; -0.66)", "(-1.05; -0.75)", "(-1.22; -0.87)", "(-1.26; -0.90)",      "(-1.06; -0.76)", "(-1.10; -0.78)", "(-1.29; -0.94)", "(-1.33; -1.00)",      "(-1.25; -0.94)", "(-1.52; -1.12)", "(-1.38; -1.03)", "(-1.48; -1.10)",      "(-1.60; -1.18)", "(-1.39; -1.06)", "(-1.38; -1.06)", "(-1.26; -0.97)",      "(-1.42; -1.08)", "(-1.45; -1.10)", "(-1.18; -0.90)", "(-1.20; -0.92)",      "(-1.44; -1.09)", "(-1.23; -0.94)", "(-1.18; -0.92)", "(-1.21; -0.94)",      "(-1.34; -1.04)", "(-1.31; -1.01)", "(-1.08; -0.84)", "(-1.08; -0.84)",      "(-1.04; -0.81)", "(-1.29; -0.99)", "(-1.10; -0.85)", "(-1.10; -0.85)",      "(-0.94; -0.72)", "(-1.15; -0.89)", "(-1.22; -0.94)", "(-1.13; -0.87)",      "(-1.03; -0.80)", "(-0.11;  0.12)"), "p-value" = c(0.058,      0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,      0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,      0, 0, 0, 0, 0, 0.919)), row.names = c(NA, -45L), class = "data.frame") |>
-    knitr::kable(
-      caption = "Full results for the weighted ERG model (Conway-Maxwell-Poisson reference) fit to pooled data.",
-      align = "lrrrr")
+    knitr::kable(align = "lrrrr")
   })

### 2023-08-07 Manuscript.qmd
Number sections

@@
     cite-method: biblatex
+    number-sections: true
     journal: nws

### 2023-08-07 Manuscript.qmd
References, acknowledgement and appendices
Complete affiliations, keywords, competing interests

--- 2023-08-07 Manuscript.qmd
@@
     orcid: "0000-0001-5043-1149"
-    affiliations: "University of Colorado Denver, USA"
+    email: "JIMI.ADAMS@ucdenver.edu"
+    affiliations: 
+      - name: "University of Colorado Denver"
+        city: "Denver, CO"
+        country: USA
+        postal-code: "80217-3364"
   - name: Michał Bojanowski
@@
     affiliations:
-      - name: "Autonomous University of Barcelona, Barcelona, Spain"
-      - name: "Kozminski University, Warsaw, Poland"
+      - name: "Autonomous University of Barcelona"
+        city: "Bellaterra (Barcelona)"
+        country: Spain
+        postal-code: "08-193"
+      - name: "Kozminski University"
+        city: Warsaw
+        country: Poland
+        postal-code: "03-301"
 date: "May 10, 2022"
@@
 We find limited-to-no support for the popular perception."
+keywords: [trade networks, nba, sports, exponential random graph models, weighted networks]
 format: 
@@
 
-## Details of model specification {.appendix}
+\begin{acknowledgement}
+We appreciate feedback we received on previous iterations of this work from Skye Bender-deMoll, Pavel Krivitsky and David Schaefer.
+\end{acknowledgement}
+
+\paragraph{Competing interests} None.
+
+# References {.unnumbered}
+
+::: {#refs}
+:::
+
+\appendix
+
+# Details of model specification {.appendix}
 
@@
 
-## Binary ERGM model fit {.appendix}
+# Binary ERGM model fit {.appendix}
 
@@
 
-## Weighted ERGM model fit {.appendix}
+# Weighted ERGM model fit {.appendix}
 
@@
 ```
-
-## Acknowledgements {.appendix}
-
-We appreciate feedback we received on previous iterations of this work from Skye Bender-deMoll, Pavel Krivitsky and David Schaefer.

### 2023-08-07 Manuscript.qmd
Fix competing interests
format keywords properly

@@
 We find limited-to-no support for the popular perception."
-keywords: [trade networks, nba, sports, exponential random graph models, weighted networks]
+keywords: ["trade networks", "NBA", "sports", "Exponential-family Random Graph Models", "weighted networks"]
 format: 
@@
 
-\paragraph{Competing interests} None.
+Competing interests. None.
 

### 2023-08-07 Manuscript.qmd
Network movie suppressed

@@
 
-<!-- Manually wrap in distill HTML tags to get a proper caption. -->
-<div class="figure">
-```{r animation, echo=FALSE}
+```{r fig-network-movie, eval=FALSE}
+#| fig-cap: "**Dynamic Visualization of Annual Trade Networks** Availabe in the online supplement.
+Note: Intra-division trades are highlighted in red.
+Node color and position indicates division.
+The upper-right menu allows the animation speed to be adjusted while visualizing the entire sequence, or the forward/back buttons at the bottom can be used to step-wise animate the visualization one annual slice (trade season) at a time."
+if(knitr::is_html_output()) {
+  cat("<div class='figure'>\n")
   htmltools::tags$iframe(src = 'NBA_Trades.html', width = '80%', height = '600px')
-```
-<p class="caption">Video 1.
+  cat('<p class="caption">Video 1.
 **Dynamic Visualization of Annual Trade Networks** Note: Intra-division trades are highlighted in red.
@@
 The upper-right menu allows the animation speed to be adjusted while visualizing the entire sequence, or the forward/back buttons at the bottom can be used to step-wise animate the visualization one annual slice (trade season) at a time.</p>
-</div>
+</div>\n')
+} else {
+  knitr::include_graphics("movie-still.png")
+}
+```
 
